



samples are here

/grid/gillis/home/hover/work/cmxci.paired/Sus_scrofa
/grid/gillis/home/hover/work/cmxci.single/Sus_scrofa



```{r}
library(ggplot2)
library(ggExtra)
library(stringr)
library(parallel)
library(dplyr)
library(ggridges)
library(MetBrewer)

```

```{r}

processed_files = list.files('/grid/gillis/home/hover/work/cmxci.paired/Sus_scrofa')
wasp_vcf_files_paired = processed_files[grepl('.wasp.4.snps_filtered.vcf',processed_files  )]
#Ignore the .vcf.idx files
wasp_vcf_files_paired = wasp_vcf_files_paired[!grepl('.idx',wasp_vcf_files_paired )]
wasp_wig_files_paired = processed_files[grepl('.wasp.4.split.filtered.wig',processed_files  )]


length(wasp_vcf_files_paired)
length(wasp_wig_files_paired)
#There's one mismatched file in the vcfs
#vcf_temp = sapply(strsplit(wasp_vcf_files_paired, split = '.', fixed = T), '[[', 1)
#wig_temp = sapply(strsplit(wasp_wig_files_paired, split = '.', fixed = T), '[[', 1)

#missing_index = which(vcf_temp %in% wig_temp)
#wasp_vcf_files_paired = wasp_vcf_files_paired[missing_index]

#length(wasp_vcf_files_paired)
#length(wasp_wig_files_paired)


processed_files = list.files('/grid/gillis/home/hover/work/cmxci.single/Sus_scrofa')
wasp_vcf_files_single = processed_files[grepl('.wasp.4.snps_filtered.vcf',processed_files  )]
#Ignore the .vcf.idx files
wasp_vcf_files_single = wasp_vcf_files_single[!grepl('.idx',wasp_vcf_files_single )]
wasp_wig_files_single = processed_files[grepl('.wasp.4.split.filtered.wig',processed_files  )]

length(wasp_vcf_files_single)
length(wasp_wig_files_single)


wig_samples_4 = sapply(strsplit(wasp_wig_files_paired, split = '.', fixed = T), '[[', 1)
wig_samples_4 = c(wig_samples_4, sapply(strsplit(wasp_wig_files_single, split = '.', fixed = T), '[[', 1) ) #Add the single samples 
vcf_samples_4 = sapply(strsplit(wasp_vcf_files_paired, split = '.', fixed = T), '[[', 1)
vcf_samples_4 = c(vcf_samples_4, sapply(strsplit(wasp_vcf_files_single, split = '.', fixed = T), '[[', 1) )


table(wig_samples_4 == vcf_samples_4)


wasp_vcf_files_paired = paste('/grid/gillis/home/hover/work/cmxci.paired/Sus_scrofa/',wasp_vcf_files_paired, sep = '')
wasp_vcf_files_single = paste('/grid/gillis/home/hover/work/cmxci.single/Sus_scrofa/',wasp_vcf_files_single, sep = '')

wasp_wig_files_paired = paste('/grid/gillis/home/hover/work/cmxci.paired/Sus_scrofa/',wasp_wig_files_paired, sep = '')
wasp_wig_files_single = paste('/grid/gillis/home/hover/work/cmxci.single/Sus_scrofa/',wasp_wig_files_single, sep = '')

wasp_vcf_files_chr4 = wasp_vcf_files_paired
wasp_wig_files_chr4 = wasp_wig_files_paired

wasp_vcf_files_chr4 = c(wasp_vcf_files_paired, wasp_vcf_files_single)
wasp_wig_files_chr4 = c(wasp_wig_files_paired, wasp_wig_files_single)


length(wasp_vcf_files_chr4)

```


```{r}

processed_files = list.files('/grid/gillis/home/hover/work/cmxci.paired/Sus_scrofa')
wasp_vcf_files_paired = processed_files[grepl('.wasp.7.snps_filtered.vcf',processed_files  )]
#Ignore the .vcf.idx files
wasp_vcf_files_paired = wasp_vcf_files_paired[!grepl('.idx',wasp_vcf_files_paired )]
wasp_wig_files_paired = processed_files[grepl('.wasp.7.split.filtered.wig',processed_files  )]

length(wasp_vcf_files_paired)
length(wasp_wig_files_paired)

head(wasp_vcf_files_paired)

processed_files = list.files('/grid/gillis/home/hover/work/cmxci.single/Sus_scrofa')
wasp_vcf_files_single = processed_files[grepl('.wasp.7.snps_filtered.vcf',processed_files  )]
#Ignore the .vcf.idx files
wasp_vcf_files_single = wasp_vcf_files_single[!grepl('.idx',wasp_vcf_files_single )]
wasp_wig_files_single = processed_files[grepl('.wasp.7.split.filtered.wig',processed_files  )]

length(wasp_vcf_files_single)
length(wasp_wig_files_single)


wig_samples_7 = sapply(strsplit(wasp_wig_files_paired, split = '.', fixed = T), '[[', 1)
wig_samples_7 = c(wig_samples_7, sapply(strsplit(wasp_wig_files_single, split = '.', fixed = T), '[[', 1) ) #Add the single samples 
vcf_samples_7 = sapply(strsplit(wasp_vcf_files_paired, split = '.', fixed = T), '[[', 1)
vcf_samples_7 = c(vcf_samples_7, sapply(strsplit(wasp_vcf_files_single, split = '.', fixed = T), '[[', 1) )
table(wig_samples_7 == vcf_samples_7)


wasp_vcf_files_paired = paste('/grid/gillis/home/hover/work/cmxci.paired/Sus_scrofa/',wasp_vcf_files_paired, sep = '')
wasp_vcf_files_single = paste('/grid/gillis/home/hover/work/cmxci.single/Sus_scrofa/',wasp_vcf_files_single, sep = '')

wasp_wig_files_paired = paste('/grid/gillis/home/hover/work/cmxci.paired/Sus_scrofa/',wasp_wig_files_paired, sep = '')
wasp_wig_files_single = paste('/grid/gillis/home/hover/work/cmxci.single/Sus_scrofa/',wasp_wig_files_single, sep = '')


wasp_vcf_files_chr7 = wasp_vcf_files_paired
wasp_wig_files_chr7 = wasp_wig_files_paired

wasp_vcf_files_chr7 = c(wasp_vcf_files_paired, wasp_vcf_files_single)
wasp_wig_files_chr7 = c(wasp_wig_files_paired, wasp_wig_files_single)

length(wasp_vcf_files_chr7)
```


Function for merging the wig and vcf info, filtering for heterozygous SNPs with 2 alleles and a minimum of 10 reads per allele

```{r}
source('get_exp_vcf.R')
```



```{r}
Sys.time()
wasp_filt_vcf_chr4 = mclapply(1:length(wasp_vcf_files_chr4), function(i) get_exp_vcf(vcf_file_path = wasp_vcf_files_chr4[i], 
                                                                           wig_file_path = wasp_wig_files_chr4[i]), mc.cores = 10)
Sys.time()

head(wasp_filt_vcf_chr4)
length(wasp_filt_vcf_chr4)


Sys.time()
wasp_filt_vcf_chr7 = mclapply(1:length(wasp_vcf_files_chr7), function(i) get_exp_vcf(vcf_file_path = wasp_vcf_files_chr7[i], 
                                                                           wig_file_path = wasp_wig_files_chr7[i]), mc.cores = 10)
Sys.time()

head(wasp_filt_vcf_chr7)
length(wasp_filt_vcf_chr7)


```





Number of filtered SNPs kept
```{r}

num_snps = vector(mode = 'numeric', length = length(wasp_filt_vcf_chr4))

for(i in 1:length(wasp_filt_vcf_chr4)){
  data = wasp_filt_vcf_chr4[[i]]
  
  if(is.null(dim(data))){
    num_snps[i] = 0
  }else{num_snps[i] = dim(data)[1]}
}

hist(num_snps)


num_snps = vector(mode = 'numeric', length = length(wasp_filt_vcf_chr7))

for(i in 1:length(wasp_filt_vcf_chr7)){
  data = wasp_filt_vcf_chr7[[i]]
  
  if(is.null(dim(data))){
    num_snps[i] = 0
  }else{num_snps[i] = dim(data)[1]}
}

hist(num_snps)


```


Getting the genome annotations


```{r}

pig_gtf = rtracklayer::import('/home/werner/projects/cross_species_XCI/genomes/Sus_scrofa_auto/refseq/annotation.gtf')
pig_gtf = as.data.frame(pig_gtf)

#Genome specific 4 chromosome name
pig_4_chrom_name = readLines('/home/werner/projects/cross_species_XCI/genomes/Sus_scrofa_auto/refseq/4label.txt', n = 1)

pig_4_gtf = pig_gtf[pig_gtf$seqnames == pig_4_chrom_name, ]

pig_4_genenames = unique(pig_4_gtf$gene)
length(pig_4_genenames)

#Genome specific 1 chromosome name
pig_7_chrom_name = readLines('/home/werner/projects/cross_species_XCI/genomes/Sus_scrofa_auto/refseq/7label.txt', n = 1)

pig_7_gtf = pig_gtf[pig_gtf$seqnames == pig_7_chrom_name, ]

pig_7_genenames = unique(pig_7_gtf$gene)
length(pig_7_genenames)
```



Adding gene annotations to each sample


```{r}
pig_4_meta = pig_4_gtf[pig_4_gtf$gbkey == 'Gene',]
pig_4_meta = pig_4_meta[!is.na(pig_4_meta$gbkey), ] #A lot of NA entries in the pig GTF file
pig_4_meta = pig_4_meta[!duplicated(pig_4_meta$gene), ]
gene_starts_4 = pig_4_meta$start
gene_ends_4 = pig_4_meta$end


annot_pig_vcfs_chr4 = wasp_filt_vcf_chr4

for(j in 1:length(annot_pig_vcfs_chr4)){
  test = annot_pig_vcfs_chr4[[j]]
  if(is.null(dim(test))){next}
  
  
  gene_annots = vector(mode = 'character', length = length(test$Pos))
  gene_biotype = vector(mode = 'character', length = length(test$Pos))
  for( i in 1:dim(test)[1]){
  
    test_gene_start = test$Pos[i]
    test_gene_start = test$Pos[i]
    gene_index = findInterval(test_gene_start, gene_starts_4)
    if(gene_index == 0){next}
    
    if(test_gene_start <= gene_ends_4[gene_index]){
      
      gene_annots[i] = pig_4_meta$gene[gene_index]
      gene_biotype[i] = pig_4_meta$gene_biotype[gene_index]
    }
  }
  
  test$gene = gene_annots
  test$gene_biotype = gene_biotype
  annot_pig_vcfs_chr4[[j]] = test
}

head(annot_pig_vcfs_chr4)

```



```{r}

save(annot_pig_vcfs_chr4, file = '/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/chr4/all_snp_vcf_7_11_23.Rdata')
```



```{r}
pig_7_meta = pig_7_gtf[pig_7_gtf$gbkey == 'Gene',]
pig_7_meta = pig_7_meta[!is.na(pig_7_meta$gbkey), ] #A lot of NA entries in the pig GTF file
pig_7_meta = pig_7_meta[!duplicated(pig_7_meta$gene), ]
gene_starts_7 = pig_7_meta$start
gene_ends_7 = pig_7_meta$end


annot_pig_vcfs_chr7 = wasp_filt_vcf_chr7

for(j in 1:length(annot_pig_vcfs_chr7)){
  test = annot_pig_vcfs_chr7[[j]]
  if(is.null(dim(test))){next}
  
  
  gene_annots = vector(mode = 'character', length = length(test$Pos))
  gene_biotype = vector(mode = 'character', length = length(test$Pos))
  for( i in 1:dim(test)[1]){
  
    test_gene_start = test$Pos[i]
    test_gene_start = test$Pos[i]
    gene_index = findInterval(test_gene_start, gene_starts_7)
    if(gene_index == 0){next}
    
    if(test_gene_start <= gene_ends_7[gene_index]){
      
      gene_annots[i] = pig_7_meta$gene[gene_index]
      gene_biotype[i] = pig_7_meta$gene_biotype[gene_index]
    }
  }
  
  test$gene = gene_annots
  test$gene_biotype = gene_biotype
  annot_pig_vcfs_chr7[[j]] = test
}

head(annot_pig_vcfs_chr7)

```


```{r}
save(annot_pig_vcfs_chr7, file = '/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/chr7/all_snp_vcf_7_11_23.Rdata')
```




Looking at global distribution of reference ratios versus read counts for all SNPs
```{r}
all_total_counts = c()
all_ref_ratios = c()

for(i in 1:length(annot_pig_vcfs_chr4)){

  data = annot_pig_vcfs_chr4[[i]]
  if(is.null(dim(data))){next}
  total_counts = data$ref_counts + data$alt_counts
  ref_ratios = data$ref_counts / total_counts
  
  all_total_counts = c(all_total_counts, total_counts)
  all_ref_ratios = c(all_ref_ratios, ref_ratios)

}

length(all_total_counts)


plot(all_ref_ratios, log10(all_total_counts), cex = .5)


all_total_counts = c()
all_ref_ratios = c()

for(i in 1:length(annot_pig_vcfs_chr7)){

  data = annot_pig_vcfs_chr7[[i]]
  if(is.null(dim(data))){next}
  total_counts = data$ref_counts + data$alt_counts
  ref_ratios = data$ref_counts / total_counts
  
  all_total_counts = c(all_total_counts, total_counts)
  all_ref_ratios = c(all_ref_ratios, ref_ratios)

}

length(all_total_counts)


plot(all_ref_ratios, log10(all_total_counts), cex = .5)


```


Looking at gene specific expression 


```{r}

captured_genes = c()
for( i in 1:length(annot_pig_vcfs_chr4)){
  suppressWarnings(if(is.null(dim(annot_pig_vcfs_chr4[[i]]))){next})
  captured_genes = c(captured_genes, annot_pig_vcfs_chr4[[i]]$gene)
}

captured_genes = unique(captured_genes)
captured_genes = captured_genes[captured_genes != ""]

length(captured_genes)
```


```{r}
for(i in 1:100){
  
  current_gene = captured_genes[i]
  
  gene_ref_exp = c()
  for(j in 1:length(annot_pig_vcfs_chr4)){
    suppressWarnings(if(is.null(dim(annot_pig_vcfs_chr4[[j]]))){next})
    gene_index = annot_pig_vcfs_chr4[[j]]$gene == current_gene
    ref_counts = annot_pig_vcfs_chr4[[j]]$ref_counts[gene_index]
    alt_counts = annot_pig_vcfs_chr4[[j]]$alt_counts[gene_index]
    total_counts = ref_counts + alt_counts
    gene_ref_exp = c(gene_ref_exp, ref_counts / total_counts )

  }
  
  
  hist(gene_ref_exp, main = current_gene, xlim = c(0,1), breaks = seq(0,1, .025))  
  
}

```


chr7


```{r}

captured_genes = c()
for( i in 1:length(annot_pig_vcfs_chr7)){
  suppressWarnings(if(is.null(dim(annot_pig_vcfs_chr7[[i]]))){next})
  captured_genes = c(captured_genes, annot_pig_vcfs_chr7[[i]]$gene)
}

captured_genes = unique(captured_genes)
captured_genes = captured_genes[captured_genes != ""]

length(captured_genes)
```


```{r}
for(i in 1:100){
  
  current_gene = captured_genes[i]
  
  gene_ref_exp = c()
  for(j in 1:length(annot_pig_vcfs_chr7)){
    suppressWarnings(if(is.null(dim(annot_pig_vcfs_chr7[[j]]))){next})
    gene_index = annot_pig_vcfs_chr7[[j]]$gene == current_gene
    ref_counts = annot_pig_vcfs_chr7[[j]]$ref_counts[gene_index]
    alt_counts = annot_pig_vcfs_chr7[[j]]$alt_counts[gene_index]
    total_counts = ref_counts + alt_counts
    gene_ref_exp = c(gene_ref_exp, ref_counts / total_counts )

  }
  
  
  hist(gene_ref_exp, main = current_gene, xlim = c(0,1), breaks = seq(0,1, .025))  
  
}

```



chr1

```{r}

captured_SNPs = c()
for( i in 1:length(annot_pig_vcfs_chr4)){
  suppressWarnings(if(is.null(dim(annot_pig_vcfs_chr4[[i]]))){next})
  captured_SNPs = c(captured_SNPs, annot_pig_vcfs_chr4[[i]]$Pos)
}

captured_SNPs = unique(captured_SNPs)
captured_SNPs = captured_SNPs[captured_SNPs != ""]

length(captured_SNPs)
```


```{r}

snp_stat_df_chr4 = data.frame(snp_pos = integer(length(captured_SNPs)), num_samples_present = integer(length(captured_SNPs)), 
                         mean_ref_ratio = integer(length(captured_SNPs)), mean_total_counts = integer(length(captured_SNPs)), sd_total_counts = integer(length(captured_SNPs)))



for(i in 1:length(captured_SNPs)){

  if(i %% 1000 == 0){print(i)}
  current_SNP = captured_SNPs[i]
  
  snp_ref_exp = c()
  snp_total_reads = c()
  for(j in 1:length(annot_pig_vcfs_chr4)){
    suppressWarnings(if(is.null(dim(annot_pig_vcfs_chr4[[j]]))){next})
    gene_index = annot_pig_vcfs_chr4[[j]]$Pos == current_SNP
    ref_counts = annot_pig_vcfs_chr4[[j]]$ref_counts[gene_index]
    alt_counts = annot_pig_vcfs_chr4[[j]]$alt_counts[gene_index]
    total_counts = ref_counts + alt_counts
    snp_ref_exp = c(snp_ref_exp, ref_counts / total_counts )
    snp_total_reads = c(snp_total_reads, total_counts)
    
    
  }
  
  snp_stat_df_chr4$snp_pos[i] = current_SNP
  snp_stat_df_chr4$num_samples_present[i] = length(snp_ref_exp)
  snp_stat_df_chr4$mean_ref_ratio[i] = mean(snp_ref_exp)
  snp_stat_df_chr4$mean_total_counts[i] = mean(snp_total_reads)
  snp_stat_df_chr4$sd_total_counts[i] = sd(snp_total_reads)
}

```



```{r}
hist(snp_stat_df_chr4$mean_ref_ratio, breaks = seq(0,1,.01))
hist(snp_stat_df_chr4$num_samples_present)
par(pty = 's')
plot(snp_stat_df_chr4$mean_ref_ratio,snp_stat_df_chr4$num_samples_present, cex = .25, xlab = 'Mean reference ratio', 
     ylab = '# of samples SNP is detected', main = 'Sus scrofa Het SNPs')
abline(v = .5, col = 'red')
abline(v = .4, col = 'blue')
abline(v = .6, col = 'blue')
```

```{r}


snp_stat_df_chr4$filter = !(snp_stat_df_chr4$mean_ref_ratio < .4 | snp_stat_df_chr4$mean_ref_ratio > .6)
table(snp_stat_df_chr4$filter)


```


Save the SNP dataframe

```{r}

save(snp_stat_df_chr4, file = '/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/chr4/detected_het_snps_df_7_11_23.Rdata')
```

```{r}
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/chr4/detected_het_snps_df_7_11_23.Rdata')

```



chr7

```{r}

captured_SNPs = c()
for( i in 1:length(annot_pig_vcfs_chr7)){
  suppressWarnings(if(is.null(dim(annot_pig_vcfs_chr7[[i]]))){next})
  captured_SNPs = c(captured_SNPs, annot_pig_vcfs_chr7[[i]]$Pos)
}

captured_SNPs = unique(captured_SNPs)
captured_SNPs = captured_SNPs[captured_SNPs != ""]

length(captured_SNPs)
```


```{r}

snp_stat_df_chr7 = data.frame(snp_pos = integer(length(captured_SNPs)), num_samples_present = integer(length(captured_SNPs)), 
                         mean_ref_ratio = integer(length(captured_SNPs)), mean_total_counts = integer(length(captured_SNPs)), sd_total_counts = integer(length(captured_SNPs)))



for(i in 1:length(captured_SNPs)){

  if(i %% 1000 == 0){print(i)}
  current_SNP = captured_SNPs[i]
  
  snp_ref_exp = c()
  snp_total_reads = c()
  for(j in 1:length(annot_pig_vcfs_chr7)){
    suppressWarnings(if(is.null(dim(annot_pig_vcfs_chr7[[j]]))){next})
    gene_index = annot_pig_vcfs_chr7[[j]]$Pos == current_SNP
    ref_counts = annot_pig_vcfs_chr7[[j]]$ref_counts[gene_index]
    alt_counts = annot_pig_vcfs_chr7[[j]]$alt_counts[gene_index]
    total_counts = ref_counts + alt_counts
    snp_ref_exp = c(snp_ref_exp, ref_counts / total_counts )
    snp_total_reads = c(snp_total_reads, total_counts)
    
    
  }
  
  snp_stat_df_chr7$snp_pos[i] = current_SNP
  snp_stat_df_chr7$num_samples_present[i] = length(snp_ref_exp)
  snp_stat_df_chr7$mean_ref_ratio[i] = mean(snp_ref_exp)
  snp_stat_df_chr7$mean_total_counts[i] = mean(snp_total_reads)
  snp_stat_df_chr7$sd_total_counts[i] = sd(snp_total_reads)
}

```



```{r}
hist(snp_stat_df_chr7$mean_ref_ratio, breaks = seq(0,1,.01))
hist(snp_stat_df_chr7$num_samples_present)
par(pty = 's')
plot(snp_stat_df_chr7$mean_ref_ratio,snp_stat_df_chr7$num_samples_present, cex = .25, xlab = 'Mean reference ratio', 
     ylab = '# of samples SNP is detected', main = 'Canis lupus familiaris Het SNPs')
abline(v = .5, col = 'red')
abline(v = .4, col = 'blue')
abline(v = .6, col = 'blue')
```

```{r}


snp_stat_df_chr7$filter = !(snp_stat_df_chr7$mean_ref_ratio < .4 | snp_stat_df_chr7$mean_ref_ratio > .6)
table(snp_stat_df_chr7$filter)


```


Save the SNP dataframe

```{r}

save(snp_stat_df_chr7, file = '/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/chr7/detected_het_snps_df_7_11_23.Rdata')
```

```{r}
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/chr7/detected_het_snps_df_7_11_23.Rdata')

```








Filtering the vcfs for only annotated genes and the max powered SNP per gene
chr4

```{r}
filt_pig_vcfs_chr4 = annot_pig_vcfs_chr4

for(i in 1:length(filt_pig_vcfs_chr4)){
  test = filt_pig_vcfs_chr4[[i]]
  suppressWarnings(if(is.null(dim(test))){next})
  #Get rid of SNPs in nonannotated genes
  test = test[test$gene != "", ]
  suppressWarnings(if(dim(test)[1] == 0){filt_pig_vcfs_chr4[[i]] = NA; next})

  #Get the unique number of genes
  sample_genes = unique(test$gene)
  test_2 = data.frame(matrix(NA, 
                             nrow = length(sample_genes), 
                             ncol = dim(test)[2]))
  colnames(test_2) = colnames(test)
  
  
  for(j in 1:length(sample_genes)){
    
    current_gene_data = test[test$gene == sample_genes[j], ]
    if(dim(current_gene_data)[1] == 1){
      test_2[j, ] = current_gene_data
    }else{
      
      current_tot_exp = current_gene_data$ref_counts + current_gene_data$alt_counts
      test_2[j, ] = current_gene_data[which.max(current_tot_exp), ]
    }
  }

  filt_pig_vcfs_chr4[[i]] = test_2
}

```



save the filtered vcf list

```{r}

save(filt_pig_vcfs_chr4, file = '/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/chr4/max_snp_vcf_7_11_23.Rdata')

```

```{r}
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/chr4/max_snp_vcf_7_11_23.Rdata')

```


chr 7
```{r}
filt_pig_vcfs_chr7 = annot_pig_vcfs_chr7

for(i in 1:length(filt_pig_vcfs_chr7)){
  test = filt_pig_vcfs_chr7[[i]]
  suppressWarnings(if(is.null(dim(test))){next})
  #Get rid of SNPs in nonannotated genes
  test = test[test$gene != "", ]
  suppressWarnings(if(dim(test)[1] == 0){filt_pig_vcfs_chr7[[i]] = NA; next})

  #Get the unique number of genes
  sample_genes = unique(test$gene)
  test_2 = data.frame(matrix(NA, 
                             nrow = length(sample_genes), 
                             ncol = dim(test)[2]))
  colnames(test_2) = colnames(test)
  
  
  for(j in 1:length(sample_genes)){
    
    current_gene_data = test[test$gene == sample_genes[j], ]
    if(dim(current_gene_data)[1] == 1){
      test_2[j, ] = current_gene_data
    }else{
      
      current_tot_exp = current_gene_data$ref_counts + current_gene_data$alt_counts
      test_2[j, ] = current_gene_data[which.max(current_tot_exp), ]
    }
  }

  filt_pig_vcfs_chr7[[i]] = test_2
}

```



save the filtered vcf list

```{r}

save(filt_pig_vcfs_chr7, file = '/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/chr7/max_snp_vcf_7_11_23.Rdata')

```

```{r}
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/chr7/max_snp_vcf_7_11_23.Rdata')

```







```{r}
num_filt_genes = vector(mode = 'numeric', length = length(filt_pig_vcfs_chr4))
for(i in 1:length(filt_pig_vcfs_chr4)){
  
  suppressWarnings(if(is.null(dim(filt_pig_vcfs_chr4[[i]]))){
    num_filt_genes[i] = 0
    next
    
  })
  
  num_filt_genes[i] = sum(filt_pig_vcfs_chr4[[i]]$Pos %in% snp_stat_df_chr4$snp_pos[snp_stat_df_chr4$filter])

}

hist(num_filt_genes)
sum(num_filt_genes >= 10)


num_filt_genes = vector(mode = 'numeric', length = length(filt_pig_vcfs_chr7))
for(i in 1:length(filt_pig_vcfs_chr7)){
  
  suppressWarnings(if(is.null(dim(filt_pig_vcfs_chr7[[i]]))){
    num_filt_genes[i] = 0
    next
    
  })
  
  num_filt_genes[i] = sum(filt_pig_vcfs_chr7[[i]]$Pos %in% snp_stat_df_chr7$snp_pos[snp_stat_df_chr7$filter])

}

hist(num_filt_genes)
sum(num_filt_genes >= 10)

```




look at the SNP distributions, with and without including the potentially reference biased SNPs


```{r}


index_num_samps = which(num_filt_genes >= 30)
length(index_num_samps)


for(i in 1:length(index_num_samps)){

  data = filt_pig_vcfs_chr4[[index_num_samps[i]]]
  
  snp_filter = data$Pos %in% snp_stat_df_chr4$snp_pos[snp_stat_df_chr4$filter]
  
  total_exp = data$ref_counts + data$alt_counts
  ref_ratios_filt = data$ref_counts[snp_filter] / total_exp[snp_filter]
  ref_ratios = data$ref_counts / total_exp

  
  hist(ref_ratios, xlim = c(0,1), breaks = seq(0,1, .025))
  abline(v = .5, col = 'red')
  hist(ref_ratios_filt, xlim = c(0,1), breaks = seq(0,1, .025), main = index_num_samps[i])
  abline(v = .5, col = 'red')

}
```



chr7


```{r}


index_num_samps = which(num_filt_genes >= 30)
length(index_num_samps)


for(i in 1:length(index_num_samps)){

  data = filt_pig_vcfs_chr7[[index_num_samps[i]]]
  
  snp_filter = data$Pos %in% snp_stat_df_chr7$snp_pos[snp_stat_df_chr7$filter]
  
  total_exp = data$ref_counts + data$alt_counts
  ref_ratios_filt = data$ref_counts[snp_filter] / total_exp[snp_filter]
  ref_ratios = data$ref_counts / total_exp

  
  hist(ref_ratios, xlim = c(0,1), breaks = seq(0,1, .025))
  abline(v = .5, col = 'red')
  hist(ref_ratios_filt, xlim = c(0,1), breaks = seq(0,1, .025), main = index_num_samps[i])
  abline(v = .5, col = 'red')

}
```





```{r}


# Functions to fold and unfold ratios 
folded <- function(x) { apply( cbind(x,1-x), 1, max)} 
unfold <- function(x) { c(x,1-x) } 


mle_folded <- function(x){ 
  mus = seq(0.5,1, by = 0.001)
  sigmas = seq(0,0.5, by = 0.01)
  
  mles = sapply(mus, function(mu)
    sapply(sigmas, function(sigma)
      -sum( log(dnorm( x,   mean = mu, sd = sigma )
                + dnorm( x,   mean = 1-mu, sd = sigma ) )  )))
  mles[!is.finite(mles)] = NA
  coefs = which(mles==min(mles, na.rm=T), arr.ind=T)
  return (list( mus[coefs[2]] , sigmas[coefs[1]]))
} 

```


```{r}

est_skew_func = function(list.skew.max, Ns, ratios.max, snp_stat_df, filter_snps = T){
  
  #If there's only 1 or 0 SNPs, don't estimate a skew
  if(Ns <= 0){return(NA)}
  
  #Filter for the good snps
  if(filter_snps){
    snp_index = list.skew.max$Pos %in% snp_stat_df$snp_pos[snp_stat_df$filter]
    #Fold the reference skews
    folded_ref_skews = folded(ratios.max[snp_index])
  }else{
    folded_ref_skews = folded(ratios.max)
  }
  #return the mean and sigma estimates
  return(mle_folded(folded_ref_skews))
}


Ns = vector(mode='numeric', length = length(filt_pig_vcfs_chr4))

for(i in 1:length(filt_pig_vcfs_chr4)){
  
  suppressWarnings(if(is.null(dim(filt_pig_vcfs_chr4[[i]]))){Ns[i] = 0}
  else{Ns[i] = dim(filt_pig_vcfs_chr4[[i]])[1]})
  
}


ratios.max = list()
for(i in 1: length(filt_pig_vcfs_chr4)){
  
  if(Ns[i] == 0){next}
  
  total_exp = filt_pig_vcfs_chr4[[i]]$ref_counts + filt_pig_vcfs_chr4[[i]]$alt_counts
  ref_ratios = filt_pig_vcfs_chr4[[i]]$ref_counts / total_exp
  ratios.max[[i]] = ref_ratios

}



```

```{r}

Sys.time()
folded_norm_fits = mclapply(1:length(filt_pig_vcfs_chr4), function(i) est_skew_func(filt_pig_vcfs_chr4[[i]], Ns[i], ratios.max[[i]], snp_stat_df_chr4), mc.cores = 20)
Sys.time()

Sys.time()
folded_norm_fits_bad_snps = mclapply(1:length(filt_pig_vcfs_chr4), function(i) est_skew_func(filt_pig_vcfs_chr4[[i]], Ns[i], ratios.max[[i]], snp_stat_df_chr4, filter_snps = F), mc.cores = 20)
Sys.time()

#Grab the skew and variance estimates
est_skew_chr4 = vector(mode = 'numeric', length = length(folded_norm_fits))
est_var_chr4 = vector(mode = 'numeric', length = length(folded_norm_fits))
for(i in 1:length(folded_norm_fits)){
  
  if(is.na(folded_norm_fits[i])){
    est_skew_chr4[i] = NA
    est_var_chr4[i] = NA
    next
  }
  
  est_skew_chr4[i] = folded_norm_fits[[i]][[1]]
  est_var_chr4[i] = folded_norm_fits[[i]][[2]]
}

#Grab the skew and variance estimates
est_skew_bad_snps_chr4 = vector(mode = 'numeric', length = length(folded_norm_fits_bad_snps))
est_var_bad_snps_chr4 = vector(mode = 'numeric', length = length(folded_norm_fits_bad_snps))
for(i in 1:length(folded_norm_fits_bad_snps)){
  
  if(is.na(folded_norm_fits_bad_snps[i])){
    est_skew_bad_snps_chr4[i] = NA
    est_var_bad_snps_chr4[i] = NA
    next
  }
  
  est_skew_bad_snps_chr4[i] = folded_norm_fits_bad_snps[[i]][[1]]
  est_var_bad_snps_chr4[i] = folded_norm_fits_bad_snps[[i]][[2]]
}

hist(est_skew_chr4, main = 'pig skew estimates filtering out bad SNPs', breaks=seq(.5, 1, .005))
hist(est_skew_bad_snps_chr4, main = 'pig skew estimates filtering out bad SNPs', breaks=seq(.5, 1, .005))
hist(est_var_chr4, main = 'Variance estimates', breaks = 32)
plot(est_skew_chr4, est_var_chr4)
```


```{r}
par(pty = 's')
plot(est_skew_chr4, est_skew_bad_snps_chr4, cex = .5)
abline(a = 0, b = 1, col = 'red')
```

Make a dataframe from with the estimated skew stats

```{r}

num_good_snps_chr4 = vector(length = length(filt_pig_vcfs_chr4), mode = 'numeric')
total_snps_chr4 = vector(length = length(filt_pig_vcfs_chr4), mode = 'numeric')

for(i in 1:length(filt_pig_vcfs_chr4)){
  
  data = filt_pig_vcfs_chr4[[i]]
  suppressWarnings(if(is.null(dim(data))){num_good_snps_chr4[i] = 0; total_snps_chr4[i] = 0; next})
  
  snp_filter = data$Pos %in% snp_stat_df_chr4$snp_pos[snp_stat_df_chr4$filter]
  
  total_snps_chr4[i] = length(data$Pos)
  num_good_snps_chr4[i] = length(data$Pos[snp_filter])
}

plot(total_snps_chr4, num_good_snps_chr4)

```


```{r}
plot(est_skew_chr4, num_good_snps_chr4)
abline(h = 10, col = 'red')


hist(unfold(est_skew_chr4[num_good_snps_chr4 >= 10]), breaks = seq(0,1,.01), xlab = 'Estimated sample skew', main = 'pig chromosome 4')
```



chr7


```{r}
Ns = vector(mode='numeric', length = length(filt_pig_vcfs_chr7))

for(i in 1:length(filt_pig_vcfs_chr7)){
  
  suppressWarnings(if(is.null(dim(filt_pig_vcfs_chr7[[i]]))){Ns[i] = 0}
  else{Ns[i] = dim(filt_pig_vcfs_chr7[[i]])[1]})
  
}


ratios.max = list()
for(i in 1: length(filt_pig_vcfs_chr7)){
  
  if(Ns[i] == 0){next}
  
  total_exp = filt_pig_vcfs_chr7[[i]]$ref_counts + filt_pig_vcfs_chr7[[i]]$alt_counts
  ref_ratios = filt_pig_vcfs_chr7[[i]]$ref_counts / total_exp
  ratios.max[[i]] = ref_ratios

}


```


```{r}

Sys.time()
folded_norm_fits = mclapply(1:length(filt_pig_vcfs_chr7), function(i) est_skew_func(filt_pig_vcfs_chr7[[i]], Ns[i], ratios.max[[i]], snp_stat_df_chr7), mc.cores = 20)
Sys.time()

Sys.time()
folded_norm_fits_bad_snps = mclapply(1:length(filt_pig_vcfs_chr7), function(i) est_skew_func(filt_pig_vcfs_chr7[[i]], Ns[i], ratios.max[[i]], snp_stat_df_chr7, filter_snps = F), mc.cores = 20)
Sys.time()

#Grab the skew and variance estimates
est_skew_chr7 = vector(mode = 'numeric', length = length(folded_norm_fits))
est_var_chr7 = vector(mode = 'numeric', length = length(folded_norm_fits))
for(i in 1:length(folded_norm_fits)){
  
  if(is.na(folded_norm_fits[i])){
    est_skew_chr7[i] = NA
    est_var_chr7[i] = NA
    next
  }
  
  est_skew_chr7[i] = folded_norm_fits[[i]][[1]]
  est_var_chr7[i] = folded_norm_fits[[i]][[2]]
}

#Grab the skew and variance estimates
est_skew_bad_snps_chr7 = vector(mode = 'numeric', length = length(folded_norm_fits_bad_snps))
est_var_bad_snps_chr7 = vector(mode = 'numeric', length = length(folded_norm_fits_bad_snps))
for(i in 1:length(folded_norm_fits_bad_snps)){
  
  if(is.na(folded_norm_fits_bad_snps[i])){
    est_skew_bad_snps_chr7[i] = NA
    est_var_bad_snps_chr7[i] = NA
    next
  }
  
  est_skew_bad_snps_chr7[i] = folded_norm_fits_bad_snps[[i]][[1]]
  est_var_bad_snps_chr7[i] = folded_norm_fits_bad_snps[[i]][[2]]
}

hist(est_skew_chr7, main = 'pig skew estimates filtering out bad SNPs', breaks=seq(.5, 1, .005))
hist(est_skew_bad_snps_chr7, main = 'pig skew estimates filtering out bad SNPs', breaks=seq(.5, 1, .005))
hist(est_var_chr7, main = 'Variance estimates', breaks = 32)
plot(est_skew_chr7, est_var_chr7)
```

```{r}
par(pty = 's')
plot(est_skew_chr7, est_skew_bad_snps_chr7, cex = .5)
abline(a = 0, b = 1, col = 'red')
```

Make a dataframe from with the estimated skew stats

```{r}

num_good_snps_chr7 = vector(length = length(filt_pig_vcfs_chr7), mode = 'numeric')
total_snps_chr7 = vector(length = length(filt_pig_vcfs_chr7), mode = 'numeric')

for(i in 1:length(filt_pig_vcfs_chr7)){
  
  data = filt_pig_vcfs_chr7[[i]]
  suppressWarnings(if(is.null(dim(data))){num_good_snps_chr7[i] = 0; total_snps_chr7[i] = 0; next})
  
  snp_filter = data$Pos %in% snp_stat_df_chr7$snp_pos[snp_stat_df_chr7$filter]
  
  total_snps_chr7[i] = length(data$Pos)
  num_good_snps_chr7[i] = length(data$Pos[snp_filter])
}

plot(total_snps_chr7, num_good_snps_chr7)

```



```{r}
plot(est_skew_chr7, num_good_snps_chr7)
abline(h = 10, col = 'red')


hist(unfold(est_skew_chr7[num_good_snps_chr7 >= 10]), breaks = seq(0,1,.01), xlab = 'Estimated sample skew', main = 'pig chromosome 3')
```


######################################
Looking at the chromosome 4 samples that are supposedly skewed
What's going on

######################################


```{r}

chr4_skewed_index = which(est_skew_chr4 >= .6 & num_good_snps_chr4 >= 10)

all_chr4_skewed_snps = c()
for(i in 1:length(chr4_skewed_index)){
  
  data = filt_pig_vcfs_chr4[[chr4_skewed_index[i]]]
  #data = data[data$Pos %in% snp_stat_df$snp_pos[snp_stat_df$filter], ]
  ref_ratios = data$ref_counts / (data$ref_counts + data$alt_counts)
  hist(ref_ratios, breaks = seq(0,1,.025), main = sprintf('Sample index: %i est skew: %0.3f',chr4_skewed_index[i], est_skew_chr4[chr4_skewed_index[i]]))
  all_chr4_skewed_snps = c(all_chr4_skewed_snps, ref_ratios)
  
}

hist(all_chr4_skewed_snps, breaks = seq(0,1,.025), main = 'All compiled SNPs')


chr4_Nonskewed_index = which(est_skew_chr4 < .6 & num_good_snps_chr4 >= 10)

all_chr4_Nonskewed_snps = c()
for(i in 1:length(chr4_Nonskewed_index)){
  
  data = filt_pig_vcfs_chr4[[chr4_Nonskewed_index[i]]]
  #data = data[data$Pos %in% snp_stat_df$snp_pos[snp_stat_df$filter], ]
  ref_ratios = data$ref_counts / (data$ref_counts + data$alt_counts)
  all_chr4_Nonskewed_snps = c(all_chr4_Nonskewed_snps, ref_ratios)
  
}

hist(all_chr4_Nonskewed_snps, breaks = seq(0,1,.025), main = 'All compiled SNPs')

```



```{r}
pig_chr4_skew_and_stats_df = data.frame(sample_id = wig_samples_4, 
                                   est_skew = est_skew_chr4, est_var = est_var_chr4,  #Skew estimates filtering out bad genes
                                   est_skew_total_snps = est_skew_bad_snps_chr4, est_var_total_snps = est_var_bad_snps_chr4,   #Skew estimates with all genes 
                                   num_good_snps = num_good_snps_chr4, total_snps = total_snps_chr4)

pig_chr4_skew_and_stats_df
```


```{r}

save(pig_chr4_skew_and_stats_df, file = '/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/chr4/pig_chr4_skew_and_stats_df_7_11_23.Rdata')
```

```{r}

load( '/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/chr4/pig_chr4_skew_and_stats_df_7_11_23.Rdata')

```





##################################

Looking at chromosome 3 now

check if the same samples that were skewed on chromosome 1 anre skewed on chromosome 3

#################################


```{r}

chr7_skewed_index = which(est_skew_chr7 >= .6 & num_good_snps_chr7 >= 10)

all_chr7_skewed_snps = c()
for(i in 1:length(chr7_skewed_index)){
  
  data = filt_pig_vcfs_chr7[[chr7_skewed_index[i]]]
  #data = data[data$Pos %in% snp_stat_df$snp_pos[snp_stat_df$filter], ]
  ref_ratios = data$ref_counts / (data$ref_counts + data$alt_counts)
  hist(ref_ratios, breaks = seq(0,1,.025), main = sprintf('Sample index: %i est skew: %0.3f',chr7_skewed_index[i], est_skew_chr7[chr7_skewed_index[i]]))
  all_chr7_skewed_snps = c(all_chr7_skewed_snps, ref_ratios)
  
}

hist(all_chr7_skewed_snps, breaks = seq(0,1,.025), main = 'All compiled SNPs')


chr7_Nonskewed_index = which(est_skew_chr7 < .6 & num_good_snps_chr7 >= 10)

all_chr7_Nonskewed_snps = c()
for(i in 1:length(chr7_Nonskewed_index)){
  
  data = filt_pig_vcfs_chr7[[chr7_Nonskewed_index[i]]]
  #data = data[data$Pos %in% snp_stat_df$snp_pos[snp_stat_df$filter], ]
  ref_ratios = data$ref_counts / (data$ref_counts + data$alt_counts)
  all_chr7_Nonskewed_snps = c(all_chr7_Nonskewed_snps, ref_ratios)
  
}

hist(all_chr7_Nonskewed_snps, breaks = seq(0,1,.025), main = 'All compiled SNPs')

```


```{r}
pig_chr7_skew_and_stats_df = data.frame(sample_id = wig_samples_7, 
                                   est_skew = est_skew_chr7, est_var = est_var_chr7,  #Skew estimates filtering out bad genes
                                   est_skew_total_snps = est_skew_bad_snps_chr7, est_var_total_snps = est_var_bad_snps_chr7,   #Skew estimates with all genes 
                                   num_good_snps = num_good_snps_chr7, total_snps = total_snps_chr7)

pig_chr7_skew_and_stats_df
```


```{r}

save(pig_chr7_skew_and_stats_df, file = '/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/chr7/pig_chr7_skew_and_stats_df_7_11_23.Rdata')
```

```{r}

load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/chr7/pig_chr7_skew_and_stats_df_7_11_23.Rdata')

```


################################
Compare across the chromosomes and see if the same samples are throwing the skewed estimates
################################



```{r}

chr4_weird_samps = pig_chr4_skew_and_stats_df$sample_id[pig_chr4_skew_and_stats_df$est_skew >= .6 & pig_chr4_skew_and_stats_df$num_good_snps >= 10]

chr7_weird_samps = pig_chr7_skew_and_stats_df$sample_id[pig_chr7_skew_and_stats_df$est_skew >= .6 & pig_chr7_skew_and_stats_df$num_good_snps >= 10]


index = match(pig_chr7_skew_and_stats_df$sample_id, pig_chr4_skew_and_stats_df$sample_id )
#Reorder to match chromosome 7
pig_chr4_skew_and_stats_df = pig_chr4_skew_and_stats_df[index, ]
```


```{r}
par(pty = 's')
plot(pig_chr4_skew_and_stats_df$est_skew, pig_chr7_skew_and_stats_df$est_skew)
abline(a = 0, b = 1, col ='red' )


num_snp_index = pig_chr4_skew_and_stats_df$num_good_snps >= 10 & pig_chr7_skew_and_stats_df$num_good_snps >= 10
plot(pig_chr4_skew_and_stats_df$est_skew[num_snp_index], pig_chr7_skew_and_stats_df$est_skew[num_snp_index], xlab = 'chr4 estimated skew', 
     ylab = 'chr7 estimated skew', main = 'pig autosome skew estimates')
abline(a = 0, b = 1, col ='red' )
```






