

Looking at the current data in lab for cross mammal RNA-seq samples

```{r}
library(dplyr)
library(ggplot2)

```



```{r}

genus_samples_df = read.table(file ='/home/werner/projects/cross_species_XCI/code/all_mammal_metadata.tsv', sep = '\t', header = T, fill = T, quote = "")

dim(genus_samples_df)
head(genus_samples_df)

```

looking for specific mouse studies that used the DO mice

SRP296549
Publication: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10404230/#CR60
Mix of female and male, no metadata that I can find

```{r}


table(genus_samples_df$SRAStudy == 'SRP296549')


study_1_df = genus_samples_df[genus_samples_df$SRAStudy == 'SRP296549', ]
study_1_df

```

SRP267741
Publication: https://www.nature.com/articles/s41467-021-23649-0#Sec12
96 female and 96 male samples, no metadata that I can find
There are multiple SRR-IDs per sample. The sample IDs are in the SampleName column, match up to the GEO IDs, there are 192 of them as expected

```{r}


table(genus_samples_df$SRAStudy == 'SRP267741')


study_2_df = genus_samples_df[genus_samples_df$SRAStudy == 'SRP267741', ]
study_2_df


unique(study_2_df$SampleName)
```


SRP125176
publication: https://academic.oup.com/genetics/article/209/1/335/5931013?login=false
All 188 are annotated as female
```{r}


table(genus_samples_df$SRAStudy == 'SRP125176')


study_3_df = genus_samples_df[genus_samples_df$SRAStudy == 'SRP125176', ]
study_3_df


table(study_3_df$Sex)
```


Currently skipping over the study with the multiple SRRs for the same sample, not sure what is going on there
Just use SRP296549 and SRP125176

```{r}

length(c(study_1_df$Run, study_3_df$Run))
writeLines(c(study_1_df$Run, study_3_df$Run), '/home/werner/projects/cross_species_XCI/code/female_metadata/mus_musculus_female_SRR.txt')



```









```{r}

table(genus_samples_df$Sex)

```




```{r}

sex_index = genus_samples_df$Sex %in% c('female', 'F', 'Female')
plot_temp_df = genus_samples_df[sex_index, ] 
species_count = plot_temp_df %>% group_by(ScientificName) %>% summarise(n())
species_count = species_count[order(species_count[ ,2], decreasing = T), ]
species_count = species_count[!species_count$ScientificName %in% c('Mus musculus'), ]


species_10 = species_count$ScientificName[species_count[ ,2] >=10]
species_100 = species_count$ScientificName[species_count[ ,2] >=100]
species_1000 = species_count$ScientificName[species_count[ ,2] >=1000]

plot_temp_10 =plot_temp_df[plot_temp_df$ScientificName %in% species_10, ] 
plot_temp_10$ScientificName = factor(plot_temp_10$ScientificName, levels = species_10)

plot_temp_100 =plot_temp_df[plot_temp_df$ScientificName %in% species_100, ] 
plot_temp_100$ScientificName = factor(plot_temp_100$ScientificName, levels = species_100)

plot_temp_1000 =plot_temp_df[plot_temp_df$ScientificName %in% species_1000, ] 
plot_temp_1000$ScientificName = factor(plot_temp_1000$ScientificName, levels = species_1000)


ggplot(plot_temp_10, aes(x = ScientificName)) + geom_bar(stat = 'count') +
  ggtitle('Species with >= 10 female samples') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(fill = 'white', linetype = 1),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'annotated_female_10_species.pdf', device = 'pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/metadata_summary/', 
       useDingbats = F, dpi = 300, height = 5, width = 8)


ggplot(plot_temp_100, aes(x = ScientificName)) + geom_bar(stat = 'count') +
  ggtitle('Species with >= 100 female samples') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(fill = 'white', linetype = 1),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'annotated_female_100_species.pdf', device = 'pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/metadata_summary/', 
       useDingbats = F, dpi = 300, height = 5, width = 8)

ggplot(plot_temp_1000, aes(x = ScientificName)) + geom_bar(stat = 'count') +
  ggtitle('Species with >= 1000 female samples') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(fill = 'white', linetype = 1),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'annotated_female_1000_species.pdf', device = 'pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/metadata_summary/', 
       useDingbats = F, dpi = 300, height = 5, width = 8)
```
```{r}
plot_temp_100 %>% group_by(ScientificName) %>% summarize(n = n())

```





```{r}
#Unannotated or mixed sex studies

sex_index = !genus_samples_df$Sex %in% c('female', 'F', 'Female')
plot_temp_df = genus_samples_df[sex_index, ] 
species_count = plot_temp_df %>% group_by(ScientificName) %>% summarise(n())
species_count = species_count[order(species_count[ ,2], decreasing = T), ]
species_count = species_count[!species_count$ScientificName %in% c('Mus musculus'), ]

species_10 = species_count$ScientificName[species_count[ ,2] >=10]
species_100 = species_count$ScientificName[species_count[ ,2] >=100]
species_1000 = species_count$ScientificName[species_count[ ,2] >=1000]

plot_temp_10 =plot_temp_df[plot_temp_df$ScientificName %in% species_10, ] 
plot_temp_10$ScientificName = factor(plot_temp_10$ScientificName, levels = species_10)

plot_temp_100 =plot_temp_df[plot_temp_df$ScientificName %in% species_100, ] 
plot_temp_100$ScientificName = factor(plot_temp_100$ScientificName, levels = species_100)

plot_temp_1000 =plot_temp_df[plot_temp_df$ScientificName %in% species_1000, ] 
plot_temp_1000$ScientificName = factor(plot_temp_1000$ScientificName, levels = species_1000)


ggplot(plot_temp_10, aes(x = ScientificName)) + geom_bar(stat = 'count') +
  ggtitle('Species with >= 10 unannotated samples') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(fill = 'white', linetype = 1),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'unannotated_10_species.pdf', device = 'pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/metadata_summary/', 
       useDingbats = F, dpi = 300, height = 5, width = 10)


ggplot(plot_temp_100, aes(x = ScientificName)) + geom_bar(stat = 'count') +
  ggtitle('Species with >= 100 unannotated samples') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(fill = 'white', linetype = 1),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'unannotated_100_species.pdf', device = 'pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/metadata_summary/', 
       useDingbats = F, dpi = 300, height = 5, width = 8)


ggplot(plot_temp_1000, aes(x = ScientificName)) + geom_bar(stat = 'count') +
  ggtitle('Species with >= 1000 unannotated samples') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(fill = 'white', linetype = 1),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
ggsave(filename = 'unannotated_1000_species.pdf', device = 'pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/metadata_summary/', 
       useDingbats = F, dpi = 300, height = 5, width = 8)
```








```{r}

table(genus_samples_df$ScientificName[sex_index])[table(genus_samples_df$ScientificName[sex_index]) >= 10]


```

```{r}

species_10

tab_obj = table(genus_samples_df$ScientificName[genus_samples_df$ScientificName %in% species_10])
tab_obj
```




Save lists of female samples for some species to start with


Check the metadata

```{r}


genus_samples_df = read.table(file ='/home/werner/projects/cross_species_XCI/code/all_mammal_metadata.tsv', sep = '\t', header = T, fill = T, quote = "")

dim(genus_samples_df)
head(genus_samples_df)

```


```{r}
#Five with the most annotated female samples
cow_meta = genus_samples_df[genus_samples_df$ScientificName == 'Bos taurus', ]
pig_meta = genus_samples_df[genus_samples_df$ScientificName == 'Sus scrofa', ]
macaca_meta = genus_samples_df[genus_samples_df$ScientificName == 'Macaca mulatta', ]
goat_meta = genus_samples_df[genus_samples_df$ScientificName == 'Ovis aries', ]
rat_meta = genus_samples_df[genus_samples_df$ScientificName == 'Rattus norvegicus', ]

#species with fewer samples
monodelphis_meta = genus_samples_df[genus_samples_df$ScientificName == 'Monodelphis domestica', ]
capra_meta = genus_samples_df[genus_samples_df$ScientificName == 'Capra hircus', ]
canis_meta = genus_samples_df[genus_samples_df$ScientificName == 'Canis lupus familiaris', ]
equus_meta = genus_samples_df[genus_samples_df$ScientificName == 'Equus caballus', ]
rabbit_meta = genus_samples_df[genus_samples_df$ScientificName == 'Oryctolagus cuniculus', ]

dim(cow_meta)
dim(pig_meta)
dim(macaca_meta)
dim(goat_meta)
dim(rat_meta)
dim(monodelphis_meta)
dim(capra_meta)
dim(canis_meta)
dim(equus_meta)
dim(rabbit_meta)
```

Check the metadata of the test samples so far

```{r}

cow_meta[cow_meta$Run %in% cow_samples, ]

```


```{r}
table(cow_meta$Sex)

table(pig_meta$Sex)

table(macaca_meta$Sex)

table(goat_meta$Sex)

table(rat_meta$Sex)
```


Get the samples that have female annotations
```{r}

female_srr = cow_meta[cow_meta$Sex == 'female', "Run" ]
length(female_srr)
writeLines(female_srr, '/home/werner/projects/cross_species_XCI/code/female_metadata/bos_taurus_female_SRR.txt')

female_srr = pig_meta[pig_meta$Sex == 'female', "Run" ]
length(female_srr)
writeLines(female_srr, '/home/werner/projects/cross_species_XCI/code/female_metadata/sus_scrofa_female_SRR.txt')

female_srr = macaca_meta[macaca_meta$Sex == 'female', "Run" ]
length(female_srr)
writeLines(female_srr, '/home/werner/projects/cross_species_XCI/code/female_metadata/macaca_mulatta_female_SRR.txt')

female_srr = goat_meta[goat_meta$Sex == 'female', "Run" ]
length(female_srr)
writeLines(female_srr, '/home/werner/projects/cross_species_XCI/code/female_metadata/ovis_aries_female_SRR.txt')

female_srr = rat_meta[rat_meta$Sex == 'female', "Run" ]
length(female_srr)
writeLines(female_srr, '/home/werner/projects/cross_species_XCI/code/female_metadata/rattus_norvegicus_female_SRR.txt')

```

Get female samples for some of the other species
```{r}
plot_temp_100 %>% group_by(ScientificName) %>% summarize(n = n())

```

```{r}
female_srr = monodelphis_meta[monodelphis_meta$Sex == 'female', "Run" ]
length(female_srr)
writeLines(female_srr, '/home/werner/projects/cross_species_XCI/code/female_metadata/Monodelphis_domestica_female_SRR.txt')

female_srr = capra_meta[capra_meta$Sex == 'female', "Run" ]
length(female_srr)
writeLines(female_srr, '/home/werner/projects/cross_species_XCI/code/female_metadata/Capra_hircus_female_SRR.txt')

female_srr = canis_meta[canis_meta$Sex == 'female', "Run" ]
length(female_srr)
writeLines(female_srr, '/home/werner/projects/cross_species_XCI/code/female_metadata/Canis_lupus_familiaris_female_SRR.txt')

female_srr = equus_meta[equus_meta$Sex == 'female', "Run" ]
length(female_srr)
writeLines(female_srr, '/home/werner/projects/cross_species_XCI/code/female_metadata/Equus_caballus_female_SRR.txt')

female_srr = rabbit_meta[rabbit_meta$Sex == 'female', "Run" ]
length(female_srr)
writeLines(female_srr, '/home/werner/projects/cross_species_XCI/code/female_metadata/Oryctolagus_cuniculus_female_SRR.txt')



```


Get some of the nonannotated samples for macaca, going to try and screen some of them for additional female samples
```{r}
macaca_meta[macaca_meta$Sex != 'female',  ]

unannot_macaca_meta = filter(macaca_meta, Sex != 'female' & Tumor == 'no' & Consent == 'public' & LibraryStrategy == 'RNA-Seq' & LibrarySource == 'TRANSCRIPTOMIC' & 
         LibraryLayout == 'PAIRED' & size_MB != 0) %>%
  arrange(size_MB)

unannot_macaca_meta

table(unannot_macaca_meta[ 1:500, ]$SRAStudy)


female_srr = unannot_macaca_meta[ 1:500, ]$Run
length(female_srr)
writeLines(female_srr, '/home/werner/projects/cross_species_XCI/code/female_metadata/unannot_macaca_mulatta_1_500_SRR.txt')


table(unannot_macaca_meta[ 501:1000, ]$SRAStudy)

female_srr = unannot_macaca_meta[ 501:1000, ]$Run
length(female_srr)
writeLines(female_srr, '/home/werner/projects/cross_species_XCI/code/female_metadata/unannot_macaca_mulatta_501_1000_SRR.txt')

table(unannot_macaca_meta[ 1001:1500, ]$SRAStudy)

female_srr = unannot_macaca_meta[ 1001:1500, ]$Run
length(female_srr)
writeLines(female_srr, '/home/werner/projects/cross_species_XCI/code/female_metadata/unannot_macaca_mulatta_1001_1500_SRR.txt')


table(unannot_macaca_meta[ 1501:2500, ]$SRAStudy)

female_srr = unannot_macaca_meta[ 1501:2500, ]$Run
length(female_srr)
writeLines(female_srr, '/home/werner/projects/cross_species_XCI/code/female_metadata/unannot_macaca_mulatta_1501_2500_SRR.txt')



```


Save a smaller metadata file of the samples we processed

```{r}


keep_species = c('Bos taurus','Sus scrofa','Macaca mulatta','Ovis aries','Rattus norvegicus','Capra hircus', 'Canis lupus familiaris', 'Equus caballus')
keep_species_meta = genus_samples_df %>% filter(ScientificName %in% keep_species & Sex == 'female')

mouse_keep_meta = rbind(genus_samples_df %>% filter(SRAStudy == 'SRP125176'), genus_samples_df %>% filter(SRAStudy == 'SRP296549'))

all_keep_species_meta = rbind(keep_species_meta, mouse_keep_meta)
all_keep_species_meta

save(all_keep_species_meta, file = '/home/werner/projects/cross_species_XCI/final_plots/R/data_for_plots/all_keep_species_meta.Rdata')
write.csv(all_keep_species_meta, file = '/home/werner/projects/cross_species_XCI/final_plots/R/data_for_plots/all_keep_species_meta.csv')

```









Check out the metadata available for the samples we're looking at

```{r}

cow_samples = read.table('/home/werner/projects/cross_species_XCI/code/female_metadata/bos_taurus_female_SRR.txt')
cow_samples = cow_samples[ ,1]

cow_metadata = genus_samples_df %>% filter(Run %in% cow_samples)



ggplot(cow_metadata, aes(x = factor(Tumor))) + geom_bar(stat="count", width=0.7)
ggplot(cow_metadata, aes(x = factor(Disease))) + geom_bar(stat="count", width=0.7)
ggplot(cow_metadata, aes(x = factor(Histological_Type))) + geom_bar(stat="count", width=0.7)
ggplot(cow_metadata, aes(x = factor(Body_Site))) + geom_bar(stat="count", width=0.7)

table(cow_metadata$Body_Site)

#105 separate studies for cow
length(table(cow_metadata$SRAStudy))
```

```{r}

rat_samples = read.table('/home/werner/projects/cross_species_XCI/code/female_metadata/rattus_norvegicus_female_SRR.txt')
rat_samples = rat_samples[ ,1]

rat_metadata = genus_samples_df %>% filter(Run %in% rat_samples)

ggplot(rat_metadata, aes(x = factor(Tumor))) + geom_bar(stat="count", width=0.7)
ggplot(rat_metadata, aes(x = factor(Disease))) + geom_bar(stat="count", width=0.7)
ggplot(rat_metadata, aes(x = factor(Histological_Type))) + geom_bar(stat="count", width=0.7)
ggplot(rat_metadata, aes(x = factor(Body_Site))) + geom_bar(stat="count", width=0.7)
table(rat_metadata$Body_Site)

#68 separate studies for rat
length(table(rat_metadata$SRAStudy))
```


```{r}

pig_samples = read.table('/home/werner/projects/cross_species_XCI/code/female_metadata/sus_scrofa_female_SRR.txt')
pig_samples = pig_samples[ ,1]

pig_metadata = genus_samples_df %>% filter(Run %in% pig_samples)

ggplot(pig_metadata, aes(x = factor(Tumor))) + geom_bar(stat="count", width=0.7)
ggplot(pig_metadata, aes(x = factor(Disease))) + geom_bar(stat="count", width=0.7)
ggplot(pig_metadata, aes(x = factor(Histological_Type))) + geom_bar(stat="count", width=0.7)
ggplot(pig_metadata, aes(x = factor(Body_Site))) + geom_bar(stat="count", width=0.7)
table(pig_metadata$Body_Site)

#142 separate studies for pig
length(table(pig_metadata$SRAStudy))
```


```{r}

sheep_samples = read.table('/home/werner/projects/cross_species_XCI/code/female_metadata/ovis_aries_female_SRR.txt')
sheep_samples = sheep_samples[ ,1]

sheep_metadata = genus_samples_df %>% filter(Run %in% sheep_samples)

ggplot(sheep_metadata, aes(x = factor(Tumor))) + geom_bar(stat="count", width=0.7)
ggplot(sheep_metadata, aes(x = factor(Disease))) + geom_bar(stat="count", width=0.7)
ggplot(sheep_metadata, aes(x = factor(Histological_Type))) + geom_bar(stat="count", width=0.7)
ggplot(sheep_metadata, aes(x = factor(Body_Site))) + geom_bar(stat="count", width=0.7)
table(sheep_metadata$Body_Site)

#76 separate studies for sheep
length(table(sheep_metadata$SRAStudy))
```





```{r}

goat_samples = read.table('/home/werner/projects/cross_species_XCI/code/female_metadata/Capra_hircus_female_SRR.txt')
goat_samples = goat_samples[ ,1]

goat_metadata = genus_samples_df %>% filter(Run %in% goat_samples)

ggplot(goat_metadata, aes(x = factor(Tumor))) + geom_bar(stat="count", width=0.7)
ggplot(goat_metadata, aes(x = factor(Disease))) + geom_bar(stat="count", width=0.7)
ggplot(goat_metadata, aes(x = factor(Histological_Type))) + geom_bar(stat="count", width=0.7)
ggplot(goat_metadata, aes(x = factor(Body_Site))) + geom_bar(stat="count", width=0.7)
table(goat_metadata$Body_Site)

#67 separate studies for goat
length(table(goat_metadata$SRAStudy))
```



```{r}

dog_samples = read.table('/home/werner/projects/cross_species_XCI/code/female_metadata/Canis_lupus_familiaris_female_SRR.txt')
dog_samples = dog_samples[ ,1]

dog_metadata = genus_samples_df %>% filter(Run %in% dog_samples)

ggplot(dog_metadata, aes(x = factor(Tumor))) + geom_bar(stat="count", width=0.7)
ggplot(dog_metadata, aes(x = factor(Disease))) + geom_bar(stat="count", width=0.7)
ggplot(dog_metadata, aes(x = factor(Histological_Type))) + geom_bar(stat="count", width=0.7)
ggplot(dog_metadata, aes(x = factor(Body_Site))) + geom_bar(stat="count", width=0.7)
table(dog_metadata$Body_Site)

#39 separate studies for dog
length(table(dog_metadata$SRAStudy))
```




```{r}

horse_samples = read.table('/home/werner/projects/cross_species_XCI/code/female_metadata/Equus_caballus_female_SRR.txt')
horse_samples = horse_samples[ ,1]

horse_metadata = genus_samples_df %>% filter(Run %in% horse_samples)

ggplot(horse_metadata, aes(x = factor(Tumor))) + geom_bar(stat="count", width=0.7)
ggplot(horse_metadata, aes(x = factor(Disease))) + geom_bar(stat="count", width=0.7)
ggplot(horse_metadata, aes(x = factor(Histological_Type))) + geom_bar(stat="count", width=0.7)
ggplot(horse_metadata, aes(x = factor(Body_Site))) + geom_bar(stat="count", width=0.7)
table(horse_metadata$Body_Site)

#34 separate studies for horse
length(table(horse_metadata$SRAStudy))
```




```{r}

macaca_samples = read.table('/home/werner/projects/cross_species_XCI/code/female_metadata/macaca_mulatta_female_SRR.txt')
macaca_samples = macaca_samples[ ,1]

macaca_metadata = genus_samples_df %>% filter(Run %in% macaca_samples)

ggplot(macaca_metadata, aes(x = factor(Tumor))) + geom_bar(stat="count", width=0.7)
ggplot(macaca_metadata, aes(x = factor(Disease))) + geom_bar(stat="count", width=0.7)
ggplot(macaca_metadata, aes(x = factor(Histological_Type))) + geom_bar(stat="count", width=0.7)
ggplot(macaca_metadata, aes(x = factor(Body_Site))) + geom_bar(stat="count", width=0.7)
table(macaca_metadata$Body_Site)

#46 separate studies for macaca
length(table(macaca_metadata$SRAStudy))
```



```{r}
length(unique(c(names(table(cow_metadata$SRAStudy)), names(table(rat_metadata$SRAStudy)), names(table(pig_metadata$SRAStudy)),
                names(table(goat_metadata$SRAStudy)),names(table(sheep_metadata$SRAStudy)),names(table(dog_metadata$SRAStudy)),
                names(table(horse_metadata$SRAStudy)),names(table(macaca_metadata$SRAStudy)))))



```






