

Analysis to quantify the relationship between XCI ratios and individual variants
Quantify with an AUROC, ranking on the XCI ratios and calling the samples with any given individual variant as the positives

```{r}
library(ggplot2)
library(dplyr)
library(MetBrewer)
library(ggrepel)
library(parallel)
library(stringr)
```


get colors for the species

```{r}

species_colors = met.brewer("Tiepolo", n=10,type="continuous")

#Going by the ranking of estimated cell counts
species_palette = c('Macaca' = species_colors[1], 'Rat' = species_colors[2], 'Pig' = species_colors[3],
                    'Goat' = species_colors[4], 'Horse' = species_colors[5], 'Sheep' = species_colors[6],
                    'Human' = species_colors[8], 'Mouse' = species_colors[7], 'Cow' = species_colors[9], 'Dog' = species_colors[10])

species_palette

```


skew and stat dataframes with escape bins filtered and the autosome filter
```{r}

load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/equus_caballus/horse_skew_and_stats_autoFilt_df_7_12_23.Rdata')
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/bos_taurus/cow_skew_and_stats_autoFilt_df_7_12_23.Rdata')
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/pig_skew_and_stats_autoFilt_df_7_12_23.Rdata')
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/ovis_aries/sheep_skew_and_stats_autoFilt_df_7_12_23.Rdata')
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/capra_hircus/goat_skew_and_stats_autoFilt_df_7_12_23.Rdata')
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/macaca_mulatta/macaca_skew_and_stats_autoFilt_df_7_12_23.Rdata')
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/rattus_norvegicus/rat_skew_and_stats_autoFilt_df_7_12_23.Rdata')
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/canis_lupus_familiaris/dog_skew_and_stats_autoFilt_df_7_12_23.Rdata')
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/mus_musculus/mouse_skew_and_stats_autoFilt_df_4_22_24.Rdata')
```


human data
```{r}

load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/homo_sapiens/human_skew_and_stats_autoFilt_df_7_12_23.Rdata')
dim(human_skew_and_stats_df)
```


snp stat dataframes
```{r}
x = load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/canis_lupus_familiaris/detected_het_snps_df_3_15_23.Rdata')
dog_snp_stat_df = get(x)

x = load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/equus_caballus/detected_het_snps_df_3_15_23.Rdata')
horse_snp_stat_df = get(x)

x = load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/bos_taurus/detected_het_snps_df_3_15_23.Rdata')
cow_snp_stat_df = get(x)

x = load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/ovis_aries/detected_het_snps_df_3_15_23.Rdata')
sheep_snp_stat_df = get(x)

x = load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/capra_hircus/detected_het_snps_df_3_15_23.Rdata')
goat_snp_stat_df = get(x)

x = load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/rattus_norvegicus/detected_het_snps_df_3_15_23.Rdata')
rat_snp_stat_df = get(x)

x = load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/macaca_mulatta/detected_het_snps_df_3_15_23.Rdata')
macaca_snp_stat_df = get(x)

x = load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/detected_het_snps_df_3_15_23.Rdata')
pig_snp_stat_df = get(x)

x = load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/mus_musculus/detected_het_snps_df_4_22_24.Rdata')
mouse_snp_stat_df = get(x)

x = load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/homo_sapiens/detected_het_snps_df_8_3_23.Rdata')
human_snp_stat_df = get(x)

```


vcfs with the annotated snp ratios

```{r}
#Dog variants
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/canis_lupus_familiaris/all_snp_vcf_3_15_23.Rdata')
#Horse variants
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/equus_caballus/all_snp_vcf_3_15_23.Rdata')
#Cow variants
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/bos_taurus/all_snp_vcf_3_15_23.Rdata')
#Sheep variants
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/ovis_aries/all_snp_vcf_3_15_23.Rdata')
#Goat variants
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/capra_hircus/all_snp_vcf_3_15_23.Rdata')
#Rat variants
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/rattus_norvegicus/all_snp_vcf_3_15_23.Rdata')
#Macaca variants
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/macaca_mulatta/all_snp_vcf_3_15_23.Rdata')
#Pig variants
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/sus_scrofa/all_snp_vcf_3_15_23.Rdata')

#Mouse variants
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/mus_musculus/all_snp_vcf_4_22_24.Rdata')

#Human
#All the detected snps from the GTEx data
#in the list.vcf object
load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/homo_sapiens/all_v8_GATK_mod.Rdata')

```



```{r}

length(annot_cow_vcfs)
cow_skew_and_stats_df
cow_snp_stat_df
```



```{r}
#Dog genome
dog_gtf = rtracklayer::import('/home/werner/projects/cross_species_XCI/genomes/Canis_lupus_familiaris/annotation.gtf')
dog_gtf = as.data.frame(dog_gtf)
#Genome specific x chromosome name
dog_x_chrom_name = readLines('/home/werner/projects/cross_species_XCI/genomes/Canis_lupus_familiaris/chrXlabel.txt', n = 1)
dog_x_gtf = dog_gtf[dog_gtf$seqnames == dog_x_chrom_name, ]
dog_x_genenames = unique(dog_x_gtf$gene)

dog_chromX_end = max(dog_x_gtf$end)

dog_x_meta = dog_x_gtf[dog_x_gtf$gbkey == 'Gene',]
dog_x_meta = dog_x_meta[!is.na(dog_x_meta$gbkey), ] #A lot of NA entries in the dog GTF file
dog_x_meta = dog_x_meta[!duplicated(dog_x_meta$gene), ]



#Horse genome
horse_gtf = rtracklayer::import('/home/werner/projects/cross_species_XCI/genomes/Equus_caballus/annotation.gtf')
horse_gtf = as.data.frame(horse_gtf)

#Genome specific x chromosome name
horse_x_chrom_name = readLines('/home/werner/projects/cross_species_XCI/genomes/Equus_caballus/chrXlabel.txt', n = 1)
horse_x_gtf = horse_gtf[horse_gtf$seqnames == horse_x_chrom_name, ]
horse_x_genenames = unique(horse_x_gtf$gene)
horse_chromX_end = max(horse_x_gtf$end)

horse_x_meta = horse_x_gtf[horse_x_gtf$gbkey == 'Gene',]



#Cow genome
cow_gtf = rtracklayer::import('/home/werner/projects/cross_species_XCI/genomes/Bos_taurus/annotation.gtf')
cow_gtf = as.data.frame(cow_gtf)

#Genome specific x chromosome name
cow_x_chrom_name = readLines('/home/werner/projects/cross_species_XCI/genomes/Bos_taurus/chrXlabel.txt', n = 1)
cow_x_gtf = cow_gtf[cow_gtf$seqnames == cow_x_chrom_name, ]
cow_x_genenames = unique(cow_x_gtf$gene)
cow_chromX_end = max(cow_x_gtf$end)

cow_x_meta = cow_x_gtf[cow_x_gtf$gbkey == 'Gene',]



#Sheep genome
sheep_gtf = rtracklayer::import('/home/werner/projects/cross_species_XCI/genomes/Ovis_aries/annotation.gtf')
sheep_gtf = as.data.frame(sheep_gtf)

#Genome specific x chromosome name
sheep_x_chrom_name = readLines('/home/werner/projects/cross_species_XCI/genomes/Ovis_aries/chrXlabel.txt', n = 1)
sheep_x_gtf = sheep_gtf[sheep_gtf$seqnames == sheep_x_chrom_name, ]
sheep_x_genenames = unique(sheep_x_gtf$gene)
sheep_chromX_end = max(sheep_x_gtf$end)

sheep_x_meta = sheep_x_gtf[sheep_x_gtf$gbkey == 'Gene',]
sheep_x_meta = sheep_x_meta[!is.na(sheep_x_meta$gbkey), ] #A lot of NA entries in the sheep GTF file
sheep_x_meta = sheep_x_meta[!duplicated(sheep_x_meta$gene), ]



#Goat genome
goat_gtf = rtracklayer::import('/home/werner/projects/cross_species_XCI/genomes/Capra_hircus/annotation.gtf')
goat_gtf = as.data.frame(goat_gtf)

#Genome specific x chromosome name
goat_x_chrom_name = readLines('/home/werner/projects/cross_species_XCI/genomes/Capra_hircus/chrXlabel.txt', n = 1)
goat_x_gtf = goat_gtf[goat_gtf$seqnames == goat_x_chrom_name, ]
goat_x_genenames = unique(goat_x_gtf$gene)
goat_chromX_end = max(goat_x_gtf$end)

goat_x_meta = goat_x_gtf[goat_x_gtf$gbkey == 'Gene',]



#Pig genome
pig_gtf = rtracklayer::import('/home/werner/projects/cross_species_XCI/genomes/Sus_scrofa/annotation.gtf')
pig_gtf = as.data.frame(pig_gtf)

#Genome specific x chromosome name
pig_x_chrom_name = readLines('/home/werner/projects/cross_species_XCI/genomes/Sus_scrofa/chrXlabel.txt', n = 1)
pig_x_gtf = pig_gtf[pig_gtf$seqnames == pig_x_chrom_name, ]
pig_x_genenames = unique(pig_x_gtf$gene)
pig_chromX_end = max(pig_x_gtf$end)

pig_x_meta = pig_x_gtf[pig_x_gtf$gbkey == 'Gene',]



#Macaca genome
macaca_gtf = rtracklayer::import('/home/werner/projects/cross_species_XCI/genomes/Macaca_mulatta/annotation.gtf')
macaca_gtf = as.data.frame(macaca_gtf)

#Genome specific x chromosome name
macaca_x_chrom_name = readLines('/home/werner/projects/cross_species_XCI/genomes/Macaca_mulatta/chrXlabel.txt', n = 1)
macaca_x_gtf = macaca_gtf[macaca_gtf$seqnames == macaca_x_chrom_name, ]
macaca_x_genenames = unique(macaca_x_gtf$gene)
macaca_chromX_end = max(macaca_x_gtf$end)

macaca_x_meta = macaca_x_gtf[macaca_x_gtf$gbkey == 'Gene',]



#Rat genome
rat_gtf = rtracklayer::import('/home/werner/projects/cross_species_XCI/genomes/Rattus_norvegicus/annotation.gtf')
rat_gtf = as.data.frame(rat_gtf)

#Genome specific x chromosome name
rat_x_chrom_name = readLines('/home/werner/projects/cross_species_XCI/genomes/Rattus_norvegicus/chrXlabel.txt', n = 1)
rat_x_gtf = rat_gtf[rat_gtf$seqnames == rat_x_chrom_name, ]
rat_x_genenames = unique(rat_x_gtf$gene)
rat_chromX_end = max(rat_x_gtf$end)

rat_x_meta = rat_x_gtf[rat_x_gtf$gbkey == 'Gene',]
rat_x_meta = rat_x_meta[!is.na(rat_x_meta$gbkey), ] #A lot of NA entries in the rat GTF file
rat_x_meta = rat_x_meta[!duplicated(rat_x_meta$gene), ]


#Mouse genome
mouse_gtf = rtracklayer::import('/home/werner/projects/cross_species_XCI/genomes/Mus_musculus/GCF_000001635.27_GRCm39_genomic.gtf')
mouse_gtf = as.data.frame(mouse_gtf)

#Genome specific x chromosome name
mouse_x_chrom_name = readLines('/home/werner/projects/cross_species_XCI/genomes/Mus_musculus/chrXlabel.txt', n = 1)
mouse_x_gtf = mouse_gtf[mouse_gtf$seqnames == mouse_x_chrom_name, ]
mouse_x_genenames = unique(mouse_x_gtf$gene)
mouse_chromX_end = max(mouse_x_gtf$end)

mouse_x_meta = mouse_x_gtf[mouse_x_gtf$gbkey == 'Gene',]

#Human genome
human_gtf = rtracklayer::import('/home/werner/projects/cross_species_XCI/genomes/Homo_sapiens/gencode.v25.annotation.gtf')
human_gtf = as.data.frame(human_gtf)
human_x_gtf = human_gtf[human_gtf$seqnames == 'chrX', ]

human_x_genenames = unique(human_x_gtf$gene_name)
human_chromX_end = max(human_x_gtf$end)

human_x_meta = human_x_gtf[human_x_gtf$type== 'gene',]



```


Add gene annotations to the species snp dataframes for ease of use later

```{r}
#horse
x_gene_starts = horse_x_meta$start
x_gene_ends = horse_x_meta$end
#Add gene_annotations
positions = horse_snp_stat_df$snp_pos
horse_gene_annots = vector(mode = 'character', length = length(positions))
for( i in 1:length(positions)){

  test_gene_start = positions[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    horse_gene_annots[i] = horse_x_meta$gene[gene_index]
  }
}

horse_snp_stat_df$gene = horse_gene_annots

#cow
x_gene_starts = cow_x_meta$start
x_gene_ends = cow_x_meta$end
#Add gene_annotations
positions = cow_snp_stat_df$snp_pos
cow_gene_annots = vector(mode = 'character', length = length(positions))
for( i in 1:length(positions)){

  test_gene_start = positions[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    cow_gene_annots[i] = cow_x_meta$gene[gene_index]
  }
}

cow_snp_stat_df$gene = cow_gene_annots


#pig
x_gene_starts = pig_x_meta$start
x_gene_ends = pig_x_meta$end
#Add gene_annotations
positions = pig_snp_stat_df$snp_pos
pig_gene_annots = vector(mode = 'character', length = length(positions))
for( i in 1:length(positions)){

  test_gene_start = positions[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    pig_gene_annots[i] = pig_x_meta$gene[gene_index]
  }
}

pig_snp_stat_df$gene = pig_gene_annots

#sheep
x_gene_starts = sheep_x_meta$start
x_gene_ends = sheep_x_meta$end
#Add gene_annotations
positions = sheep_snp_stat_df$snp_pos
sheep_gene_annots = vector(mode = 'character', length = length(positions))
for( i in 1:length(positions)){
  
  test_gene_start = positions[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    sheep_gene_annots[i] = sheep_x_meta$gene[gene_index]
  }
}

sheep_snp_stat_df$gene = sheep_gene_annots

#goat
x_gene_starts = goat_x_meta$start
x_gene_ends = goat_x_meta$end
#Add gene_annotations
positions = goat_snp_stat_df$snp_pos
goat_gene_annots = vector(mode = 'character', length = length(positions))
for( i in 1:length(positions)){

  test_gene_start = positions[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    goat_gene_annots[i] = goat_x_meta$gene[gene_index]
  }
}

goat_snp_stat_df$gene = goat_gene_annots

#macaca
x_gene_starts = macaca_x_meta$start
x_gene_ends = macaca_x_meta$end
#Add gene_annotations
positions = macaca_snp_stat_df$snp_pos
macaca_gene_annots = vector(mode = 'character', length = length(positions))
for( i in 1:length(positions)){

  test_gene_start = positions[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    macaca_gene_annots[i] = macaca_x_meta$gene[gene_index]
  }
}

macaca_snp_stat_df$gene = macaca_gene_annots

#dog
x_gene_starts = dog_x_meta$start
x_gene_ends = dog_x_meta$end
#Add gene_annotations
positions = dog_snp_stat_df$snp_pos
dog_gene_annots = vector(mode = 'character', length = length(positions))
for( i in 1:length(positions)){

  test_gene_start = positions[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    dog_gene_annots[i] = dog_x_meta$gene[gene_index]
  }
}

dog_snp_stat_df$gene = dog_gene_annots

#rat
x_gene_starts = rat_x_meta$start
x_gene_ends = rat_x_meta$end
#Add gene_annotations
positions = rat_snp_stat_df$snp_pos
rat_gene_annots = vector(mode = 'character', length = length(positions))
for( i in 1:length(positions)){

  test_gene_start = positions[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    rat_gene_annots[i] = rat_x_meta$gene[gene_index]
  }
}

rat_snp_stat_df$gene = rat_gene_annots


#mouse
x_gene_starts = mouse_x_meta$start
x_gene_ends = mouse_x_meta$end
#Add gene_annotations
positions = mouse_snp_stat_df$snp_pos
mouse_gene_annots = vector(mode = 'character', length = length(positions))
for( i in 1:length(positions)){

  test_gene_start = positions[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    mouse_gene_annots[i] = mouse_x_meta$gene[gene_index]
  }
}

mouse_snp_stat_df$gene = mouse_gene_annots


#human
x_gene_starts = human_x_meta$start
x_gene_ends = human_x_meta$end
#Add gene_annotations
positions =  as.numeric(stringr::str_extract(human_snp_stat_df$variant, "[0-9]+"))
human_snp_stat_df$snp_pos = positions
human_gene_annots = vector(mode = 'character', length = length(positions))
for( i in 1:length(positions)){

  test_gene_start = positions[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    human_gene_annots[i] = human_x_meta$gene_name[gene_index]
  }
}


human_snp_stat_df$gene = human_gene_annots

```





###################################
Get the number of SNPs around XIST for species with XIST annotations, see if heterzygosity tracks with estimated skew
##################################



```{r}


dog_x_genenames[grepl('X',dog_x_genenames)] 

cow_x_genenames[grepl('X',cow_x_genenames)] 

pig_x_genenames[grepl('X',pig_x_genenames)] 

sheep_x_genenames[grepl('X',sheep_x_genenames)] 

goat_x_genenames[grepl('X',goat_x_genenames)] 

rat_x_genenames[grepl('x',rat_x_genenames)] 

macaca_x_genenames[grepl('X',macaca_x_genenames)] 

horse_x_genenames[grepl('X',horse_x_genenames)] 

human_x_genenames[grepl('X',human_x_genenames)] 
```


```{r}
#Rat XIST is LOC100911498
#XIC is roughly 68426668 to 68630338
#So approximately 68400000 to 68650000
rat_x_meta[!is.na(rat_x_meta$start), ]


rat_xic_start = 68400000
rat_xic_end = 68650000
rat_xic_hetSNPs = vector(mode = 'numeric', length = length(annot_rat_vcfs))
for(i in 1:length(annot_rat_vcfs)){
  
  if(is.null(dim(annot_rat_vcfs[[i]]))){rat_xic_hetSNPs[i] = NA; next}
  data = annot_rat_vcfs[[i]]
  rat_xic_hetSNPs[i] = sum(data$Pos >= rat_xic_start & data$Pos <= rat_xic_end)
}

hist(rat_xic_hetSNPs)

plot(rat_skew_and_stats_df$est_skew_no_escape, rat_xic_hetSNPs)
plot(rat_skew_and_stats_df$est_skew_total_snps, rat_xic_hetSNPs)
```


```{r}


compute_auc = function(scores, label){
  w_test = wilcox.test(scores[label],scores[!label], alternative = 'two.sided', exact = F)
  p_value = w_test$p.value
  auc = as.numeric(w_test$statistic / (sum(label) * sum(!label)))
  return(list('p_value' = p_value, 'auc' = auc))
}


bootstrap_auc = function(scores, label){
  #Randomly sample with replacement, same number as the original sample
  positive_sample = sample(scores[label], size = sum(label), replace = T)  
  negative_sample = sample(scores[!label], size = sum(!label), replace = T)  
  w_test = wilcox.test(positive_sample, negative_sample, alternative = 'two.sided', exact = F)
  p_value = w_test$p.value
  auc = as.numeric(w_test$statistic / (sum(label) * sum(!label)))
  #auc = compute_auc(scores = c(positive_sample, negative_sample), label = c(rep(TRUE, length = length(positive_sample)), rep(FALSE, length = length(negative_sample))) )
  return(list('p_value' = p_value, 'auc' = auc))
}


compute_auc_power = function(scores, label, num_sims = 2000, sig_level = .05){
  bootstrap_results = mclapply(1:num_sims, function(i) bootstrap_auc(scores, label), mc.cores = 10)
  bootstrap_pvals = sapply(bootstrap_results, '[[', 1)
  bootstrap_aucs = sapply(bootstrap_results, '[[', 2)
  
  power_sig = sum(bootstrap_pvals <= sig_level) / num_sims
  power_effect_55 = sum(bootstrap_aucs >= .55) / num_sims
  power_effect_75 = sum(bootstrap_aucs >= .75) / num_sims
  auc_var = var(bootstrap_aucs)
  return(list('power_sig' = power_sig, 'power_effect_55' = power_effect_55, 'power_effect_75' = power_effect_75, 'bootstrap_auc_var' = auc_var))

}

# Functions to fold and unfold ratios 
folded <- function(x) { apply( cbind(x,1-x), 1, max)} 
unfold <- function(x) { c(x,1-x) } 

```


```{r}
#sample index is a vector of indices for samples with at least 10 good SNPs
xci_skew_predict_variant_experiment = function(skew_and_stats_df, annot_vcfs, sample_index, snp_stats){
  #Go through all the good samples and get the list of all the variants present
  test_variants = c()
  for(i in 1:length(sample_index)){
  data_variants = paste0(annot_vcfs[[sample_index[i]]]$Pos, paste0(annot_vcfs[[sample_index[i]]]$ref, annot_vcfs[[sample_index[i]]]$alt))
  test_variants = c(test_variants, data_variants)
  }
  
  #Get the sample counts of each test variant
  test_variants = table(test_variants)
  total_samples = length(sample_index)
  #Number of positives and negatives for each variant and the prevalence of each variant
  n_pos = as.numeric(test_variants[test_variants >= 10])
  n_neg = total_samples - n_pos
  variant_prevalence = n_pos / total_samples
  test_variants = names(test_variants[test_variants >= 10])
  
  #Just using the samples with XCI ratios from 10 genes or more
  filt_annot_vcfs = annot_vcfs[sample_index]
  auc_test_variants = vector(mode = 'numeric', length = length(test_variants))
  auc_pval = vector(mode = 'numeric', length = length(test_variants))
  auc_power_sig = vector(mode = 'numeric', length = length(test_variants))
  auc_power_effect_55 = vector(mode = 'numeric', length = length(test_variants))
  auc_power_effect_75 = vector(mode = 'numeric', length = length(test_variants))
  bootstrap_auc_var = vector(mode = 'numeric', length = length(test_variants))
  auc_null = vector(mode = 'numeric', length = length(test_variants))
  
  for(j in 1:length(test_variants)){
    if(j%%100 == 0){print(j)}
    #Go through all samples and get a boolean vector for which samples have that variant
    variant_present = vector(mode = 'logical', length = length(filt_annot_vcfs))
    for(i in 1:length(filt_annot_vcfs)){
      data_variants = paste0(filt_annot_vcfs[[i]]$Pos, paste0(filt_annot_vcfs[[i]]$ref, filt_annot_vcfs[[i]]$alt))
      variant_present[i] = test_variants[j] %in% data_variants
    }
    #AUC using the estimated skew without escape
    scores = skew_and_stats_df$est_skew_no_escape[sample_index]
    #Get the auroc and it's p_value
    auc_results = compute_auc(scores, variant_present)
    auc_test_variants[j] = auc_results$auc
    auc_pval[j] = auc_results$p_value
    #Compute the power calculation
    power_results = compute_auc_power(scores, variant_present)
    auc_power_sig[j] = power_results$power_sig
    auc_power_effect_55[j] = power_results$power_effect_55
    auc_power_effect_75[j] = power_results$power_effect_75
    bootstrap_auc_var[j] = power_results$bootstrap_auc_var
    #Get a permuted auroc
    auc_null[j] = compute_auc(scores, sample(variant_present))$auc #AUC with the labels randomly shuffled for a null
  }
  
  #Variant positions
  positions =  as.numeric(stringr::str_extract(test_variants, "[0-9]+"))
  #reference biased positions
  ref_biased_pos = snp_stats$snp_pos[snp_stats$filter == FALSE]
  #Which of the variants are reference biased
  ref_index = positions %in% ref_biased_pos
  
  results = list('variants' = test_variants, 'reference_biased_index' = ref_index, 'n_pos' = n_pos, 'n_neg' = n_neg, 'variant_prevalence' = variant_prevalence, 
                 'auc' = auc_test_variants, 'auc_pval' = auc_pval, 'auc_power_significance' = auc_power_sig, 'auc_power_effect_55' = auc_power_effect_55, 
                 'auc_power_effect_75' = auc_power_effect_75, 'bootstrap_auc_var' = bootstrap_auc_var, 'auc_null' = auc_null)
  return(results)
}

```







```{r}

start = Sys.time()
sample_index = which(rat_skew_and_stats_df$num_good_snps_no_escape >= 10 & rat_skew_and_stats_df$autosome_imbalance == F)
rat_results = xci_skew_predict_variant_experiment(rat_skew_and_stats_df, annot_rat_vcfs, sample_index, rat_snp_stat_df)
Sys.time() - start

hist(rat_results$auc_null, breaks = seq(0,1,.01), main = 'rat skew predicting SNP aurocs: null')
abline(v = .5, col = 'red')
hist(rat_results$auc, breaks = seq(0,1,.01), main = 'rat skew predicting SNP aurocs')
abline(v = .5, col = 'red')
plot(rat_results$auc, rat_results$auc_power_significance, main = 'rat AUROC against significance power', xlab = 'AUROC', ylab = 'Simulated significance power')
plot(rat_results$auc, rat_results$auc_power_effect_55, main = 'rat effect size power thresh: 0.55', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(rat_results$auc, rat_results$auc_power_effect_75, main = 'rat effect size power thresh: 0.75', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(rat_results$variant_prevalence, rat_results$auc_power_effect_55, main = 'rat variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(rat_results$variant_prevalence, rat_results$auc_power_effect_75, main = 'rat variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(rat_results$variant_prevalence, rat_results$bootstrap_auc_var, main = 'rat variant % against boostrap auc variance', xlab = 'Variant prevalence', ylab = 'Variance of bootstrap aurocs')
plot(rat_results$auc, rat_results$bootstrap_auc_var, xlab = 'AUROC', ylab = 'Variance of bootstrap AUROC distribution', main = 'rat samples')


hist(rat_results$auc_pval, main = 'Raw auroc p-values', breaks = seq(0,1,.05))
adj_pval = p.adjust(rat_results$auc_pval, method = 'BH')
hist(adj_pval, main = 'FDR adjusted auroc p-values', breaks = seq(0,1,.05))
fdr_index = which(adj_pval <= .05)
hist(rat_results$auc[fdr_index], main = 'Significant aurocs', breaks = seq(0,1,.01))


rat_ref_check_df = data.frame(auc = rat_results$auc, ref_biased_index = rat_results$reference_biased_index)
ggplot(rat_ref_check_df, aes(x = ref_biased_index, y = auc, fill = ref_biased_index)) + geom_violin(scale = 'width') + geom_boxplot(width = .25) + ggtitle('Rat') +
  xlab('Reference biased SNPs') + ylab('XCI ratio and variant AUROC') 
#ggsave(filename = 'rat_ref_biased_xci_ratio_auroc_violins_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/',
#       device = 'pdf', useDingbats = F, height = 4, width = 5)

```



```{r}

start = Sys.time()
sample_index = which(mouse_skew_and_stats_df$num_good_snps_no_escape >= 10 & mouse_skew_and_stats_df$autosome_imbalance == F)
mouse_results = xci_skew_predict_variant_experiment(mouse_skew_and_stats_df, annot_mouse_vcfs, sample_index, mouse_snp_stat_df)
Sys.time() - start

hist(mouse_results$auc_null, breaks = seq(0,1,.01), main = 'mouse skew predicting SNP aurocs: null')
abline(v = .5, col = 'red')
hist(mouse_results$auc, breaks = seq(0,1,.01), main = 'mouse skew predicting SNP aurocs')
abline(v = .5, col = 'red')
plot(mouse_results$auc, mouse_results$auc_power_significance, main = 'mouse AUROC against significance power', xlab = 'AUROC', ylab = 'Simulated significance power')
plot(mouse_results$auc, mouse_results$auc_power_effect_55, main = 'mouse effect size power thresh: 0.55', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(mouse_results$auc, mouse_results$auc_power_effect_75, main = 'mouse effect size power thresh: 0.75', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(mouse_results$variant_prevalence, mouse_results$auc_power_effect_55, main = 'mouse variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(mouse_results$variant_prevalence, mouse_results$auc_power_effect_75, main = 'mouse variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(mouse_results$variant_prevalence, mouse_results$bootstrap_auc_var, main = 'mouse variant % against boostrap auc variance', xlab = 'Variant prevalence', ylab = 'Variance of bootstrap aurocs')
plot(mouse_results$auc, mouse_results$bootstrap_auc_var, xlab = 'AUROC', ylab = 'Variance of bootstrap AUROC distribution', main = 'mouse samples')


hist(mouse_results$auc_pval, main = 'Raw auroc p-values', breaks = seq(0,1,.05))
adj_pval = p.adjust(mouse_results$auc_pval, method = 'BH')
hist(adj_pval, main = 'FDR adjusted auroc p-values', breaks = seq(0,1,.05))
fdr_index = which(adj_pval <= .05)
hist(mouse_results$auc[fdr_index], main = 'Significant aurocs', breaks = seq(0,1,.01))


mouse_ref_check_df = data.frame(auc = mouse_results$auc, ref_biased_index = mouse_results$reference_biased_index)
ggplot(mouse_ref_check_df, aes(x = ref_biased_index, y = auc, fill = ref_biased_index)) + geom_violin(scale = 'width') + geom_boxplot(width = .25) + ggtitle('Mouse') +
  xlab('Reference biased SNPs') + ylab('XCI ratio and variant AUROC') 
#ggsave(filename = 'mouse_ref_biased_xci_ratio_auroc_violins_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/',
#       device = 'pdf', useDingbats = F, height = 4, width = 5)

```






```{r}
#macaca
start = Sys.time()
sample_index = which(macaca_skew_and_stats_df$num_good_snps_no_escape >= 10 & macaca_skew_and_stats_df$autosome_imbalance == F)
macaca_results = xci_skew_predict_variant_experiment(macaca_skew_and_stats_df, annot_macaca_vcfs, sample_index, macaca_snp_stat_df)
Sys.time() - start
#dog
start = Sys.time()
sample_index = which(dog_skew_and_stats_df$num_good_snps_no_escape >= 10 & dog_skew_and_stats_df$autosome_imbalance == F)
dog_results = xci_skew_predict_variant_experiment(dog_skew_and_stats_df, annot_dog_vcfs, sample_index, dog_snp_stat_df)
Sys.time() - start
#goat
start = Sys.time()
sample_index = which(goat_skew_and_stats_df$num_good_snps_no_escape >= 10 & goat_skew_and_stats_df$autosome_imbalance == F)
goat_results = xci_skew_predict_variant_experiment(goat_skew_and_stats_df, annot_goat_vcfs, sample_index, goat_snp_stat_df)
Sys.time() - start
#pig
start = Sys.time()
sample_index = which(pig_skew_and_stats_df$num_good_snps_no_escape >= 10 & pig_skew_and_stats_df$autosome_imbalance == F)
pig_results = xci_skew_predict_variant_experiment(pig_skew_and_stats_df, annot_pig_vcfs, sample_index, pig_snp_stat_df)
Sys.time() - start
#horse
start = Sys.time()
sample_index = which(horse_skew_and_stats_df$num_good_snps_no_escape >= 10 & horse_skew_and_stats_df$autosome_imbalance == F)
horse_results = xci_skew_predict_variant_experiment(horse_skew_and_stats_df, annot_horse_vcfs, sample_index, horse_snp_stat_df)
Sys.time() - start
#cow
start = Sys.time()
sample_index = which(cow_skew_and_stats_df$num_good_snps_no_escape >= 10 & cow_skew_and_stats_df$autosome_imbalance == F)
cow_results = xci_skew_predict_variant_experiment(cow_skew_and_stats_df, annot_cow_vcfs, sample_index, cow_snp_stat_df)
Sys.time() - start
#sheep
start = Sys.time()
sample_index = which(sheep_skew_and_stats_df$num_good_snps_no_escape >= 10 & sheep_skew_and_stats_df$autosome_imbalance == F)
sheep_results = xci_skew_predict_variant_experiment(sheep_skew_and_stats_df, annot_sheep_vcfs, sample_index, sheep_snp_stat_df)
Sys.time() - start

```


```{r}

hist(macaca_results$auc_null, breaks = seq(0,1,.01), main = 'macaca skew predicting SNP aurocs: null')
abline(v = .5, col = 'red')
hist(macaca_results$auc, breaks = seq(0,1,.01), main = 'macaca skew predicting SNP aurocs')
abline(v = .5, col = 'red')
plot(macaca_results$auc, macaca_results$auc_power_significance, main = 'macaca AUROC against significance power', xlab = 'AUROC', ylab = 'Simulated significance power')
plot(macaca_results$auc, macaca_results$auc_power_effect_55, main = 'macaca effect size power thresh: 0.55', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(macaca_results$auc, macaca_results$auc_power_effect_75, main = 'macaca effect size power thresh: 0.75', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(macaca_results$variant_prevalence, macaca_results$auc_power_effect_55, main = 'macaca variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(macaca_results$variant_prevalence, macaca_results$auc_power_effect_75, main = 'macaca variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(macaca_results$variant_prevalence, macaca_results$bootstrap_auc_var, main = 'macaca variant % against boostrap auc variance', xlab = 'Variant prevalence', ylab = 'Variance of bootstrap aurocs')
plot(macaca_results$auc, macaca_results$bootstrap_auc_var, xlab = 'AUROC', ylab = 'Variance of bootstrap AUROC distribution', main = 'macaca samples')


hist(macaca_results$auc_pval, main = 'Raw auroc p-values', breaks = seq(0,1,.05))
adj_pval = p.adjust(macaca_results$auc_pval, method = 'BH')
hist(adj_pval, main = 'FDR adjusted auroc p-values', breaks = seq(0,1,.05))
fdr_index = which(adj_pval <= .05)
hist(macaca_results$auc[fdr_index], main = 'Significant aurocs', breaks = seq(0,1,.01))


macaca_ref_check_df = data.frame(auc = macaca_results$auc, ref_biased_index = macaca_results$reference_biased_index)
ggplot(macaca_ref_check_df, aes(x = ref_biased_index, y = auc, fill = ref_biased_index)) + geom_violin(scale = 'width') + geom_boxplot(width = .25) + ggtitle('macaca') +
  xlab('Reference biased SNPs') + ylab('XCI ratio and variant AUROC') 
#ggsave(filename = 'macaca_ref_biased_xci_ratio_auroc_violins_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/',
#       device = 'pdf', useDingbats = F, height = 4, width = 5)
```


```{r}

hist(dog_results$auc_null, breaks = seq(0,1,.01), main = 'dog skew predicting SNP aurocs: null')
abline(v = .5, col = 'red')
hist(dog_results$auc, breaks = seq(0,1,.01), main = 'dog skew predicting SNP aurocs')
abline(v = .5, col = 'red')
plot(dog_results$auc, dog_results$auc_power_significance, main = 'dog AUROC against significance power', xlab = 'AUROC', ylab = 'Simulated significance power')
plot(dog_results$auc, dog_results$auc_power_effect_55, main = 'dog effect size power thresh: 0.55', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(dog_results$auc, dog_results$auc_power_effect_75, main = 'dog effect size power thresh: 0.75', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(dog_results$variant_prevalence, dog_results$auc_power_effect_55, main = 'dog variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(dog_results$variant_prevalence, dog_results$auc_power_effect_75, main = 'dog variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(dog_results$variant_prevalence, dog_results$bootstrap_auc_var, main = 'dog variant % against boostrap auc variance', xlab = 'Variant prevalence', ylab = 'Variance of bootstrap aurocs')
plot(dog_results$auc, dog_results$bootstrap_auc_var, xlab = 'AUROC', ylab = 'Variance of bootstrap AUROC distribution', main = 'dog samples')


hist(dog_results$auc_pval, main = 'Raw auroc p-values', breaks = seq(0,1,.05))
adj_pval = p.adjust(dog_results$auc_pval, method = 'BH')
hist(adj_pval, main = 'FDR adjusted auroc p-values', breaks = seq(0,1,.05))
fdr_index = which(adj_pval <= .05)
hist(dog_results$auc[fdr_index], main = 'Significant aurocs', breaks = seq(0,1,.01))


dog_ref_check_df = data.frame(auc = dog_results$auc, ref_biased_index = dog_results$reference_biased_index)
ggplot(dog_ref_check_df, aes(x = ref_biased_index, y = auc, fill = ref_biased_index)) + geom_violin(scale = 'width') + geom_boxplot(width = .25) + ggtitle('dog') +
  xlab('Reference biased SNPs') + ylab('XCI ratio and variant AUROC') 
#ggsave(filename = 'dog_ref_biased_xci_ratio_auroc_violins_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/',
#       device = 'pdf', useDingbats = F, height = 4, width = 5)
```


```{r}

hist(goat_results$auc_null, breaks = seq(0,1,.01), main = 'goat skew predicting SNP aurocs: null')
abline(v = .5, col = 'red')
hist(goat_results$auc, breaks = seq(0,1,.01), main = 'goat skew predicting SNP aurocs')
abline(v = .5, col = 'red')
plot(goat_results$auc, goat_results$auc_power_significance, main = 'goat AUROC against significance power', xlab = 'AUROC', ylab = 'Simulated significance power')
plot(goat_results$auc, goat_results$auc_power_effect_55, main = 'goat effect size power thresh: 0.55', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(goat_results$auc, goat_results$auc_power_effect_75, main = 'goat effect size power thresh: 0.75', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(goat_results$variant_prevalence, goat_results$auc_power_effect_55, main = 'goat variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(goat_results$variant_prevalence, goat_results$auc_power_effect_75, main = 'goat variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(goat_results$variant_prevalence, goat_results$bootstrap_auc_var, main = 'goat variant % against boostrap auc variance', xlab = 'Variant prevalence', ylab = 'Variance of bootstrap aurocs')
plot(goat_results$auc, goat_results$bootstrap_auc_var, xlab = 'AUROC', ylab = 'Variance of bootstrap AUROC distribution', main = 'goat samples')


hist(goat_results$auc_pval, main = 'Raw auroc p-values', breaks = seq(0,1,.05))
adj_pval = p.adjust(goat_results$auc_pval, method = 'BH')
hist(adj_pval, main = 'FDR adjusted auroc p-values', breaks = seq(0,1,.05))
fdr_index = which(adj_pval <= .05)
hist(goat_results$auc[fdr_index], main = 'Significant aurocs', breaks = seq(0,1,.01))

goat_ref_check_df = data.frame(auc = goat_results$auc, ref_biased_index = goat_results$reference_biased_index)
ggplot(goat_ref_check_df, aes(x = ref_biased_index, y = auc, fill = ref_biased_index)) + geom_violin(scale = 'width') + geom_boxplot(width = .25) + ggtitle('goat') +
  xlab('Reference biased SNPs') + ylab('XCI ratio and variant AUROC') 
#ggsave(filename = 'goat_ref_biased_xci_ratio_auroc_violins_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/',
#       device = 'pdf', useDingbats = F, height = 4, width = 5)
```


```{r}

hist(pig_results$auc_null, breaks = seq(0,1,.01), main = 'pig skew predicting SNP aurocs: null')
abline(v = .5, col = 'red')
hist(pig_results$auc, breaks = seq(0,1,.01), main = 'pig skew predicting SNP aurocs')
abline(v = .5, col = 'red')
plot(pig_results$auc, pig_results$auc_power_significance, main = 'pig AUROC against significance power', xlab = 'AUROC', ylab = 'Simulated significance power')
plot(pig_results$auc, pig_results$auc_power_effect_55, main = 'pig effect size power thresh: 0.55', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(pig_results$auc, pig_results$auc_power_effect_75, main = 'pig effect size power thresh: 0.75', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(pig_results$variant_prevalence, pig_results$auc_power_effect_55, main = 'pig variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(pig_results$variant_prevalence, pig_results$auc_power_effect_75, main = 'pig variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(pig_results$variant_prevalence, pig_results$bootstrap_auc_var, main = 'pig variant % against boostrap auc variance', xlab = 'Variant prevalence', ylab = 'Variance of bootstrap aurocs')
plot(pig_results$auc, pig_results$bootstrap_auc_var, xlab = 'AUROC', ylab = 'Variance of bootstrap AUROC distribution', main = 'pig samples')


hist(pig_results$auc_pval, main = 'Raw auroc p-values', breaks = seq(0,1,.05))
adj_pval = p.adjust(pig_results$auc_pval, method = 'BH')
hist(adj_pval, main = 'FDR adjusted auroc p-values', breaks = seq(0,1,.05))
fdr_index = which(adj_pval <= .05)
hist(pig_results$auc[fdr_index], main = 'Significant aurocs', breaks = seq(0,1,.01))

pig_ref_check_df = data.frame(auc = pig_results$auc, ref_biased_index = pig_results$reference_biased_index)
ggplot(pig_ref_check_df, aes(x = ref_biased_index, y = auc, fill = ref_biased_index)) + geom_violin(scale = 'width') + geom_boxplot(width = .25) + ggtitle('pig') +
  xlab('Reference biased SNPs') + ylab('XCI ratio and variant AUROC') 
#ggsave(filename = 'pig_ref_biased_xci_ratio_auroc_violins_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/',
#       device = 'pdf', useDingbats = F, height = 4, width = 5)
```



```{r}

hist(horse_results$auc_null, breaks = seq(0,1,.01), main = 'horse skew predicting SNP aurocs: null')
abline(v = .5, col = 'red')
hist(horse_results$auc, breaks = seq(0,1,.01), main = 'horse skew predicting SNP aurocs')
abline(v = .5, col = 'red')
plot(horse_results$auc, horse_results$auc_power_significance, main = 'horse AUROC against significance power', xlab = 'AUROC', ylab = 'Simulated significance power')
plot(horse_results$auc, horse_results$auc_power_effect_55, main = 'horse effect size power thresh: 0.55', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(horse_results$auc, horse_results$auc_power_effect_75, main = 'horse effect size power thresh: 0.75', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(horse_results$variant_prevalence, horse_results$auc_power_effect_55, main = 'horse variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(horse_results$variant_prevalence, horse_results$auc_power_effect_75, main = 'horse variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(horse_results$variant_prevalence, horse_results$bootstrap_auc_var, main = 'horse variant % against boostrap auc variance', xlab = 'Variant prevalence', ylab = 'Variance of bootstrap aurocs')
plot(horse_results$auc, horse_results$bootstrap_auc_var, xlab = 'AUROC', ylab = 'Variance of bootstrap AUROC distribution', main = 'horse samples')


hist(horse_results$auc_pval, main = 'Raw auroc p-values', breaks = seq(0,1,.05))
adj_pval = p.adjust(horse_results$auc_pval, method = 'BH')
hist(adj_pval, main = 'FDR adjusted auroc p-values', breaks = seq(0,1,.05))
fdr_index = which(adj_pval <= .05)
hist(horse_results$auc[fdr_index], main = 'Significant aurocs', breaks = seq(0,1,.01))


horse_ref_check_df = data.frame(auc = horse_results$auc, ref_biased_index = horse_results$reference_biased_index)
ggplot(horse_ref_check_df, aes(x = ref_biased_index, y = auc, fill = ref_biased_index)) + geom_violin(scale = 'width') + geom_boxplot(width = .25) + ggtitle('horse') +
  xlab('Reference biased SNPs') + ylab('XCI ratio and variant AUROC') 
#ggsave(filename = 'horse_ref_biased_xci_ratio_auroc_violins_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/',
#       device = 'pdf', useDingbats = F, height = 4, width = 5)
```



```{r}

hist(cow_results$auc_null, breaks = seq(0,1,.01), main = 'cow skew predicting SNP aurocs: null')
abline(v = .5, col = 'red')
hist(cow_results$auc, breaks = seq(0,1,.01), main = 'cow skew predicting SNP aurocs')
abline(v = .5, col = 'red')
plot(cow_results$auc, cow_results$auc_power_significance, main = 'cow AUROC against significance power', xlab = 'AUROC', ylab = 'Simulated significance power')
plot(cow_results$auc, cow_results$auc_power_effect_55, main = 'cow effect size power thresh: 0.55', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(cow_results$auc, cow_results$auc_power_effect_75, main = 'cow effect size power thresh: 0.75', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(cow_results$variant_prevalence, cow_results$auc_power_effect_55, main = 'cow variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(cow_results$variant_prevalence, cow_results$auc_power_effect_75, main = 'cow variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(cow_results$variant_prevalence, cow_results$bootstrap_auc_var, main = 'cow variant % against boostrap auc variance', xlab = 'Variant prevalence', ylab = 'Variance of bootstrap aurocs')
plot(cow_results$auc, cow_results$bootstrap_auc_var, xlab = 'AUROC', ylab = 'Variance of bootstrap AUROC distribution', main = 'cow samples')


hist(cow_results$auc_pval, main = 'Raw auroc p-values', breaks = seq(0,1,.05))
adj_pval = p.adjust(cow_results$auc_pval, method = 'BH')
hist(adj_pval, main = 'FDR adjusted auroc p-values', breaks = seq(0,1,.05))
fdr_index = which(adj_pval <= .05)
hist(cow_results$auc[fdr_index], main = 'Significant aurocs', breaks = seq(0,1,.01))

cow_ref_check_df = data.frame(auc = cow_results$auc, ref_biased_index = cow_results$reference_biased_index)
ggplot(cow_ref_check_df, aes(x = ref_biased_index, y = auc, fill = ref_biased_index)) + geom_violin(scale = 'width') + geom_boxplot(width = .25) + ggtitle('cow') +
  xlab('Reference biased SNPs') + ylab('XCI ratio and variant AUROC') 
#ggsave(filename = 'cow_ref_biased_xci_ratio_auroc_violins_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/',
#       device = 'pdf', useDingbats = F, height = 4, width = 5)
```


```{r}

hist(sheep_results$auc_null, breaks = seq(0,1,.01), main = 'sheep skew predicting SNP aurocs: null')
abline(v = .5, col = 'red')
hist(sheep_results$auc, breaks = seq(0,1,.01), main = 'sheep skew predicting SNP aurocs')
abline(v = .5, col = 'red')
plot(sheep_results$auc, sheep_results$auc_power_significance, main = 'sheep AUROC against significance power', xlab = 'AUROC', ylab = 'Simulated significance power')
plot(sheep_results$auc, sheep_results$auc_power_effect_55, main = 'sheep effect size power thresh: 0.55', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(sheep_results$auc, sheep_results$auc_power_effect_75, main = 'sheep effect size power thresh: 0.75', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(sheep_results$variant_prevalence, sheep_results$auc_power_effect_55, main = 'sheep variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(sheep_results$variant_prevalence, sheep_results$auc_power_effect_75, main = 'sheep variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(sheep_results$variant_prevalence, sheep_results$bootstrap_auc_var, main = 'sheep variant % against boostrap auc variance', xlab = 'Variant prevalence', ylab = 'Variance of bootstrap aurocs')
plot(sheep_results$auc, sheep_results$bootstrap_auc_var, xlab = 'AUROC', ylab = 'Variance of bootstrap AUROC distribution', main = 'sheep samples')


hist(sheep_results$auc_pval, main = 'Raw auroc p-values', breaks = seq(0,1,.05))
adj_pval = p.adjust(sheep_results$auc_pval, method = 'BH')
hist(adj_pval, main = 'FDR adjusted auroc p-values', breaks = seq(0,1,.05))
fdr_index = which(adj_pval <= .05)
hist(sheep_results$auc[fdr_index], main = 'Significant aurocs', breaks = seq(0,1,.01))

sheep_ref_check_df = data.frame(auc = sheep_results$auc, ref_biased_index = sheep_results$reference_biased_index)
ggplot(sheep_ref_check_df, aes(x = ref_biased_index, y = auc, fill = ref_biased_index)) + geom_violin(scale = 'width') + geom_boxplot(width = .25) + ggtitle('sheep') +
  xlab('Reference biased SNPs') + ylab('XCI ratio and variant AUROC') 
#ggsave(filename = 'sheep_ref_biased_xci_ratio_auroc_violins_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/',
#       device = 'pdf', useDingbats = F, height = 4, width = 5)
```

#human samples

```{r}

#Go through the samples that have XCI ratio estimates with at least 10 genes and grab all the variants present across all samples
skew_index = human_skew_and_stats_df$sample_index[human_skew_and_stats_df$num_snps >= 10 & human_skew_and_stats_df$autosome_imbalance == F]
human_scores = human_skew_and_stats_df$skew[human_skew_and_stats_df$num_snps >= 10 & human_skew_and_stats_df$autosome_imbalance == F]
total_human_samps = length(skew_index) #Total number of rat samples
test_variants = c()
for(i in 1:length(skew_index)){
  data_variants = paste0(list.vcf[[skew_index[i]]][ ,1], paste0(list.vcf[[skew_index[i]]]$V3, list.vcf[[skew_index[i]]]$V4))
  test_variants = c(test_variants, data_variants)
}
#Get the sample counts of each test variant
test_variants = table(test_variants)
#Number of positives and negatives for each variant and the prevalence of each variant
n_pos = as.numeric(test_variants[test_variants >= 10])
n_neg = total_human_samps - n_pos
variant_prevalence = n_pos / total_human_samps
test_variants = names(test_variants[test_variants >= 10])
length(test_variants)

#Just using the samples with XCI ratios from 10 genes or more
skew_filt_annot_human_vcfs = list.vcf[skew_index]
auc_test_variants = vector(mode = 'numeric', length = length(test_variants))
auc_pval = vector(mode = 'numeric', length = length(test_variants))
auc_power_sig = vector(mode = 'numeric', length = length(test_variants))
auc_power_effect_55 = vector(mode = 'numeric', length = length(test_variants))
auc_power_effect_75 = vector(mode = 'numeric', length = length(test_variants))
bootstrap_auc_var = vector(mode = 'numeric', length = length(test_variants))
auc_null = vector(mode = 'numeric', length = length(test_variants))

for(j in 1:length(test_variants)){
  if(j%%100 == 0){print(j)}
  #Go through all samples and get a boolean vector for which samples have that variant
  variant_present = vector(mode = 'logical', length = length(skew_filt_annot_human_vcfs))
  for(i in 1:length(skew_filt_annot_human_vcfs)){
  
    data_variants = paste0(skew_filt_annot_human_vcfs[[i]][ ,1], paste0(skew_filt_annot_human_vcfs[[i]]$V3, skew_filt_annot_human_vcfs[[i]]$V4))
    variant_present[i] = test_variants[j] %in% data_variants
  }
  
  #Get the auroc and it's p_value
  auc_results = compute_auc(human_scores, variant_present)
  auc_test_variants[j] = auc_results$auc
  auc_pval[j] = auc_results$p_value
  #Compute the power calculation
  power_results = compute_auc_power(human_scores, variant_present)
  auc_power_sig[j] = power_results$power_sig
  auc_power_effect_55[j] = power_results$power_effect_55
  auc_power_effect_75[j] = power_results$power_effect_75
  bootstrap_auc_var[j] = power_results$bootstrap_auc_var
  #Get a permuted auroc
  auc_null[j] = compute_auc(human_scores, sample(variant_present))$auc #AUC with the labels randomly shuffled for a null
}

#Get the index for all the SNPs classified as reference biased
ref_index = test_variants %in% human_snp_stat_df$variant[human_snp_stat_df$filter == FALSE]

human_results = list('variants' = test_variants,'reference_biased_index' = ref_index, 'n_pos' = n_pos, 'n_neg' = n_neg, 'variant_prevalence' = variant_prevalence, 
                     'auc' = auc_test_variants, 'auc_pval' = auc_pval, 'auc_power_significance' = auc_power_sig, 'auc_power_effect_55' = auc_power_effect_55,
                     'auc_power_effect_75' = auc_power_effect_75, 'bootstrap_auc_var' = bootstrap_auc_var, 'auc_null' = auc_null)


hist(human_results$auc_null, breaks = seq(0,1,.01), main = 'human skew predicting SNP aurocs: null')
abline(v = .5, col = 'red')
hist(human_results$auc, breaks = seq(0,1,.01), main = 'human skew predicting SNP aurocs')
abline(v = .5, col = 'red')
plot(human_results$auc, human_results$auc_power_significance, main = 'human AUROC against significance power', xlab = 'AUROC', ylab = 'Simulated significance power')
plot(human_results$auc, human_results$auc_power_effect_55, main = 'human effect size power thresh: 0.55', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(human_results$auc, human_results$auc_power_effect_75, main = 'human effect size power thresh: 0.75', xlab = 'AUROC', ylab = 'Simulated effect size power')
plot(human_results$variant_prevalence, human_results$auc_power_effect_55, main = 'human variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(human_results$variant_prevalence, human_results$auc_power_effect_75, main = 'human variant % against effect size power', xlab = 'Variant prevalence', ylab = 'Simulated effect size power')
plot(human_results$variant_prevalence, human_results$bootstrap_auc_var, main = 'human variant % against boostrap auc variance', xlab = 'Variant prevalence', ylab = 'Variance of bootstrap aurocs')
plot(human_results$auc, human_results$bootstrap_auc_var, xlab = 'AUROC', ylab = 'Variance of bootstrap AUROC distribution', main = 'human samples')


hist(human_results$auc_pval, main = 'Raw auroc p-values', breaks = seq(0,1,.05))
adj_pval = p.adjust(human_results$auc_pval, method = 'BH')
hist(adj_pval, main = 'FDR adjusted auroc p-values', breaks = seq(0,1,.05))
fdr_index = which(adj_pval <= .05)
hist(human_results$auc[fdr_index], main = 'Significant aurocs', breaks = seq(0,1,.01))

human_ref_check_df = data.frame(auc = human_results$auc, ref_biased_index = human_results$reference_biased_index)
ggplot(human_ref_check_df, aes(x = ref_biased_index, y = auc, fill = ref_biased_index)) + geom_violin(scale = 'width') + geom_boxplot(width = .25) + ggtitle('human') +
  xlab('Reference biased SNPs') + ylab('XCI ratio and variant AUROC') 
#ggsave(filename = 'human_ref_biased_xci_ratio_auroc_violins_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/',
#       device = 'pdf', useDingbats = F, height = 4, width = 5)

```


Save the ref_check dataframes
```{r}


save(human_ref_check_df, cow_ref_check_df, pig_ref_check_df, sheep_ref_check_df, goat_ref_check_df, horse_ref_check_df, rat_ref_check_df, 
     mouse_ref_check_df, macaca_ref_check_df, dog_ref_check_df, file = '/home/werner/projects/cross_species_XCI/final_plots/R/data_for_plots/all_ref_check_dfs.Rdata')



```





put all the null and test distributions together in a plot

```{r}



all_aucs = c(sheep_results$auc_null, sheep_results$auc, cow_results$auc_null, cow_results$auc, horse_results$auc_null, horse_results$auc,
             goat_results$auc_null, goat_results$auc, dog_results$auc_null, dog_results$auc, pig_results$auc_null, pig_results$auc, 
             macaca_results$auc_null, macaca_results$auc, rat_results$auc_null, rat_results$auc, mouse_results$auc_null, mouse_results$auc, 
  human_results$auc_null, human_results$auc)

all_pvalues = c(rep(NA, length = length(sheep_results$auc_null)), sheep_results$auc_pval, rep(NA, length = length(cow_results$auc_null)), cow_results$auc_pval, 
                rep(NA, length = length(horse_results$auc_null)), horse_results$auc_pval, rep(NA, length = length(goat_results$auc_null)), goat_results$auc_pval, 
                rep(NA, length = length(dog_results$auc_null)), dog_results$auc_pval, rep(NA, length = length(pig_results$auc_null)), pig_results$auc_pval,
                rep(NA, length = length(macaca_results$auc_null)), macaca_results$auc_pval, rep(NA, length = length(rat_results$auc_null)), rat_results$auc_pval, 
                rep(NA, length = length(mouse_results$auc_null)), mouse_results$auc_pval, 
                rep(NA, length = length(human_results$auc_null)), human_results$auc_pval)

all_power_sig = c(rep(NA, length = length(sheep_results$auc_null)), sheep_results$auc_power_significance, 
                  rep(NA, length = length(cow_results$auc_null)), cow_results$auc_power_significance, 
                rep(NA, length = length(horse_results$auc_null)), horse_results$auc_power_significance, 
                rep(NA, length = length(goat_results$auc_null)), goat_results$auc_power_significance, 
                rep(NA, length = length(dog_results$auc_null)), dog_results$auc_power_significance, 
                rep(NA, length = length(pig_results$auc_null)), pig_results$auc_power_significance,
                rep(NA, length = length(macaca_results$auc_null)), macaca_results$auc_power_significance, 
                rep(NA, length = length(rat_results$auc_null)), rat_results$auc_power_significance, 
                rep(NA, length = length(mouse_results$auc_null)), mouse_results$auc_power_significance, 
                rep(NA, length = length(human_results$auc_null)), human_results$auc_power_significance)

all_power_effect_55 = c(rep(NA, length = length(sheep_results$auc_null)), sheep_results$auc_power_effect_55, 
                        rep(NA, length = length(cow_results$auc_null)), cow_results$auc_power_effect_55, 
                rep(NA, length = length(horse_results$auc_null)), horse_results$auc_power_effect_55, 
                rep(NA, length = length(goat_results$auc_null)), goat_results$auc_power_effect_55, 
                rep(NA, length = length(dog_results$auc_null)), dog_results$auc_power_effect_55, 
                rep(NA, length = length(pig_results$auc_null)), pig_results$auc_power_effect_55,
                rep(NA, length = length(macaca_results$auc_null)), macaca_results$auc_power_effect_55, 
                rep(NA, length = length(rat_results$auc_null)), rat_results$auc_power_effect_55, 
                rep(NA, length = length(mouse_results$auc_null)), mouse_results$auc_power_effect_55, 
                rep(NA, length = length(human_results$auc_null)), human_results$auc_power_effect_55)

all_power_effect_75 = c(rep(NA, length = length(sheep_results$auc_null)), sheep_results$auc_power_effect_75, 
                        rep(NA, length = length(cow_results$auc_null)), cow_results$auc_power_effect_75, 
                rep(NA, length = length(horse_results$auc_null)), horse_results$auc_power_effect_75,
                rep(NA, length = length(goat_results$auc_null)), goat_results$auc_power_effect_75, 
                rep(NA, length = length(dog_results$auc_null)), dog_results$auc_power_effect_75, 
                rep(NA, length = length(pig_results$auc_null)), pig_results$auc_power_effect_75,
                rep(NA, length = length(macaca_results$auc_null)), macaca_results$auc_power_effect_75,
                rep(NA, length = length(rat_results$auc_null)), rat_results$auc_power_effect_75, 
                rep(NA, length = length(mouse_results$auc_null)), mouse_results$auc_power_effect_75, 
                rep(NA, length = length(human_results$auc_null)), human_results$auc_power_effect_75)

all_bootstrap_auc_var = c(rep(NA, length = length(sheep_results$auc_null)), sheep_results$bootstrap_auc_var, 
                          rep(NA, length = length(cow_results$auc_null)), cow_results$bootstrap_auc_var, 
                rep(NA, length = length(horse_results$auc_null)), horse_results$bootstrap_auc_var, 
                rep(NA, length = length(goat_results$auc_null)), goat_results$bootstrap_auc_var, 
                rep(NA, length = length(dog_results$auc_null)), dog_results$bootstrap_auc_var, 
                rep(NA, length = length(pig_results$auc_null)), pig_results$bootstrap_auc_var,
                rep(NA, length = length(macaca_results$auc_null)), macaca_results$bootstrap_auc_var,
                rep(NA, length = length(rat_results$auc_null)), rat_results$bootstrap_auc_var, 
                rep(NA, length = length(mouse_results$auc_null)), mouse_results$bootstrap_auc_var, 
                rep(NA, length = length(human_results$auc_null)), human_results$bootstrap_auc_var)

all_n_pos = c(rep(NA, length = length(sheep_results$auc_null)), sheep_results$n_pos, rep(NA, length = length(cow_results$auc_null)), cow_results$n_pos, 
                rep(NA, length = length(horse_results$auc_null)), horse_results$n_pos, rep(NA, length = length(goat_results$auc_null)), goat_results$n_pos, 
                rep(NA, length = length(dog_results$auc_null)), dog_results$n_pos, rep(NA, length = length(pig_results$auc_null)), pig_results$n_pos,
                rep(NA, length = length(macaca_results$auc_null)), macaca_results$n_pos, rep(NA, length = length(rat_results$auc_null)), rat_results$n_pos, 
              rep(NA, length = length(mouse_results$auc_null)), mouse_results$n_pos, 
                rep(NA, length = length(human_results$auc_null)), human_results$n_pos)

all_n_neg = c(rep(NA, length = length(sheep_results$auc_null)), sheep_results$n_neg, rep(NA, length = length(cow_results$auc_null)), cow_results$n_neg, 
                rep(NA, length = length(horse_results$auc_null)), horse_results$n_neg, rep(NA, length = length(goat_results$auc_null)), goat_results$n_neg, 
                rep(NA, length = length(dog_results$auc_null)), dog_results$n_neg, rep(NA, length = length(pig_results$auc_null)), pig_results$n_neg,
                rep(NA, length = length(macaca_results$auc_null)), macaca_results$n_neg, rep(NA, length = length(rat_results$auc_null)), rat_results$n_neg,
               rep(NA, length = length(mouse_results$auc_null)), mouse_results$n_neg, 
                rep(NA, length = length(human_results$auc_null)), human_results$n_neg)

all_variant_prevalence = c(rep(NA, length = length(sheep_results$auc_null)), sheep_results$variant_prevalence,
                           rep(NA, length = length(cow_results$auc_null)), cow_results$variant_prevalence, 
                rep(NA, length = length(horse_results$auc_null)), horse_results$variant_prevalence, 
                rep(NA, length = length(goat_results$auc_null)), goat_results$variant_prevalence, 
                rep(NA, length = length(dog_results$auc_null)), dog_results$variant_prevalence, 
                rep(NA, length = length(pig_results$auc_null)), pig_results$variant_prevalence,
                rep(NA, length = length(macaca_results$auc_null)), macaca_results$variant_prevalence, 
                rep(NA, length = length(rat_results$auc_null)), rat_results$variant_prevalence, 
                rep(NA, length = length(mouse_results$auc_null)), mouse_results$variant_prevalence, 
                rep(NA, length = length(human_results$auc_null)), human_results$variant_prevalence)


all_variants = c(sheep_results$variants, sheep_results$variants, cow_results$variants, cow_results$variants, horse_results$variants, horse_results$variants, 
                 goat_results$variants, goat_results$variants, dog_results$variants, dog_results$variants, 
                 pig_results$variants, pig_results$variants, macaca_results$variants, macaca_results$variants, 
                 rat_results$variants, rat_results$variants, mouse_results$variants, mouse_results$variants, human_results$variants, human_results$variants)

all_ref_biased = c(sheep_results$reference_biased_index, sheep_results$reference_biased_index, cow_results$reference_biased_index, cow_results$reference_biased_index,
                   horse_results$reference_biased_index, horse_results$reference_biased_index, goat_results$reference_biased_index, goat_results$reference_biased_index,
                   dog_results$reference_biased_index, dog_results$reference_biased_index, pig_results$reference_biased_index, pig_results$reference_biased_index,
                   macaca_results$reference_biased_index, macaca_results$reference_biased_index, rat_results$reference_biased_index, rat_results$reference_biased_index,
                   mouse_results$reference_biased_index, mouse_results$reference_biased_index,
                   human_results$reference_biased_index, human_results$reference_biased_index)


null_label = c(rep('null', length = length(sheep_results$auc_null)), rep('data', length = length(sheep_results$auc)),
  rep('null', length = length(cow_results$auc_null)), rep('data', length = length(cow_results$auc)),
  rep('null', length = length(horse_results$auc_null)), rep('data', length = length(horse_results$auc)),
  rep('null', length = length(goat_results$auc_null)), rep('data', length = length(goat_results$auc)),
  rep('null', length = length(dog_results$auc_null)), rep('data', length = length(dog_results$auc)),
  rep('null', length = length(pig_results$auc_null)), rep('data', length = length(pig_results$auc)),
  rep('null', length = length(macaca_results$auc_null)), rep('data', length = length(macaca_results$auc)),
  rep('null', length = length(rat_results$auc_null)), rep('data', length = length(rat_results$auc)),
  rep('null', length = length(mouse_results$auc_null)), rep('data', length = length(mouse_results$auc)),
  rep('null', length = length(human_results$auc_null)), rep('data', length = length(human_results$auc)))

species_label = c(rep('Sheep', length = length(c(sheep_results$auc_null, sheep_results$auc))),
  rep('Cow', length = length(c(cow_results$auc_null, cow_results$auc))),
  rep('Horse', length = length(c(horse_results$auc_null, horse_results$auc))),
  rep('Goat', length = length(c(goat_results$auc_null, goat_results$auc))),
  rep('Dog', length = length(c(dog_results$auc_null, dog_results$auc))),
  rep('Pig', length = length(c(pig_results$auc_null, pig_results$auc))),
  rep('Macaca', length = length(c(macaca_results$auc_null, macaca_results$auc))),
  rep('Rat', length = length(c(rat_results$auc_null, rat_results$auc))),
  rep('Mouse', length = length(c(mouse_results$auc_null, mouse_results$auc))),
  rep('Human', length = length(c(human_results$auc_null, human_results$auc))))


snp_skew_auc_df = data.frame(variants = all_variants, reference_bias_index = all_ref_biased, aucs = all_aucs, p_values = all_pvalues, power_significance = all_power_sig,
                             power_effect_55 = all_power_effect_55, power_effect_75 = all_power_effect_75, bootstrap_auc_var = all_bootstrap_auc_var,
                             n_pos = all_n_pos, n_neg = all_n_neg, 
                             variant_prevalence = all_variant_prevalence, null_label = null_label, species = species_label)
#FDR adjust pvalues and add a significance column
snp_skew_auc_df$adj_p_values = p.adjust(snp_skew_auc_df$p_values, method = 'BH')
snp_skew_auc_df$auc_significance = snp_skew_auc_df$adj_p_values <= .05 
snp_skew_auc_df$auc_significance[is.na(snp_skew_auc_df$p_values)] = NA


species_order_df = snp_skew_auc_df %>% filter(null_label == 'data') %>% group_by(species) %>% summarize(median = median(aucs)) %>% arrange(desc(median))
snp_skew_auc_df$species = factor(snp_skew_auc_df$species, levels = species_order_df$species )



ggplot(filter(snp_skew_auc_df, reference_bias_index == FALSE), aes(y = species, x = aucs, color = null_label, fill = species, alpha = null_label)) + 
  geom_violin(scale = 'width', width = .75) +
  #geom_boxplot() +
  scale_y_discrete(labels = species_order_df$species, expand = c(.0005, 0)) +
  scale_fill_manual(values = species_palette, breaks = levels(snp_skew_auc_df$species)) +
  scale_color_manual(values = c('null' = 'grey', 'data' = 'black')) +
  scale_alpha_manual(values = c('null' = .5, 'data' = 1)) +
  xlim(0, 1) +
  geom_vline(xintercept = .5, color = 'red', linetype = 'dashed')
ggsave(filename = 'skew_snp_auc_violins_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/', device = 'pdf',
       useDingbats = F, height = 6, width = 4)


ggplot(filter(snp_skew_auc_df, reference_bias_index == FALSE), aes(x = species, y = aucs, color = null_label, fill = species, alpha = null_label)) + 
  geom_violin(scale = 'width', width = .75) +
  scale_fill_manual(values = species_palette, breaks = levels(snp_skew_auc_df$species)) +
  scale_color_manual(values = c('null' = 'grey', 'data' = 'black')) +
  scale_alpha_manual(values = c('null' = .5, 'data' = 1)) +
  ylim(0, 1) +
  geom_hline(yintercept = .5, color = 'red', linetype = 'dashed')

ggsave(filename = 'skew_snp_auc_violins_horz_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/', device = 'pdf',
       useDingbats = F, height = 4, width = 8)
```

```{r}

save(snp_skew_auc_df, file = '/home/werner/projects/cross_species_XCI/data/species_data_analysis/snp_skew_auc/snp_skew_auc_df_4_10_24.Rdata')

```


```{r}

load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/snp_skew_auc/snp_skew_auc_df_4_10_24.Rdata')

```



```{r}

ggplot(snp_skew_auc_df, aes(x = aucs, y = power_significance)) + geom_point()
```


Look at the mean expression of each snp and how it relates to the skew AUC

```{r}
#Get the variant, variant position, mean counts per variant, and proportion of samples per variant for each species
#sheep
sheep_variants = filter(snp_skew_auc_df, null_label == 'data' & species == 'Sheep' & reference_bias_index == FALSE)$variants
sheep_variant_auc = filter(snp_skew_auc_df, null_label == 'data' & species == 'Sheep' & reference_bias_index == FALSE)$aucs
sheep_variant_auc_pvals = filter(snp_skew_auc_df, null_label == 'data' & species == 'Sheep' & reference_bias_index == FALSE)$p_values
sheep_variant_auc_power_sig = filter(snp_skew_auc_df, null_label == 'data' & species == 'Sheep' & reference_bias_index == FALSE)$power_significance
sheep_variant_auc_power_effect_55 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Sheep' & reference_bias_index == FALSE)$power_effect_55
sheep_variant_auc_power_effect_75 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Sheep' & reference_bias_index == FALSE)$power_effect_75
sheep_variant_bootstrap_auc_var = filter(snp_skew_auc_df, null_label == 'data' & species == 'Sheep' & reference_bias_index == FALSE)$bootstrap_auc_var
sheep_variant_n_pos = filter(snp_skew_auc_df, null_label == 'data' & species == 'Sheep' & reference_bias_index == FALSE)$n_pos
sheep_variant_n_neg = filter(snp_skew_auc_df, null_label == 'data' & species == 'Sheep' & reference_bias_index == FALSE)$n_neg
sheep_variant_prevalence = filter(snp_skew_auc_df, null_label == 'data' & species == 'Sheep' & reference_bias_index == FALSE)$variant_prevalence
sheep_snp_pos = as.numeric(sapply(1:length(sheep_variants), function(i) substr(sheep_variants[i],1,nchar(sheep_variants[i])-2)))
index = match(sheep_snp_pos, sheep_snp_stat_df$snp_pos)
sheep_exp = sheep_snp_stat_df[index, 'mean_total_counts' ]

sheep_x_meta = sheep_x_gtf[sheep_x_gtf$gbkey == 'Gene'& !is.na(sheep_x_gtf$gbkey),]
x_gene_starts = sheep_x_meta$start
x_gene_ends = sheep_x_meta$end
#Add gene_annotations
sheep_gene_annots = vector(mode = 'character', length = length(sheep_snp_pos))
for( i in 1:length(sheep_snp_pos)){

  test_gene_start = sheep_snp_pos[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    sheep_gene_annots[i] = sheep_x_meta$gene[gene_index]
  }
}


#cow
cow_variants = filter(snp_skew_auc_df, null_label == 'data' & species == 'Cow'& reference_bias_index == FALSE)$variants
cow_variant_auc = filter(snp_skew_auc_df, null_label == 'data' & species == 'Cow' & reference_bias_index == FALSE)$aucs
cow_variant_auc_pvals = filter(snp_skew_auc_df, null_label == 'data' & species == 'Cow' & reference_bias_index == FALSE)$p_values
cow_variant_auc_power_sig = filter(snp_skew_auc_df, null_label == 'data' & species == 'Cow' & reference_bias_index == FALSE)$power_significance
cow_variant_auc_power_effect_55 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Cow' & reference_bias_index == FALSE)$power_effect_55
cow_variant_auc_power_effect_75 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Cow' & reference_bias_index == FALSE)$power_effect_75
cow_variant_bootstrap_auc_var = filter(snp_skew_auc_df, null_label == 'data' & species == 'Cow' & reference_bias_index == FALSE)$bootstrap_auc_var
cow_variant_n_pos = filter(snp_skew_auc_df, null_label == 'data' & species == 'Cow' & reference_bias_index == FALSE)$n_pos
cow_variant_n_neg = filter(snp_skew_auc_df, null_label == 'data' & species == 'Cow' & reference_bias_index == FALSE)$n_neg
cow_variant_prevalence = filter(snp_skew_auc_df, null_label == 'data' & species == 'Cow' & reference_bias_index == FALSE)$variant_prevalence
cow_snp_pos = as.numeric(sapply(1:length(cow_variants), function(i) substr(cow_variants[i],1,nchar(cow_variants[i])-2)))
index = match(cow_snp_pos, cow_snp_stat_df$snp_pos)
cow_exp = cow_snp_stat_df[index, 'mean_total_counts' ]
#cow_prop = cow_snp_stat_df[index, 'num_samples_present' ] / length(annot_cow_vcfs)

cow_x_meta = cow_x_gtf[cow_x_gtf$gbkey == 'Gene'& !is.na(cow_x_gtf$gbkey),]
x_gene_starts = cow_x_meta$start
x_gene_ends = cow_x_meta$end
#Add gene annotations
cow_gene_annots = vector(mode = 'character', length = length(cow_snp_pos))
for( i in 1:length(cow_snp_pos)){

  test_gene_start = cow_snp_pos[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    cow_gene_annots[i] = cow_x_meta$gene[gene_index]
  }
}

#horse
horse_variants = filter(snp_skew_auc_df, null_label == 'data' & species == 'Horse' & reference_bias_index == FALSE)$variants
horse_variant_auc = filter(snp_skew_auc_df, null_label == 'data' & species == 'Horse' & reference_bias_index == FALSE)$aucs
horse_variant_auc_pvals = filter(snp_skew_auc_df, null_label == 'data' & species == 'Horse' & reference_bias_index == FALSE)$p_values
horse_variant_auc_power_sig = filter(snp_skew_auc_df, null_label == 'data' & species == 'Horse' & reference_bias_index == FALSE)$power_significance
horse_variant_auc_power_effect_55 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Horse' & reference_bias_index == FALSE)$power_effect_55
horse_variant_auc_power_effect_75 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Horse' & reference_bias_index == FALSE)$power_effect_75
horse_variant_bootstrap_auc_var = filter(snp_skew_auc_df, null_label == 'data' & species == 'Horse' & reference_bias_index == FALSE)$bootstrap_auc_var
horse_variant_n_pos = filter(snp_skew_auc_df, null_label == 'data' & species == 'Horse' & reference_bias_index == FALSE)$n_pos
horse_variant_n_neg = filter(snp_skew_auc_df, null_label == 'data' & species == 'Horse' & reference_bias_index == FALSE)$n_neg
horse_variant_prevalence = filter(snp_skew_auc_df, null_label == 'data' & species == 'Horse' & reference_bias_index == FALSE)$variant_prevalence
horse_snp_pos = as.numeric(sapply(1:length(horse_variants), function(i) substr(horse_variants[i],1,nchar(horse_variants[i])-2)))
index = match(horse_snp_pos, horse_snp_stat_df$snp_pos)
horse_exp = horse_snp_stat_df[index, 'mean_total_counts' ]
#horse_prop = horse_snp_stat_df[index, 'num_samples_present' ] / length(annot_horse_vcfs)

horse_x_meta = horse_x_gtf[horse_x_gtf$gbkey == 'Gene'& !is.na(horse_x_gtf$gbkey),]
x_gene_starts = horse_x_meta$start
x_gene_ends = horse_x_meta$end
#Add gene_annotations
horse_gene_annots = vector(mode = 'character', length = length(horse_snp_pos))
for( i in 1:length(horse_snp_pos)){

  test_gene_start = horse_snp_pos[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    horse_gene_annots[i] = horse_x_meta$gene[gene_index]
  }
}

#goat
goat_variants = filter(snp_skew_auc_df, null_label == 'data' & species == 'Goat' & reference_bias_index == FALSE)$variants
goat_variant_auc = filter(snp_skew_auc_df, null_label == 'data' & species == 'Goat' & reference_bias_index == FALSE)$aucs
goat_variant_auc_pvals = filter(snp_skew_auc_df, null_label == 'data' & species == 'Goat' & reference_bias_index == FALSE)$p_values
goat_variant_auc_power_sig = filter(snp_skew_auc_df, null_label == 'data' & species == 'Goat' & reference_bias_index == FALSE)$power_significance
goat_variant_auc_power_effect_55 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Goat' & reference_bias_index == FALSE)$power_effect_55
goat_variant_auc_power_effect_75 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Goat' & reference_bias_index == FALSE)$power_effect_75
goat_variant_bootstrap_auc_var = filter(snp_skew_auc_df, null_label == 'data' & species == 'Goat' & reference_bias_index == FALSE)$bootstrap_auc_var
goat_variant_n_pos = filter(snp_skew_auc_df, null_label == 'data' & species == 'Goat' & reference_bias_index == FALSE)$n_pos
goat_variant_n_neg = filter(snp_skew_auc_df, null_label == 'data' & species == 'Goat' & reference_bias_index == FALSE)$n_neg
goat_variant_prevalence = filter(snp_skew_auc_df, null_label == 'data' & species == 'Goat' & reference_bias_index == FALSE)$variant_prevalence
goat_snp_pos = as.numeric(sapply(1:length(goat_variants), function(i) substr(goat_variants[i],1,nchar(goat_variants[i])-2)))
index = match(goat_snp_pos, goat_snp_stat_df$snp_pos)
goat_exp = goat_snp_stat_df[index, 'mean_total_counts' ]
#goat_prop = goat_snp_stat_df[index, 'num_samples_present' ] / length(annot_goat_vcfs)

goat_x_meta = goat_x_gtf[goat_x_gtf$gbkey == 'Gene'& !is.na(goat_x_gtf$gbkey),]
x_gene_starts = goat_x_meta$start
x_gene_ends = goat_x_meta$end
#Add gene_annotations
goat_gene_annots = vector(mode = 'character', length = length(goat_snp_pos))
for( i in 1:length(goat_snp_pos)){

  test_gene_start = goat_snp_pos[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    goat_gene_annots[i] = goat_x_meta$gene[gene_index]
  }
}

#dog
dog_variants = filter(snp_skew_auc_df, null_label == 'data' & species == 'Dog' & reference_bias_index == FALSE)$variants
dog_variant_auc = filter(snp_skew_auc_df, null_label == 'data' & species == 'Dog' & reference_bias_index == FALSE)$aucs
dog_variant_auc_pvals = filter(snp_skew_auc_df, null_label == 'data' & species == 'Dog' & reference_bias_index == FALSE)$p_values
dog_variant_auc_power_sig = filter(snp_skew_auc_df, null_label == 'data' & species == 'Dog' & reference_bias_index == FALSE)$power_significance
dog_variant_auc_power_effect_55 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Dog' & reference_bias_index == FALSE)$power_effect_55
dog_variant_auc_power_effect_75 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Dog' & reference_bias_index == FALSE)$power_effect_75
dog_variant_bootstrap_auc_var = filter(snp_skew_auc_df, null_label == 'data' & species == 'Dog' & reference_bias_index == FALSE)$bootstrap_auc_var
dog_variant_n_pos = filter(snp_skew_auc_df, null_label == 'data' & species == 'Dog' & reference_bias_index == FALSE)$n_pos
dog_variant_n_neg = filter(snp_skew_auc_df, null_label == 'data' & species == 'Dog' & reference_bias_index == FALSE)$n_neg
dog_variant_prevalence = filter(snp_skew_auc_df, null_label == 'data' & species == 'Dog' & reference_bias_index == FALSE)$variant_prevalence
dog_snp_pos = as.numeric(sapply(1:length(dog_variants), function(i) substr(dog_variants[i],1,nchar(dog_variants[i])-2)))
index = match(dog_snp_pos, dog_snp_stat_df$snp_pos)
dog_exp = dog_snp_stat_df[index, 'mean_total_counts' ]
#dog_prop = dog_snp_stat_df[index, 'num_samples_present' ] / length(annot_dog_vcfs)

dog_x_meta = dog_x_gtf[dog_x_gtf$gbkey == 'Gene',]
dog_x_meta = dog_x_meta[!is.na(dog_x_meta$gbkey), ] #A lot of NA entries in the dog GTF file
dog_x_meta = dog_x_meta[!duplicated(dog_x_meta$gene), ]
x_gene_starts = dog_x_meta$start
x_gene_ends = dog_x_meta$end
#Add gene_annotations
dog_gene_annots = vector(mode = 'character', length = length(dog_snp_pos))
for( i in 1:length(dog_snp_pos)){

  test_gene_start = dog_snp_pos[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    dog_gene_annots[i] = dog_x_meta$gene[gene_index]
  }
}

#pig
pig_variants = filter(snp_skew_auc_df, null_label == 'data' & species == 'Pig' & reference_bias_index == FALSE)$variants
pig_variant_auc = filter(snp_skew_auc_df, null_label == 'data' & species == 'Pig' & reference_bias_index == FALSE)$aucs
pig_variant_auc_pvals = filter(snp_skew_auc_df, null_label == 'data' & species == 'Pig' & reference_bias_index == FALSE)$p_values
pig_variant_auc_power_sig = filter(snp_skew_auc_df, null_label == 'data' & species == 'Pig' & reference_bias_index == FALSE)$power_significance
pig_variant_auc_power_effect_55 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Pig' & reference_bias_index == FALSE)$power_effect_55
pig_variant_auc_power_effect_75 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Pig' & reference_bias_index == FALSE)$power_effect_75
pig_variant_bootstrap_auc_var = filter(snp_skew_auc_df, null_label == 'data' & species == 'Pig' & reference_bias_index == FALSE)$bootstrap_auc_var
pig_variant_n_pos = filter(snp_skew_auc_df, null_label == 'data' & species == 'Pig' & reference_bias_index == FALSE)$n_pos
pig_variant_n_neg = filter(snp_skew_auc_df, null_label == 'data' & species == 'Pig' & reference_bias_index == FALSE)$n_neg
pig_variant_prevalence = filter(snp_skew_auc_df, null_label == 'data' & species == 'Pig' & reference_bias_index == FALSE)$variant_prevalence
pig_snp_pos = as.numeric(sapply(1:length(pig_variants), function(i) substr(pig_variants[i],1,nchar(pig_variants[i])-2)))
index = match(pig_snp_pos, pig_snp_stat_df$snp_pos)
pig_exp = pig_snp_stat_df[index, 'mean_total_counts' ]
#pig_prop = pig_snp_stat_df[index, 'num_samples_present' ] / length(annot_pig_vcfs)

pig_x_meta = pig_x_gtf[pig_x_gtf$gbkey == 'Gene'& !is.na(pig_x_gtf$gbkey),]
x_gene_starts = pig_x_meta$start
x_gene_ends = pig_x_meta$end
#Add gene_annotations
pig_gene_annots = vector(mode = 'character', length = length(pig_snp_pos))
for( i in 1:length(pig_snp_pos)){

  test_gene_start = pig_snp_pos[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    pig_gene_annots[i] = pig_x_meta$gene[gene_index]
  }
}

#macaca
macaca_variants = filter(snp_skew_auc_df, null_label == 'data' & species == 'Macaca' & reference_bias_index == FALSE)$variants
macaca_variant_auc = filter(snp_skew_auc_df, null_label == 'data' & species == 'Macaca' & reference_bias_index == FALSE)$aucs
macaca_variant_auc_pvals = filter(snp_skew_auc_df, null_label == 'data' & species == 'Macaca' & reference_bias_index == FALSE)$p_values
macaca_variant_auc_power_sig = filter(snp_skew_auc_df, null_label == 'data' & species == 'Macaca' & reference_bias_index == FALSE)$power_significance
macaca_variant_auc_power_effect_55 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Macaca' & reference_bias_index == FALSE)$power_effect_55
macaca_variant_auc_power_effect_75 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Macaca' & reference_bias_index == FALSE)$power_effect_75
macaca_variant_bootstrap_auc_var = filter(snp_skew_auc_df, null_label == 'data' & species == 'Macaca' & reference_bias_index == FALSE)$bootstrap_auc_var
macaca_variant_n_pos = filter(snp_skew_auc_df, null_label == 'data' & species == 'Macaca' & reference_bias_index == FALSE)$n_pos
macaca_variant_n_neg = filter(snp_skew_auc_df, null_label == 'data' & species == 'Macaca' & reference_bias_index == FALSE)$n_neg
macaca_variant_prevalence = filter(snp_skew_auc_df, null_label == 'data' & species == 'Macaca' & reference_bias_index == FALSE)$variant_prevalence
macaca_snp_pos = as.numeric(sapply(1:length(macaca_variants), function(i) substr(macaca_variants[i],1,nchar(macaca_variants[i])-2)))
index = match(macaca_snp_pos, macaca_snp_stat_df$snp_pos)
macaca_exp = macaca_snp_stat_df[index, 'mean_total_counts' ]
#macaca_prop = macaca_snp_stat_df[index, 'num_samples_present' ] / length(annot_macaca_vcfs)

macaca_x_meta = macaca_x_gtf[macaca_x_gtf$gbkey == 'Gene'& !is.na(macaca_x_gtf$gbkey),]
x_gene_starts = macaca_x_meta$start
x_gene_ends = macaca_x_meta$end
#Add gene_annotations
macaca_gene_annots = vector(mode = 'character', length = length(macaca_snp_pos))
for( i in 1:length(macaca_snp_pos)){

  test_gene_start = macaca_snp_pos[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    macaca_gene_annots[i] = macaca_x_meta$gene[gene_index]
  }
}

#rat
rat_variants = filter(snp_skew_auc_df, null_label == 'data' & species == 'Rat' & reference_bias_index == FALSE)$variants
rat_variant_auc = filter(snp_skew_auc_df, null_label == 'data' & species == 'Rat' & reference_bias_index == FALSE)$aucs
rat_variant_auc_pvals = filter(snp_skew_auc_df, null_label == 'data' & species == 'Rat' & reference_bias_index == FALSE)$p_values
rat_variant_auc_power_sig = filter(snp_skew_auc_df, null_label == 'data' & species == 'Rat' & reference_bias_index == FALSE)$power_significance
rat_variant_auc_power_effect_55 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Rat' & reference_bias_index == FALSE)$power_effect_55
rat_variant_auc_power_effect_75 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Rat' & reference_bias_index == FALSE)$power_effect_75
rat_variant_bootstrap_auc_var = filter(snp_skew_auc_df, null_label == 'data' & species == 'Rat' & reference_bias_index == FALSE)$bootstrap_auc_var
rat_variant_n_pos = filter(snp_skew_auc_df, null_label == 'data' & species == 'Rat' & reference_bias_index == FALSE)$n_pos
rat_variant_n_neg = filter(snp_skew_auc_df, null_label == 'data' & species == 'Rat' & reference_bias_index == FALSE)$n_neg
rat_variant_prevalence = filter(snp_skew_auc_df, null_label == 'data' & species == 'Rat' & reference_bias_index == FALSE)$variant_prevalence
rat_snp_pos = as.numeric(sapply(1:length(rat_variants), function(i) substr(rat_variants[i],1,nchar(rat_variants[i])-2)))
index = match(rat_snp_pos, rat_snp_stat_df$snp_pos)
rat_exp = rat_snp_stat_df[index, 'mean_total_counts' ]
#rat_prop = rat_snp_stat_df[index, 'num_samples_present' ] / length(annot_rat_vcfs)

rat_x_meta = rat_x_gtf[rat_x_gtf$gbkey == 'Gene'& !is.na(rat_x_gtf$gbkey),]
x_gene_starts = rat_x_meta$start
x_gene_ends = rat_x_meta$end
#Add gene_annotations
rat_gene_annots = vector(mode = 'character', length = length(rat_snp_pos))
for( i in 1:length(rat_snp_pos)){

  test_gene_start = rat_snp_pos[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    rat_gene_annots[i] = rat_x_meta$gene[gene_index]
  }
}


#Mouse
mouse_variants = filter(snp_skew_auc_df, null_label == 'data' & species == 'Mouse' & reference_bias_index == FALSE)$variants
mouse_variant_auc = filter(snp_skew_auc_df, null_label == 'data' & species == 'Mouse' & reference_bias_index == FALSE)$aucs
mouse_variant_auc_pvals = filter(snp_skew_auc_df, null_label == 'data' & species == 'Mouse' & reference_bias_index == FALSE)$p_values
mouse_variant_auc_power_sig = filter(snp_skew_auc_df, null_label == 'data' & species == 'Mouse' & reference_bias_index == FALSE)$power_significance
mouse_variant_auc_power_effect_55 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Mouse' & reference_bias_index == FALSE)$power_effect_55
mouse_variant_auc_power_effect_75 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Mouse' & reference_bias_index == FALSE)$power_effect_75
mouse_variant_bootstrap_auc_var = filter(snp_skew_auc_df, null_label == 'data' & species == 'Mouse' & reference_bias_index == FALSE)$bootstrap_auc_var
mouse_variant_n_pos = filter(snp_skew_auc_df, null_label == 'data' & species == 'Mouse' & reference_bias_index == FALSE)$n_pos
mouse_variant_n_neg = filter(snp_skew_auc_df, null_label == 'data' & species == 'Mouse' & reference_bias_index == FALSE)$n_neg
mouse_variant_prevalence = filter(snp_skew_auc_df, null_label == 'data' & species == 'Mouse' & reference_bias_index == FALSE)$variant_prevalence
mouse_snp_pos = as.numeric(sapply(1:length(mouse_variants), function(i) substr(mouse_variants[i],1,nchar(mouse_variants[i])-2)))
index = match(mouse_snp_pos, mouse_snp_stat_df$snp_pos)
mouse_exp = mouse_snp_stat_df[index, 'mean_total_counts' ]
#mouse_prop = mouse_snp_stat_df[index, 'num_samples_present' ] / length(annot_mouse_vcfs)

mouse_x_meta = mouse_x_gtf[mouse_x_gtf$gbkey == 'Gene'& !is.na(mouse_x_gtf$gbkey),]
x_gene_starts = mouse_x_meta$start
x_gene_ends = mouse_x_meta$end
#Add gene_annotations
mouse_gene_annots = vector(mode = 'character', length = length(mouse_snp_pos))
for( i in 1:length(mouse_snp_pos)){

  test_gene_start = mouse_snp_pos[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    mouse_gene_annots[i] = mouse_x_meta$gene[gene_index]
  }
}


```


Go through human samples since I didnt gather the snp info prior like I did with the other species

```{r}

human_variants = filter(snp_skew_auc_df, null_label == 'data' & species == 'Human' & reference_bias_index == FALSE)$variants
human_variant_auc = filter(snp_skew_auc_df, null_label == 'data' & species == 'Human' & reference_bias_index == FALSE)$aucs
human_variant_auc_pvals = filter(snp_skew_auc_df, null_label == 'data' & species == 'Human' & reference_bias_index == FALSE)$p_values
human_variant_auc_power_sig = filter(snp_skew_auc_df, null_label == 'data' & species == 'Human' & reference_bias_index == FALSE)$power_significance
human_variant_auc_power_effect_55 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Human' & reference_bias_index == FALSE)$power_effect_55
human_variant_auc_power_effect_75 = filter(snp_skew_auc_df, null_label == 'data' & species == 'Human' & reference_bias_index == FALSE)$power_effect_75
human_variant_bootstrap_auc_var = filter(snp_skew_auc_df, null_label == 'data' & species == 'Human' & reference_bias_index == FALSE)$bootstrap_auc_var
human_variant_n_pos = filter(snp_skew_auc_df, null_label == 'data' & species == 'Human' & reference_bias_index == FALSE)$n_pos
human_variant_n_neg = filter(snp_skew_auc_df, null_label == 'data' & species == 'Human' & reference_bias_index == FALSE)$n_neg
human_variant_prevalence = filter(snp_skew_auc_df, null_label == 'data' & species == 'Human' & reference_bias_index == FALSE)$variant_prevalence
human_snp_pos = as.numeric(str_extract(human_variants, "[0-9]+")) # extract positions from human variants

#Go through and get the mean counts and proportions
human_samples_index = human_skew_and_stats_df$sample_index
snp_human_counts_mat = matrix(nrow = length(human_variants), ncol = length(human_samples_index))
rownames(snp_human_counts_mat) = human_variants

for(i in 1:length(human_samples_index)){
  
  data = list.vcf[[human_samples_index[i]]]
  data_variants = paste0(data[ ,1], paste0(data$V3, data$V4)) #Get the sample's variants
  data_variants_counts = data[ ,2] + data[ ,3] #Get the total counts
  index = match(data_variants, rownames(snp_human_counts_mat)) #Put counts in the full matrix
  na_index = !is.na(index) #Ignoring the SNPs not in the human variants list, not detected in at least 10 samples
  snp_human_counts_mat[index[na_index], i] = data_variants_counts[na_index]
}
#Get the mean expression for all snps and the sample proportion snp is detected
human_exp = rowMeans(snp_human_counts_mat, na.rm = T)
#human_prop = rowSums(!is.na(snp_human_counts_mat)) / length(human_samples_index)

human_x_meta = human_x_gtf[human_x_gtf$type == 'gene'& !is.na(human_x_gtf$type),]
x_gene_starts = human_x_meta$start
x_gene_ends = human_x_meta$end
#Add gene_annotations
human_gene_annots = vector(mode = 'character', length = length(human_snp_pos))
for( i in 1:length(human_snp_pos)){

  test_gene_start = human_snp_pos[i]
  gene_index = findInterval(test_gene_start, x_gene_starts)
  if(gene_index == 0){next}
  
  if(test_gene_start <= x_gene_ends[gene_index]){
    human_gene_annots[i] = human_x_meta$gene_name[gene_index]
  }
}

#Ignore the SNPs with 0 exp, not sure where those crept into the human data
snp_index = human_exp > 0
human_variants = human_variants[snp_index]
human_snp_pos = human_snp_pos[snp_index]
human_gene_annots = human_gene_annots[snp_index]
human_variant_auc = human_variant_auc[snp_index]
human_variant_auc_pvals = human_variant_auc_pvals[snp_index]
human_variant_auc_power_sig = human_variant_auc_power_sig[snp_index]
human_variant_auc_power_effect_55 = human_variant_auc_power_effect_55[snp_index]
human_variant_auc_power_effect_75 = human_variant_auc_power_effect_75[snp_index]
human_variant_bootstrap_auc_var = human_variant_bootstrap_auc_var[snp_index]
human_variant_n_pos = human_variant_n_pos[snp_index]
human_variant_n_neg = human_variant_n_neg[snp_index]
human_variant_prevalence = human_variant_prevalence[snp_index]
human_exp = human_exp[snp_index]
#human_prop = human_prop[snp_index]
```


```{r}

all_variants = c(sheep_variants, cow_variants, horse_variants, goat_variants, 
                 dog_variants, pig_variants, macaca_variants, rat_variants, mouse_variants, human_variants)
all_positions = c(sheep_snp_pos, cow_snp_pos, horse_snp_pos, goat_snp_pos, 
                  dog_snp_pos, pig_snp_pos, macaca_snp_pos, rat_snp_pos, mouse_snp_pos, human_snp_pos)
all_genes = c(sheep_gene_annots, cow_gene_annots, horse_gene_annots, goat_gene_annots, 
              dog_gene_annots, pig_gene_annots, macaca_gene_annots, rat_gene_annots, mouse_gene_annots, human_gene_annots)
all_aucs = c(sheep_variant_auc, cow_variant_auc, horse_variant_auc, goat_variant_auc, 
             dog_variant_auc, pig_variant_auc, macaca_variant_auc, rat_variant_auc, mouse_variant_auc, human_variant_auc)
all_auc_pvals = c(sheep_variant_auc_pvals, cow_variant_auc_pvals, horse_variant_auc_pvals, goat_variant_auc_pvals, 
                  dog_variant_auc_pvals, pig_variant_auc_pvals, 
                  macaca_variant_auc_pvals, rat_variant_auc_pvals, mouse_variant_auc_pvals, human_variant_auc_pvals)

all_adj_pvals = p.adjust(all_auc_pvals, method = 'BH')
auc_significance = all_adj_pvals <= .05
all_power_sig = c(sheep_variant_auc_power_sig, cow_variant_auc_power_sig, horse_variant_auc_power_sig, goat_variant_auc_power_sig, 
                  dog_variant_auc_power_sig, pig_variant_auc_power_sig,
              macaca_variant_auc_power_sig, rat_variant_auc_power_sig, mouse_variant_auc_power_sig, human_variant_auc_power_sig)

all_power_effect_55 = c(sheep_variant_auc_power_effect_55, cow_variant_auc_power_effect_55, horse_variant_auc_power_effect_55, goat_variant_auc_power_effect_55, 
                        dog_variant_auc_power_effect_55, pig_variant_auc_power_effect_55, macaca_variant_auc_power_effect_55, rat_variant_auc_power_effect_55,
                        mouse_variant_auc_power_effect_55, human_variant_auc_power_effect_55)

all_power_effect_75 = c(sheep_variant_auc_power_effect_75, cow_variant_auc_power_effect_75, horse_variant_auc_power_effect_75, goat_variant_auc_power_effect_75, 
                        dog_variant_auc_power_effect_75, pig_variant_auc_power_effect_75, macaca_variant_auc_power_effect_75, rat_variant_auc_power_effect_75,
                        mouse_variant_auc_power_effect_75, human_variant_auc_power_effect_75)

all_bootstrap_auc_var = c(sheep_variant_bootstrap_auc_var, cow_variant_bootstrap_auc_var, horse_variant_bootstrap_auc_var, goat_variant_bootstrap_auc_var, 
                          dog_variant_bootstrap_auc_var, pig_variant_bootstrap_auc_var, macaca_variant_bootstrap_auc_var, rat_variant_bootstrap_auc_var,
                          mouse_variant_bootstrap_auc_var, human_variant_bootstrap_auc_var)

all_exp = c(sheep_exp, cow_exp, horse_exp, goat_exp, dog_exp, pig_exp, macaca_exp, rat_exp, mouse_exp, human_exp)
all_n_pos = c(sheep_variant_n_pos, cow_variant_n_pos, horse_variant_n_pos, goat_variant_n_pos, dog_variant_n_pos, pig_variant_n_pos,
              macaca_variant_n_pos, rat_variant_n_pos, mouse_variant_n_pos, human_variant_n_pos)
all_n_neg = c(sheep_variant_n_neg, cow_variant_n_neg, horse_variant_n_neg, goat_variant_n_neg, dog_variant_n_neg, pig_variant_n_neg,
              macaca_variant_n_neg, rat_variant_n_neg, mouse_variant_n_neg, human_variant_n_neg)
all_variant_prevalence = c(sheep_variant_prevalence, cow_variant_prevalence, horse_variant_prevalence, goat_variant_prevalence, 
                           dog_variant_prevalence, pig_variant_prevalence,macaca_variant_prevalence, rat_variant_prevalence,
                           mouse_variant_prevalence, human_variant_prevalence)
species_label = c(rep('Sheep', length = length(sheep_variants)), rep('Cow', length = length(cow_variants)), rep('Horse', length = length(horse_variants)), 
                  rep('Goat', length = length(goat_variants)), rep('Dog', length = length(dog_variants)), rep('Pig', length = length(pig_variants)),
                  rep('Macaca', length = length(macaca_variants)), rep('Rat', length = length(rat_variants)), 
                  rep('Mouse', length = length(mouse_variants)), rep('Human', length = length(human_variants)))

snp_skew_auc_with_meta_df = data.frame(variants = all_variants, chrom_pos = all_positions, gene = all_genes, skew_auc = all_aucs, skew_auc_pval = all_auc_pvals, skew_auc_adj_pval = all_adj_pvals,
                                       skew_auc_significance = auc_significance, skew_auc_power_significance = all_power_sig,
                                       skew_auc_power_effect_55 = all_power_effect_55, skew_auc_power_effect_75 = all_power_effect_75,
                                       skew_auc_bootstrap_var = all_bootstrap_auc_var,
                                       skew_auc_n_pos = all_n_pos, skew_auc_n_neg = all_n_neg,  
                                       variant_prevalence = all_variant_prevalence, expression = all_exp, label = species_label)

snp_skew_auc_with_meta_df


ggplot(snp_skew_auc_with_meta_df, aes(x = log10(expression), y = skew_auc, color = label)) + geom_point(size = .5, alpha = .25) + facet_wrap(~label, ncol = 3) +
  ylim(0, 1) + geom_hline(yintercept = .5, linetype = 'dashed', color = 'red') +
  scale_color_manual(values = species_palette)
ggsave(filename = 'skewSNP_auc_and_expression_scatter_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/', 
       device = 'pdf', useDingbats = F, height = 6, width = 8)


ggplot(snp_skew_auc_with_meta_df, aes(y = variant_prevalence, x = skew_auc, color = label)) + geom_point(size = .5, alpha = .25) + facet_wrap(~label, ncol = 3) +
  xlim(0, 1) + geom_vline(xintercept = .5, linetype = 'dashed', color = 'red') +
  scale_color_manual(values = species_palette)
ggsave(filename = 'skewSNP_auc_and_sample_prop_scatter_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/', 
       device = 'pdf', useDingbats = F, height = 6, width = 8)

ggplot(snp_skew_auc_with_meta_df, aes(y = variant_prevalence, x = skew_auc)) + geom_point(size = 1, alpha = .25) + 
  geom_vline(xintercept = .5, linetype = 'dashed', color = 'red') +
  scale_color_manual(values = species_palette) +
  xlab('Strength of XCI ratio and variant association (AUROC)') + ylab('Variant prevalence')
ggsave(filename = 'all_species_skewSNP_sample_prop_scatter_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/', 
       device = 'pdf', useDingbats = F, height = 4, width = 5)


```


```{r}
snp_skew_auc_with_meta_df


ggplot(snp_skew_auc_with_meta_df, aes(y = variant_prevalence, x = skew_auc, color = skew_auc_significance)) + geom_point(size = 1, alpha = .25) + 
  geom_vline(xintercept = .75, linetype = 'dashed', color = 'red') +
  scale_color_manual(values = c('TRUE' = 'red', 'FALSE' = 'black'), name = 'AUROC significance') +
  xlab('Strength of XCI ratio and variant association (AUROC)') + ylab('Variant prevalence')
ggsave(filename = 'all_species_skewSNP_sample_prop_scatter_sig_label_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/', 
       device = 'pdf', useDingbats = F, height = 4, width = 6)


ggplot(snp_skew_auc_with_meta_df, aes(y = -log10(skew_auc_adj_pval), x = skew_auc, color = skew_auc_significance)) + geom_point(size = 1, alpha = .25) + 
  geom_vline(xintercept = .75, linetype = 'dashed', color = 'red') +
  scale_color_manual(values = c('TRUE' = 'red', 'FALSE' = 'black'), name = 'AUROC significance') +
  #scale_x_continuous(breaks=c(0, 0.25, 0.50, 0.75, 0.80, 1)) +
  xlab('Strength of XCI ratio and variant association (AUROC)') + ylab('-log10( FDR-adjusted p-value)')

ggsave(filename = 'all_species_skewSNP_auc_volcano_sig_label_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/', 
       device = 'pdf', useDingbats = F, height = 4, width = 6)



ggplot(filter(snp_skew_auc_with_meta_df, skew_auc > .5), aes(y = -log10(skew_auc_adj_pval), x = variant_prevalence, color = skew_auc_significance)) + geom_point(size = 1, alpha = .25) + 
  scale_color_manual(values = c('TRUE' = 'red', 'FALSE' = 'black'), name = 'AUROC significance') +
  #scale_x_continuous(breaks=c(0, 0.25, 0.50, 0.75, 0.80, 1)) +
  xlab('Variant prevalence') + ylab('-log10( FDR-adjusted p-value)') + ggtitle('AUROC > 0.50')

ggplot(filter(snp_skew_auc_with_meta_df, skew_auc < .5), aes(y = -log10(skew_auc_adj_pval), x = variant_prevalence, color = skew_auc_significance)) + geom_point(size = 1, alpha = .25) + 
  scale_color_manual(values = c('TRUE' = 'red', 'FALSE' = 'black'), name = 'AUROC significance') +
  #scale_x_continuous(breaks=c(0, 0.25, 0.50, 0.75, 0.80, 1)) +
  xlab('Variant prevalence') + ylab('-log10( FDR-adjusted p-value)') + ggtitle('AUROC < 0.50')




ggplot(snp_skew_auc_with_meta_df, aes(y = skew_auc_power_significance, x = variant_prevalence, color = skew_auc_significance)) + geom_point(size = 1, alpha = .25) + 
  scale_color_manual(values = c('TRUE' = 'red', 'FALSE' = 'black'), name = 'AUROC significance') +
  xlab('Variant prevalence') + ylab('Estimated AUROC power')
ggsave(filename = 'all_species_skewSNP_prevalence_and_power_scatter_sig_label_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/', 
       device = 'pdf', useDingbats = F, height = 4, width = 6)

ggplot(snp_skew_auc_with_meta_df, aes(y = skew_auc_power_significance, x = skew_auc, color = skew_auc_significance)) + geom_point(size = 1, alpha = .25) + 
  scale_color_manual(values = c('TRUE' = 'red', 'FALSE' = 'black'), name = 'AUROC significance') +
  xlab('AUROC') + ylab('Estimated AUROC power') +
  geom_vline(xintercept = .75, linetype = 'dashed', color = 'red')
ggsave(filename = 'all_species_skewSNP_auroc_and_power_scatter_sig_label_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/', 
       device = 'pdf', useDingbats = F, height = 4, width = 6)

```






```{r}

ggplot(snp_skew_auc_with_meta_df, aes(y = skew_auc_power_effect_55, x = skew_auc, color = skew_auc_significance)) + geom_point(size = 1, alpha = .25) + 
  scale_color_manual(values = c('TRUE' = 'red', 'FALSE' = 'black'), name = 'AUROC significance') +
  geom_vline(xintercept = .75, linetype = 'dashed', color = 'red') +
  xlab('AUROC') + ylab('Estimated AUROC power effect size .55')
ggsave(filename = 'all_species_skewSNP_skew_and_power_effect55_scatter_sig_label_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/', 
       device = 'pdf', useDingbats = F, height = 4, width = 6)

ggplot(snp_skew_auc_with_meta_df, aes(y = skew_auc_power_effect_75, x = skew_auc, color = skew_auc_significance)) + geom_point(size = 1, alpha = .25) + 
  scale_color_manual(values = c('TRUE' = 'red', 'FALSE' = 'black'), name = 'AUROC significance') +
  geom_vline(xintercept = .75, linetype = 'dashed', color = 'red') +
  xlab('AUROC') + ylab('Estimated AUROC power effect size .75')
ggsave(filename = 'all_species_skewSNP_skew_and_power_effect75_scatter_sig_label_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/', 
       device = 'pdf', useDingbats = F, height = 4, width = 6)

ggplot(snp_skew_auc_with_meta_df, aes(y = skew_auc_bootstrap_var, x = skew_auc, color = skew_auc_significance)) + geom_point(size = 1, alpha = .25) + 
  scale_color_manual(values = c('TRUE' = 'red', 'FALSE' = 'black'), name = 'AUROC significance') +
  geom_vline(xintercept = .75, linetype = 'dashed', color = 'red') +
  xlab('AUROC') + ylab('Variance of bootstrapped AUROC distribution')
ggsave(filename = 'all_species_skewSNP_skew_and_bootstrapVar_scatter_sig_label_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/', 
       device = 'pdf', useDingbats = F, height = 4, width = 6)


```







```{r}

temp_skew_snp_df = snp_skew_auc_with_meta_df %>% filter(skew_auc >= .75 & skew_auc_significance == TRUE & label != 'Human')
temp_skew_snp_df = temp_skew_snp_df %>% group_by(label, gene) %>% mutate(gene = replace(gene,  expression < max(expression), ""))
#Keep the full annotated gene list to include in the paper text
temp_skew_snp_df

ggplot(temp_skew_snp_df, aes(x = skew_auc, y = variant_prevalence, color = label)) + geom_point(size = 2,show.legend = F) +
  scale_color_manual(values = species_palette) + facet_wrap(~label,ncol = 3, scales = 'free_y') + geom_label_repel(aes(label = gene), size = 3, force = 3, show.legend = F, max.overlaps = 15) +
  xlab('Strength of XCI ratio and variant association (AUROC)') + ylab('Variant prevalence') 

ggsave(filename = 'gene_labeled_skewSNP_auc_scatter_v3.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/', 
       device = 'pdf', useDingbats = F, height = 7, width = 9)

```


Rat XIST is LOC100911498

```{r}
temp_skew_snp_df

rat_x_meta
```

```{r}

temp_skew_snp_df = snp_skew_auc_with_meta_df %>% filter(skew_auc >= .75 & label == 'Rat' & skew_auc_significance == TRUE)
#Adding the rat XIST label
temp_skew_snp_df$gene[temp_skew_snp_df$gene == 'LOC100911498'] = 'XIST'

ggplot(temp_skew_snp_df, aes(x = chrom_pos, y = skew_auc))  + geom_point(size = 3, color = species_palette['Rat']) +
  geom_vline(xintercept = 68474987, color = 'red', size = 1) + geom_label_repel(aes(label = gene), size = 3, force = 3, show.legend = F ) +
  ylab('Strength of XCI ratio and variant association (AUROC)') + xlab('Chromosome position')
ggsave(filename = 'rat_chromPos_skewSNP_auc_scatter_v2.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/', 
       device = 'pdf', useDingbats = F, height = 4, width = 6)

```

check if those variants are present at all

```{r}
temp_skew_snp_df = snp_skew_auc_with_meta_df %>% filter(label == 'Rat')
temp_skew_snp_df[temp_skew_snp_df$gene == 'LOC100911498', ]
```


```{r}


save(snp_skew_auc_with_meta_df, file = '/home/werner/projects/cross_species_XCI/data/species_data_analysis/snp_skew_auc/snp_skew_auc_with_meta_4_10_24.Rdata')

```



```{r}


load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/snp_skew_auc/snp_skew_auc_with_meta_4_10_24.Rdata')

```


```{r}
library(ComplexHeatmap)
genus_samples_df = read.table(file ='/home/werner/projects/cross_species_XCI/code/all_mammal_metadata.tsv', sep = '\t', header = T, fill = T, quote = "")
study_1_df = genus_samples_df[genus_samples_df$SRAStudy == 'SRP296549', ]
study_3_df = genus_samples_df[genus_samples_df$SRAStudy == 'SRP125176', ]

all_xist_mouse_pos = c()
for(i in 1:length(annot_mouse_vcfs)){
  data = annot_mouse_vcfs[[i]]
  all_xist_mouse_pos = c(all_xist_mouse_pos, data$Pos[data$gene == 'Xist'])
}
all_xist_mouse_pos = unique(all_xist_mouse_pos)
length(all_xist_mouse_pos)



#mouse_xist_variants = snp_skew_auc_with_meta_df %>% filter(label == 'Mouse' & gene == 'Xist') %>% pull(chrom_pos)
mouse_xist_present_mat = matrix(nrow = length(all_xist_mouse_pos), ncol = length(annot_mouse_vcfs))
colnames(mouse_xist_present_mat) = mouse_skew_and_stats_df$sample_id
rownames(mouse_xist_present_mat) = paste0('variant_', 1:length(all_xist_mouse_pos))


for(i in 1:ncol(mouse_xist_present_mat)){
  data = annot_mouse_vcfs[[i]]
  mouse_xist_present_mat[ ,i] = as.numeric(all_xist_mouse_pos %in% data$Pos)
}

#Only considering mouse samples with at least 20 XIST snps
sample_index = colSums(mouse_xist_present_mat) >= 20
mouse_xist_present_mat = mouse_xist_present_mat[, sample_index]


mouse_hap_df = mouse_skew_and_stats_df[sample_index, ] %>% select(est_skew_no_escape, num_good_snps_no_escape, sample_id)
mouse_hap_df$study =  rep('other', length = nrow(mouse_hap_df))
mouse_hap_df$study[mouse_hap_df$sample_id %in% study_1_df$Run] = 'Study_1_striatum'
mouse_hap_df$study[mouse_hap_df$sample_id %in% study_3_df$Run] = 'Study_2_pancreas'


col_annot = HeatmapAnnotation(study = mouse_hap_df$study,
                              col = list(study = c('Study_1_striatum' = '#F8766D', 'Study_2_pancreas' = '#00BFC4')))
colors = structure(1:2, names = c(0, 1)) # black, red
h1 = Heatmap(mouse_xist_present_mat, name = "mat", col = colors,top_annotation = col_annot,
        clustering_distance_columns = function(m) dist(m, method = 'binary'),
        clustering_distance_rows = function(m) dist(m, method = 'binary'),
        show_column_names = F, show_row_names = F,
        clustering_method_rows = "ward.D2", clustering_method_columns = "ward.D2", show_column_dend = T, column_split = 4)

h1 = draw(h1)

#pdf(file = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/mouse_xist_vars/mouse_xist_clust_matrix.pdf',
#    height = 4, width = 6, useDingbats = F)
draw(h1)
#dev.off()


mouse_hap_df$label = rep('other', length = nrow(mouse_hap_df))
mouse_hap_df$label[column_order(h1)[[1]]] = 'haplotype_group 1'
mouse_hap_df$label[column_order(h1)[[2]]] = 'haplotype_group 2'
mouse_hap_df$label[column_order(h1)[[3]]] = 'haplotype_group 3'
mouse_hap_df$label[column_order(h1)[[4]]] = 'haplotype_group 4'

ggplot(mouse_hap_df, aes(x = study, y = est_skew_no_escape, color = label)) + geom_boxplot() + xlab('Mouse XIST putative haplotypes') + ylab('Estimated XCI ratio') +
  geom_point(position=position_dodge(width=0.75),aes(group=label), color = 'grey50')
#ggsave(filename = 'mouse_xist_x_axis_study.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/mouse_xist_vars/', device = 'pdf',
#       useDingbats = F, height = 4, width = 6)

ggplot(mouse_hap_df, aes(x = label, y = est_skew_no_escape, color = study)) + geom_boxplot() + xlab('Mouse XIST putative haplotypes') + ylab('Estimated XCI ratio') +
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  geom_point(position=position_dodge(width=0.75),aes(group=study), color = 'grey50')
#ggsave(filename = 'mouse_xist_x_axis_haplotype.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/mouse_xist_vars/', device = 'pdf',
#       useDingbats = F, height = 4, width = 6)



```




```{r}


save(mouse_xist_present_mat,mouse_hap_df, file = '/home/werner/projects/cross_species_XCI/final_plots/R/data_for_plots/mouse_haplotype_data.Rdata' )


```

########################################
Downsampling results for both cow and sheep. These are our 2 highest sampled species, sheep has a bunch of detected SNPs and Cow
has a lot less. Downsample to 300 samples and run the XCI variant association experiment, there are 313 human donors we sample over
Idea is to test how robust the genetic associations are to sample size variation. 
We currently get no significant results in humans, is it sampling limitation?

##########################################



```{r}
all_cow_downsample_df = data.frame(variants = c(), ref_index = c(),
                                 n_pos = c(), n_neg = c(),
                                 auc = c(), auc_pval = c(),
                                 auc_fdr_pval = c(),
                                 gene = c(), down_sample = c())
for(s in 1:10){
  #cow
  start = Sys.time()
  sample_index = which(cow_skew_and_stats_df$num_good_snps_no_escape >= 10 & cow_skew_and_stats_df$autosome_imbalance == F)
  #Get a sample of 300
  down_sample_index = sample(sample_index, size = 300)
  cow_results_down_1 = xci_skew_predict_variant_experiment(cow_skew_and_stats_df, annot_cow_vcfs, down_sample_index, cow_snp_stat_df)
  Sys.time() - start
  
  #Get the gene annotations for the variants
  cow_variants = cow_results_down_1$variants
  cow_snp_pos = as.numeric(sapply(1:length(cow_variants), 
                                         function(i) substr(cow_variants[i],1,nchar(cow_variants[i])-2)))
  cow_x_meta = cow_x_gtf[cow_x_gtf$gbkey == 'Gene'& !is.na(cow_x_gtf$gbkey),]
  x_gene_starts = cow_x_meta$start
  x_gene_ends = cow_x_meta$end
  #Add gene annotations
  cow_gene_annots = vector(mode = 'character', length = length(cow_snp_pos))
  for( i in 1:length(cow_snp_pos)){
  
    test_gene_start = cow_snp_pos[i]
    gene_index = findInterval(test_gene_start, x_gene_starts)
    if(gene_index == 0){next}
    
    if(test_gene_start <= x_gene_ends[gene_index]){
      cow_gene_annots[i] = cow_x_meta$gene[gene_index]
    }
  }
  #Put it all in a dataframe
  cow_temp_results_df = data.frame(variants = cow_results_down_1$variants, ref_index = cow_results_down_1$reference_biased_index,
                                   n_pos = cow_results_down_1$n_pos, n_neg = cow_results_down_1$n_neg,
                                   auc = cow_results_down_1$auc, auc_pval = cow_results_down_1$auc_pval,
                                   auc_fdr_pval = p.adjust(cow_results_down_1$auc_pval, method = 'BH'),
                                   gene = cow_gene_annots, down_sample = rep(sprintf('Sample %s', s), 
                                                                             length = length(cow_results_down_1$variants)))
  
  #Add to the complete dataframe
  all_cow_downsample_df = rbind(all_cow_downsample_df,cow_temp_results_df  )

}

all_cow_downsample_df

gene_hits_down_cow = all_cow_downsample_df %>% filter(auc >= .75 & auc_fdr_pval <= .05 & ref_index == F) %>% group_by(down_sample, gene) %>% summarize(n = length(gene)) %>% group_by(gene) %>% summarize(n = length(gene)) %>% arrange(desc(n))

gene_hits_down_cow$gene = factor(gene_hits_down_cow$gene, levels = c(gene_hits_down_cow$gene))
ggplot(gene_hits_down_cow, aes(x = gene, y = n)) + geom_col() + theme(axis.text.x = element_text(angle = 90)) +
  ylab('Number of downsamples') + xlab('Genes with skew associated variants') + ggtitle('Cow')
```



```{r}

save(all_cow_downsample_df, 
     file = '/home/werner/projects/cross_species_XCI/data/species_data_analysis/snp_skew_auc/all_cow_downsample_df.Rdata')

```


```{r}

load('/home/werner/projects/cross_species_XCI/data/species_data_analysis/snp_skew_auc/all_cow_downsample_df.Rdata')

```


```{r}

ori_variant_data = filter(snp_skew_auc_with_meta_df, label == 'Cow' & gene == 'KDM6A')
downSamp_variant_data = filter(all_cow_downsample_df, gene == 'KDM6A') %>% group_by(variants) %>% summarize(mean_auc = mean(auc), max_auc = max(auc))

index = match(ori_variant_data$variants, downSamp_variant_data$variants)
downSamp_variant_data = downSamp_variant_data[index, ]
downSamp_variant_data[ ,c(2,3)][is.na(downSamp_variant_data[ ,c(2,3)])] = 0

par(pty = 's')
plot(ori_variant_data$skew_auc, downSamp_variant_data$mean_auc, ylim = c(0, 1), xlim = c(0, 1), ylab = "Mean AUC of downsampled data", xlab='Original AUC of variant',
     main = 'Cow: KDM6A' )
abline(v = .75, col = 'red')
abline(h = .75, col = 'red')

pdf(file = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/human_experiments/cow_downsample_KDM6A.pdf', height = 4, width = 4, useDingbats = F)
par(pty = 's')
plot(ori_variant_data$skew_auc, downSamp_variant_data$mean_auc, ylim = c(0, 1), xlim = c(0, 1), ylab = "Mean AUC of downsampled data", xlab='Original AUC of variant',
     main = 'Cow: KDM6A' )
abline(v = .75, col = 'red')
abline(h = .75, col = 'red')
dev.off()



ori_variant_data = filter(snp_skew_auc_with_meta_df, label == 'Cow' & gene == 'TMEM164')
downSamp_variant_data = filter(all_cow_downsample_df, gene == 'TMEM164') %>% group_by(variants) %>% summarize(mean_auc = mean(auc), max_auc = max(auc))

index = match(ori_variant_data$variants, downSamp_variant_data$variants)
downSamp_variant_data = downSamp_variant_data[index, ]
downSamp_variant_data[ ,c(2,3)][is.na(downSamp_variant_data[ ,c(2,3)])] = 0

par(pty = 's')
plot(ori_variant_data$skew_auc, downSamp_variant_data$mean_auc, ylim = c(0, 1), xlim = c(0, 1), ylab = "Mean AUC of downsampled data", xlab='Original AUC of variant',
     main = 'Cow: TMEM164' )
abline(v = .75, col = 'red')
abline(h = .75, col = 'red')

pdf(file = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/human_experiments/cow_downsample_TMEM164.pdf', height = 4, width = 4, useDingbats = F)
par(pty = 's')
plot(ori_variant_data$skew_auc, downSamp_variant_data$mean_auc, ylim = c(0, 1), xlim = c(0, 1), ylab = "Mean AUC of downsampled data", xlab='Original AUC of variant',
     main = 'Cow: TMEM164' )
abline(v = .75, col = 'red')
abline(h = .75, col = 'red')
dev.off()



ori_variant_data = filter(snp_skew_auc_with_meta_df, label == 'Cow' & gene == 'MAMLD1')
downSamp_variant_data = filter(all_cow_downsample_df, gene == 'MAMLD1') %>% group_by(variants) %>% summarize(mean_auc = mean(auc), max_auc = max(auc))

index = match(ori_variant_data$variants, downSamp_variant_data$variants)
downSamp_variant_data = downSamp_variant_data[index, ]
downSamp_variant_data[ ,c(2,3)][is.na(downSamp_variant_data[ ,c(2,3)])] = 0

par(pty = 's')
plot(ori_variant_data$skew_auc, downSamp_variant_data$mean_auc, ylim = c(0, 1), xlim = c(0, 1), ylab = "Mean AUC of downsampled data", xlab='Original AUC of variant',
     main = 'Cow: MAMLD1' )
abline(v = .75, col = 'red')
abline(h = .75, col = 'red')

pdf(file = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/human_experiments/cow_downsample_MAMLD1.pdf', height = 4, width = 4, useDingbats = F)
par(pty = 's')
plot(ori_variant_data$skew_auc, downSamp_variant_data$mean_auc, ylim = c(0, 1), xlim = c(0, 1), ylab = "Mean AUC of downsampled data", xlab='Original AUC of variant',
     main = 'Cow: MAMLD1' )
abline(v = .75, col = 'red')
abline(h = .75, col = 'red')
dev.off()



ori_variant_data = filter(snp_skew_auc_with_meta_df, label == 'Cow' & gene == 'MID2')
downSamp_variant_data = filter(all_cow_downsample_df, gene == 'MID2') %>% group_by(variants) %>% summarize(mean_auc = mean(auc), max_auc = max(auc))

index = match(ori_variant_data$variants, downSamp_variant_data$variants)
downSamp_variant_data = downSamp_variant_data[index, ]
downSamp_variant_data[ ,c(2,3)][is.na(downSamp_variant_data[ ,c(2,3)])] = 0

par(pty = 's')
plot(ori_variant_data$skew_auc, downSamp_variant_data$mean_auc, ylim = c(0, 1), xlim = c(0, 1), ylab = "Mean AUC of downsampled data", xlab='Original AUC of variant',
     main = 'Cow: MID2' )
abline(v = .75, col = 'red')
abline(h = .75, col = 'red')

pdf(file = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/human_experiments/cow_downsample_MID2.pdf', height = 4, width = 4, useDingbats = F)
par(pty = 's')
plot(ori_variant_data$skew_auc, downSamp_variant_data$mean_auc, ylim = c(0, 1), xlim = c(0, 1), ylab = "Mean AUC of downsampled data", xlab='Original AUC of variant',
     main = 'Cow: MID2' )
abline(v = .75, col = 'red')
abline(h = .75, col = 'red')
dev.off()
```



Checking out the Snps associated with XCI skews in horse
```{r}

snp_skew_auc_with_meta_df %>% filter(label == 'Horse', skew_auc >= .75 )


```


```{r}

get_var_data = function(filt_annot_vcfs, variant_pos, tissue_skews){
  
  variant_present = vector(mode = 'logical', length = length(filt_annot_vcfs))
  variant_skew = c()
  for(i in 1:length(filt_annot_vcfs)){
    #data_variants = paste0(filt_annot_vcfs[[i]]$Pos, paste0(filt_annot_vcfs[[i]]$ref, filt_annot_vcfs[[i]]$alt))
    
    variant_present[i] = variant_pos %in% filt_annot_vcfs[[i]]$Pos
    if(variant_pos %in% filt_annot_vcfs[[i]]$Pos){
      data_var_skews = folded(filt_annot_vcfs[[i]]$ref_counts / (filt_annot_vcfs[[i]]$ref_counts + filt_annot_vcfs[[i]]$alt_counts))
      variant_skew = c(variant_skew, data_var_skews[filt_annot_vcfs[[i]]$Pos == variant_pos])
    }
  }
  
  var_data_df = data.frame(skews = c(tissue_skews[variant_present], variant_skew), variant = rep(variant_pos, length = (sum(variant_present)*2)),
                           data_label = c(rep('Tissue skew', length = sum(variant_present)), rep('Variant skew', length = sum(variant_present))))
  return(var_data_df)
}

filter(horse_snp_stat_df, gene == 'ARHGAP6' & num_samples_present >= 10) %>% arrange(desc(num_samples_present))
```



```{r}

sample_index = which(horse_skew_and_stats_df$num_good_snps_no_escape >= 10 & horse_skew_and_stats_df$autosome_imbalance == F)
filt_annot_horse_vcfs = annot_horse_vcfs[sample_index]
horse_XCI_skews = horse_skew_and_stats_df$est_skew_no_escape[sample_index]

#ARHGAP6
arhgap6_var_1_df =  get_var_data(filt_annot_horse_vcfs,'8077217', horse_XCI_skews)
arhgap6_var_2_df =  get_var_data(filt_annot_horse_vcfs,'8077541', horse_XCI_skews)
arhgap6_var_3_df =  get_var_data(filt_annot_horse_vcfs,'8083294', horse_XCI_skews)

comb_arhgap6_df = Reduce(rbind, list(arhgap6_var_1_df, arhgap6_var_2_df, arhgap6_var_3_df))
ggplot(comb_arhgap6_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('Horse ARHGAP6') + ylab('XCI skews') + ggtitle('Horse ARHGAP6')
#ggsave(filename = 'horse_arhgap6_variants.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/variant_escape_examples/', 
#       device = 'pdf', height = 4, width = 6, useDingbats = F)


#GPC4
gpc4_var_1_df =  get_var_data(filt_annot_horse_vcfs,'110031329', horse_XCI_skews)
gpc4_var_2_df =  get_var_data(filt_annot_horse_vcfs,'110031201', horse_XCI_skews)
gpc4_var_3_df =  get_var_data(filt_annot_horse_vcfs,'110031506', horse_XCI_skews)
gpc4_var_4_df =  get_var_data(filt_annot_horse_vcfs,'110031002', horse_XCI_skews)
gpc4_var_5_df =  get_var_data(filt_annot_horse_vcfs,'110031664', horse_XCI_skews)

comb_gpc4_df = Reduce(rbind, list(gpc4_var_1_df,gpc4_var_2_df, gpc4_var_3_df, gpc4_var_4_df, gpc4_var_5_df))
ggplot(comb_gpc4_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('Horse GPC4') + ylab('XCI skews') + ggtitle('Horse GPC4')
#ggsave(filename = 'horse_gpc4_variants.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/variant_escape_examples/', 
#       device = 'pdf', height = 4, width = 6, useDingbats = F)


#PLCXD1
plcxd1_var_1_df =  get_var_data(filt_annot_horse_vcfs,'32437', horse_XCI_skews)
plcxd1_var_2_df =  get_var_data(filt_annot_horse_vcfs,'32442', horse_XCI_skews)
plcxd1_var_3_df =  get_var_data(filt_annot_horse_vcfs,'32592', horse_XCI_skews)
plcxd1_var_4_df =  get_var_data(filt_annot_horse_vcfs,'33358', horse_XCI_skews)
plcxd1_var_5_df =  get_var_data(filt_annot_horse_vcfs,'33906', horse_XCI_skews)
plcxd1_var_6_df =  get_var_data(filt_annot_horse_vcfs,'34231', horse_XCI_skews)
plcxd1_var_7_df =  get_var_data(filt_annot_horse_vcfs,'30849', horse_XCI_skews)

comb_plcxd1_df = Reduce(rbind, list(plcxd1_var_1_df,plcxd1_var_2_df,plcxd1_var_3_df,plcxd1_var_4_df,plcxd1_var_5_df,plcxd1_var_6_df,plcxd1_var_7_df))
ggplot(comb_plcxd1_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('Horse PLCXD1') + ylab('XCI skews') + ggtitle('Horse PLCXD1')
#ggsave(filename = 'horse_plcxd1_variants.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/variant_escape_examples/', 
#       device = 'pdf', height = 4, width = 6, useDingbats = F)


#ARSE
arse_var_1_df =  get_var_data(filt_annot_horse_vcfs,'1218988', horse_XCI_skews)
arse_var_2_df =  get_var_data(filt_annot_horse_vcfs,'1218989', horse_XCI_skews)

comb_arse_df = Reduce(rbind, list(arse_var_1_df, arse_var_2_df))
ggplot(comb_arse_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('Horse ARSE') + ylab('XCI skews') + ggtitle('Horse ARSE')
#ggsave(filename = 'horse_arse_variants.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/variant_escape_examples/', 
#       device = 'pdf', height = 4, width = 6, useDingbats = F)

save(comb_arse_df, comb_plcxd1_df, comb_gpc4_df, comb_arhgap6_df, 
     file = '/home/werner/projects/cross_species_XCI/final_plots/R/data_for_plots/horse_individual_variant_auc_boxplots.Rdata')


```

Checking out the variants associated with skew in mice

```{r}

snp_skew_auc_with_meta_df %>% filter(label == 'Mouse', skew_auc >= .75 )

```


102619182 is between Mir374c (miRNA) and Gm9159 (pseudogene)
104379565 is between GM14824 (pseudogene) and Mir384 (miRNA)

```{r}

filter(mouse_snp_stat_df, snp_pos == 102619182 & num_samples_present >= 10) %>% arrange(desc(num_samples_present))
filter(mouse_snp_stat_df, snp_pos == 104379565 & num_samples_present >= 10) %>% arrange(desc(num_samples_present))


filter(mouse_x_meta, start >= 104370000)
```



```{r}

sample_index = which(mouse_skew_and_stats_df$num_good_snps_no_escape >= 10 & mouse_skew_and_stats_df$autosome_imbalance == F)
filt_annot_mouse_vcfs = annot_mouse_vcfs[sample_index]
mouse_XCI_skews = mouse_skew_and_stats_df$est_skew_no_escape[sample_index]

#only 2 associated mouse variants are not in an annotated gene
mouse_var_1_df =  get_var_data(filt_annot_mouse_vcfs,'102619182', mouse_XCI_skews)
mouse_var_2_df =  get_var_data(filt_annot_mouse_vcfs,'104379565', mouse_XCI_skews)


comb_test_df = rbind(mouse_var_1_df, mouse_var_2_df)
ggplot(comb_test_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('Mouse variants') + ylab('XCI skews') + ggtitle('Mouse variants')

```





Checking out the Snps associated with XCI skews in goat
```{r}

snp_skew_auc_with_meta_df %>% filter(label == 'Goat', skew_auc >= .75 )

```

```{r}
filter(goat_snp_stat_df, gene == 'LOC102168615' & num_samples_present >= 10) %>% arrange(desc(num_samples_present))

```

```{r}
sample_index = which(goat_skew_and_stats_df$num_good_snps_no_escape >= 10 & goat_skew_and_stats_df$autosome_imbalance == F)
filt_annot_goat_vcfs = annot_goat_vcfs[sample_index]
goat_XCI_skews = goat_skew_and_stats_df$est_skew_no_escape[sample_index]

#MSL3
msl3_var_1_df =  get_var_data(filt_annot_goat_vcfs,'58365038', goat_XCI_skews)
msl3_var_2_df =  get_var_data(filt_annot_goat_vcfs,'58375993', goat_XCI_skews)
msl3_var_3_df =  get_var_data(filt_annot_goat_vcfs,'58377919', goat_XCI_skews)
msl3_var_4_df =  get_var_data(filt_annot_goat_vcfs,'58364846', goat_XCI_skews)

comb_msl3_df = Reduce(rbind, list(msl3_var_1_df, msl3_var_2_df, msl3_var_3_df, msl3_var_4_df))
ggplot(comb_msl3_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('goat MSL3') + ylab('XCI skews') + ggtitle('goat MSL3')
#ggsave(filename = 'goat_msl3_variants.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/variant_escape_examples/', 
#       device = 'pdf', height = 4, width = 6, useDingbats = F)

#MXRA5
MXRA5_var_1_df =  get_var_data(filt_annot_goat_vcfs,'65380180', goat_XCI_skews)
MXRA5_var_2_df =  get_var_data(filt_annot_goat_vcfs,'65381048', goat_XCI_skews)
MXRA5_var_3_df =  get_var_data(filt_annot_goat_vcfs,'65394917', goat_XCI_skews)
MXRA5_var_4_df =  get_var_data(filt_annot_goat_vcfs,'65395668', goat_XCI_skews)
MXRA5_var_5_df =  get_var_data(filt_annot_goat_vcfs,'65393424', goat_XCI_skews)
MXRA5_var_6_df =  get_var_data(filt_annot_goat_vcfs,'65393664', goat_XCI_skews)
MXRA5_var_7_df =  get_var_data(filt_annot_goat_vcfs,'65394420', goat_XCI_skews)
MXRA5_var_8_df =  get_var_data(filt_annot_goat_vcfs,'65393460', goat_XCI_skews)

comb_mxra5_df = Reduce(rbind, list(MXRA5_var_1_df, MXRA5_var_2_df, MXRA5_var_3_df, MXRA5_var_4_df, MXRA5_var_5_df, MXRA5_var_6_df, MXRA5_var_7_df, MXRA5_var_8_df))
ggplot(comb_mxra5_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('goat MXRA5') + ylab('XCI skews') + ggtitle('goat MXRA5')
#ggsave(filename = 'goat_mxra5_variants.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/variant_escape_examples/', 
#       device = 'pdf', height = 4, width = 6, useDingbats = F)

#SHROOM2
SHROOM2_var_1_df =  get_var_data(filt_annot_goat_vcfs,'60903653', goat_XCI_skews)
SHROOM2_var_2_df =  get_var_data(filt_annot_goat_vcfs,'60919507', goat_XCI_skews)
SHROOM2_var_3_df =  get_var_data(filt_annot_goat_vcfs,'60919917', goat_XCI_skews)
SHROOM2_var_4_df =  get_var_data(filt_annot_goat_vcfs,'60905961', goat_XCI_skews)
SHROOM2_var_5_df =  get_var_data(filt_annot_goat_vcfs,'60899151', goat_XCI_skews)
SHROOM2_var_6_df =  get_var_data(filt_annot_goat_vcfs,'60867055', goat_XCI_skews)
SHROOM2_var_7_df =  get_var_data(filt_annot_goat_vcfs,'60867365', goat_XCI_skews)
SHROOM2_var_8_df =  get_var_data(filt_annot_goat_vcfs,'60917891', goat_XCI_skews)

comb_shroom2_df = Reduce(rbind, list(SHROOM2_var_1_df, SHROOM2_var_2_df, SHROOM2_var_3_df, SHROOM2_var_4_df, SHROOM2_var_5_df, SHROOM2_var_6_df, SHROOM2_var_7_df, SHROOM2_var_8_df))
ggplot(comb_shroom2_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('goat SHROOM2') + ylab('XCI skews') + ggtitle('goat SHROOM2')
#ggsave(filename = 'goat_shroom2_variants.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/variant_escape_examples/', 
#       device = 'pdf', height = 4, width = 6, useDingbats = F)

#loc102
loc102_var_1_df =  get_var_data(filt_annot_goat_vcfs,'8677070', goat_XCI_skews)
loc102_var_2_df =  get_var_data(filt_annot_goat_vcfs,'8684125', goat_XCI_skews)
loc102_var_3_df =  get_var_data(filt_annot_goat_vcfs,'8716897', goat_XCI_skews)
loc102_var_4_df =  get_var_data(filt_annot_goat_vcfs,'8707809', goat_XCI_skews)
loc102_var_5_df =  get_var_data(filt_annot_goat_vcfs,'8801309', goat_XCI_skews)
loc102_var_6_df =  get_var_data(filt_annot_goat_vcfs,'8677624', goat_XCI_skews)
loc102_var_7_df =  get_var_data(filt_annot_goat_vcfs,'8685099', goat_XCI_skews)
loc102_var_8_df =  get_var_data(filt_annot_goat_vcfs,'8677596', goat_XCI_skews)

comb_loc102_df = Reduce(rbind, list(loc102_var_1_df, loc102_var_2_df, loc102_var_3_df, loc102_var_4_df, loc102_var_5_df, loc102_var_6_df, loc102_var_7_df, loc102_var_8_df))
ggplot(comb_loc102_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('goat LOC102168615') + ylab('XCI skews') + ggtitle('goat LOC102168615')
#ggsave(filename = 'goat_loc102_variants.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/variant_escape_examples/', 
#       device = 'pdf', height = 4, width = 6, useDingbats = F)

save(comb_msl3_df, comb_mxra5_df, comb_shroom2_df, comb_loc102_df, 
     file = '/home/werner/projects/cross_species_XCI/final_plots/R/data_for_plots/goat_individual_variant_auc_boxplots.Rdata')



```


Checking out the Snps associated with XCI skews in cow
```{r}

snp_skew_auc_with_meta_df %>% filter(label == 'Cow', skew_auc >= .75 )

```

```{r}
filter(cow_snp_stat_df, gene == 'KDM6A' & num_samples_present >= 10) %>% arrange(desc(num_samples_present))

```

```{r}

sample_index = which(cow_skew_and_stats_df$num_good_snps_no_escape >= 10 & cow_skew_and_stats_df$autosome_imbalance == F)
filt_annot_cow_vcfs = annot_cow_vcfs[sample_index]
cow_XCI_skews = cow_skew_and_stats_df$est_skew_no_escape[sample_index]


#MAMLD1
MAMLD1_var_1_df =  get_var_data(filt_annot_cow_vcfs,'33106543', cow_XCI_skews)
MAMLD1_var_2_df =  get_var_data(filt_annot_cow_vcfs,'33135236', cow_XCI_skews)
MAMLD1_var_3_df =  get_var_data(filt_annot_cow_vcfs,'33114156', cow_XCI_skews)
MAMLD1_var_4_df =  get_var_data(filt_annot_cow_vcfs,'33141579', cow_XCI_skews)
MAMLD1_var_5_df =  get_var_data(filt_annot_cow_vcfs,'33141584', cow_XCI_skews)
MAMLD1_var_6_df =  get_var_data(filt_annot_cow_vcfs,'33128071', cow_XCI_skews)
MAMLD1_var_7_df =  get_var_data(filt_annot_cow_vcfs,'33134000', cow_XCI_skews)
MAMLD1_var_8_df =  get_var_data(filt_annot_cow_vcfs,'33138476', cow_XCI_skews)

comb_mamld1_df = Reduce(rbind, list(MAMLD1_var_1_df, MAMLD1_var_2_df, MAMLD1_var_3_df, MAMLD1_var_4_df, MAMLD1_var_5_df, MAMLD1_var_6_df, MAMLD1_var_7_df, MAMLD1_var_8_df))
ggplot(comb_mamld1_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('cow MAMLD1') + ylab('XCI skews') + ggtitle('cow MAMLD1')
#ggsave(filename = 'cow_mamld1_variants.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/variant_escape_examples/', 
#       device = 'pdf', height = 4, width = 6, useDingbats = F)


#MID2
MID2_var_1_df =  get_var_data(filt_annot_cow_vcfs,'56469706', cow_XCI_skews)
MID2_var_2_df =  get_var_data(filt_annot_cow_vcfs,'56492919', cow_XCI_skews)
MID2_var_3_df =  get_var_data(filt_annot_cow_vcfs,'56557471', cow_XCI_skews)
MID2_var_4_df =  get_var_data(filt_annot_cow_vcfs,'56558220', cow_XCI_skews)
MID2_var_5_df =  get_var_data(filt_annot_cow_vcfs,'56555624', cow_XCI_skews)
MID2_var_6_df =  get_var_data(filt_annot_cow_vcfs,'56471967', cow_XCI_skews)
MID2_var_7_df =  get_var_data(filt_annot_cow_vcfs,'56470399', cow_XCI_skews)
MID2_var_8_df =  get_var_data(filt_annot_cow_vcfs,'56491067', cow_XCI_skews)

comb_mid2_df = Reduce(rbind, list(MID2_var_1_df, MID2_var_2_df, MID2_var_3_df, MID2_var_4_df, MID2_var_5_df, MID2_var_6_df, MID2_var_7_df, MID2_var_8_df))
ggplot(comb_mid2_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('cow MID2') + ylab('XCI skews') + ggtitle('cow MID2')
#ggsave(filename = 'cow_mid2_variants.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/variant_escape_examples/', 
#       device = 'pdf', height = 4, width = 6, useDingbats = F)

#TMEM164
TMEM164_var_1_df =  get_var_data(filt_annot_cow_vcfs,'58445235', cow_XCI_skews)
TMEM164_var_2_df =  get_var_data(filt_annot_cow_vcfs,'58445236', cow_XCI_skews)
TMEM164_var_3_df =  get_var_data(filt_annot_cow_vcfs,'58424460', cow_XCI_skews)
TMEM164_var_4_df =  get_var_data(filt_annot_cow_vcfs,'58434987', cow_XCI_skews)
TMEM164_var_5_df =  get_var_data(filt_annot_cow_vcfs,'58435002', cow_XCI_skews)
TMEM164_var_6_df =  get_var_data(filt_annot_cow_vcfs,'58432292', cow_XCI_skews)
TMEM164_var_7_df =  get_var_data(filt_annot_cow_vcfs,'58466645', cow_XCI_skews)
TMEM164_var_8_df =  get_var_data(filt_annot_cow_vcfs,'58441229', cow_XCI_skews)

comb_tmem164_df = Reduce(rbind, list(TMEM164_var_1_df, TMEM164_var_2_df, TMEM164_var_3_df, TMEM164_var_4_df, TMEM164_var_5_df, TMEM164_var_6_df, TMEM164_var_7_df, TMEM164_var_8_df))
ggplot(comb_tmem164_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('cow TMEM164') + ylab('XCI skews') + ggtitle('cow TMEM164')
#ggsave(filename = 'cow_tmem164_variants.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/variant_escape_examples/', 
#       device = 'pdf', height = 4, width = 6, useDingbats = F)

#KDM6A
KDM6A_var_1_df =  get_var_data(filt_annot_cow_vcfs,'98666547', cow_XCI_skews)
KDM6A_var_2_df =  get_var_data(filt_annot_cow_vcfs,'98670561', cow_XCI_skews)
KDM6A_var_3_df =  get_var_data(filt_annot_cow_vcfs,'98666128', cow_XCI_skews)
KDM6A_var_4_df =  get_var_data(filt_annot_cow_vcfs,'98679560', cow_XCI_skews)
KDM6A_var_5_df =  get_var_data(filt_annot_cow_vcfs,'98802379', cow_XCI_skews)
KDM6A_var_6_df =  get_var_data(filt_annot_cow_vcfs,'98801035', cow_XCI_skews)
KDM6A_var_7_df =  get_var_data(filt_annot_cow_vcfs,'98685885', cow_XCI_skews)
KDM6A_var_8_df =  get_var_data(filt_annot_cow_vcfs,'98785159', cow_XCI_skews)

comb_kdm6a_df = Reduce(rbind, list(KDM6A_var_1_df, KDM6A_var_2_df, KDM6A_var_3_df, KDM6A_var_4_df, KDM6A_var_5_df, KDM6A_var_6_df, KDM6A_var_7_df, KDM6A_var_8_df))
ggplot(comb_kdm6a_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('cow KDM6A') + ylab('XCI skews') + ggtitle('cow KDM6A')
#ggsave(filename = 'cow_kdm6a_variants.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/variant_escape_examples/', 
#       device = 'pdf', height = 4, width = 6, useDingbats = F)

save(comb_mamld1_df, comb_mid2_df, comb_tmem164_df, comb_kdm6a_df, 
     file = '/home/werner/projects/cross_species_XCI/final_plots/R/data_for_plots/cow_individual_variant_auc_boxplots.Rdata')


```




Checking out the Snps associated with XCI skews in sheep
```{r}

snp_skew_auc_with_meta_df %>% filter(label == 'Sheep', skew_auc >= .75 )

```

```{r}
filter(sheep_snp_stat_df, gene == 'IL3RA' & num_samples_present >= 10) %>% arrange(desc(num_samples_present))

```

```{r}
sample_index = which(sheep_skew_and_stats_df$num_good_snps_no_escape >= 10 & sheep_skew_and_stats_df$autosome_imbalance == F)
filt_annot_sheep_vcfs = annot_sheep_vcfs[sample_index]
sheep_XCI_skews = sheep_skew_and_stats_df$est_skew_no_escape[sample_index]


#LAS1L
LAS1L_var_1_df =  get_var_data(filt_annot_sheep_vcfs,'47731599', sheep_XCI_skews)
LAS1L_var_2_df =  get_var_data(filt_annot_sheep_vcfs,'47731971', sheep_XCI_skews)
LAS1L_var_3_df =  get_var_data(filt_annot_sheep_vcfs,'47730518', sheep_XCI_skews)
LAS1L_var_4_df =  get_var_data(filt_annot_sheep_vcfs,'47725877', sheep_XCI_skews)
LAS1L_var_5_df =  get_var_data(filt_annot_sheep_vcfs,'47728201', sheep_XCI_skews)
LAS1L_var_6_df =  get_var_data(filt_annot_sheep_vcfs,'47713895', sheep_XCI_skews)
LAS1L_var_7_df =  get_var_data(filt_annot_sheep_vcfs,'47732094', sheep_XCI_skews)
LAS1L_var_8_df =  get_var_data(filt_annot_sheep_vcfs,'47726456', sheep_XCI_skews)

comb_test_df = Reduce(rbind, list(LAS1L_var_1_df, LAS1L_var_2_df, LAS1L_var_3_df, LAS1L_var_4_df, LAS1L_var_5_df, LAS1L_var_6_df, LAS1L_var_7_df, LAS1L_var_8_df))
ggplot(comb_test_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('sheep LAS1L') + ylab('XCI skews') + ggtitle('sheep LAS1L')



#MXRA5
MXRA5_var_1_df =  get_var_data(filt_annot_sheep_vcfs,'5797063', sheep_XCI_skews)
MXRA5_var_2_df =  get_var_data(filt_annot_sheep_vcfs,'5797102', sheep_XCI_skews)
MXRA5_var_3_df =  get_var_data(filt_annot_sheep_vcfs,'5814806', sheep_XCI_skews)
MXRA5_var_4_df =  get_var_data(filt_annot_sheep_vcfs,'5814449', sheep_XCI_skews)
MXRA5_var_5_df =  get_var_data(filt_annot_sheep_vcfs,'5814644', sheep_XCI_skews)
MXRA5_var_6_df =  get_var_data(filt_annot_sheep_vcfs,'5814101', sheep_XCI_skews)
MXRA5_var_7_df =  get_var_data(filt_annot_sheep_vcfs,'5805814', sheep_XCI_skews)
MXRA5_var_8_df =  get_var_data(filt_annot_sheep_vcfs,'5797399', sheep_XCI_skews)

comb_mxra5_df = Reduce(rbind, list(MXRA5_var_1_df, MXRA5_var_2_df, MXRA5_var_3_df, MXRA5_var_4_df, MXRA5_var_5_df, MXRA5_var_6_df, MXRA5_var_7_df, MXRA5_var_8_df))
ggplot(comb_mxra5_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('sheep MXRA5') + ylab('XCI skews') + ggtitle('sheep MXRA5')
#ggsave(filename = 'sheep_mxra5_variants.pdf', path = '/home/werner/projects/cross_species_XCI/code/graphs/skew_predict_snp/variant_escape_examples/', 
#       device = 'pdf', height = 4, width = 6, useDingbats = F)

#PRKX
PRKX_var_1_df =  get_var_data(filt_annot_sheep_vcfs,'5660010', sheep_XCI_skews)
PRKX_var_2_df =  get_var_data(filt_annot_sheep_vcfs,'5658562', sheep_XCI_skews)
PRKX_var_3_df =  get_var_data(filt_annot_sheep_vcfs,'5660456', sheep_XCI_skews)
PRKX_var_4_df =  get_var_data(filt_annot_sheep_vcfs,'5660494', sheep_XCI_skews)
PRKX_var_5_df =  get_var_data(filt_annot_sheep_vcfs,'5660365', sheep_XCI_skews)
PRKX_var_6_df =  get_var_data(filt_annot_sheep_vcfs,'5659443', sheep_XCI_skews)
PRKX_var_7_df =  get_var_data(filt_annot_sheep_vcfs,'5657672', sheep_XCI_skews)
PRKX_var_8_df =  get_var_data(filt_annot_sheep_vcfs,'5661013', sheep_XCI_skews)

comb_test_df = Reduce(rbind, list(PRKX_var_1_df, PRKX_var_2_df, PRKX_var_3_df, PRKX_var_4_df, PRKX_var_5_df, PRKX_var_6_df, PRKX_var_7_df, PRKX_var_8_df))
ggplot(comb_test_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('sheep PRKX') + ylab('XCI skews') + ggtitle('sheep PRKX')


#CXCR3
CXCR3_var_1_df =  get_var_data(filt_annot_sheep_vcfs,'64273174', sheep_XCI_skews)
CXCR3_var_2_df =  get_var_data(filt_annot_sheep_vcfs,'64271204', sheep_XCI_skews)
CXCR3_var_3_df =  get_var_data(filt_annot_sheep_vcfs,'64270869', sheep_XCI_skews)


comb_test_df = Reduce(rbind, list(CXCR3_var_1_df, CXCR3_var_2_df, CXCR3_var_3_df))
ggplot(comb_test_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('sheep CXCR3') + ylab('XCI skews') + ggtitle('sheep CXCR3')


#IL3RA
IL3RA_var_1_df =  get_var_data(filt_annot_sheep_vcfs,'6648655', sheep_XCI_skews)
IL3RA_var_2_df =  get_var_data(filt_annot_sheep_vcfs,'6654816', sheep_XCI_skews)
IL3RA_var_3_df =  get_var_data(filt_annot_sheep_vcfs,'6658743', sheep_XCI_skews)
IL3RA_var_4_df =  get_var_data(filt_annot_sheep_vcfs,'6662529', sheep_XCI_skews)
IL3RA_var_5_df =  get_var_data(filt_annot_sheep_vcfs,'6648106', sheep_XCI_skews)
IL3RA_var_6_df =  get_var_data(filt_annot_sheep_vcfs,'6662912', sheep_XCI_skews)
IL3RA_var_7_df =  get_var_data(filt_annot_sheep_vcfs,'6657058', sheep_XCI_skews)
IL3RA_var_8_df =  get_var_data(filt_annot_sheep_vcfs,'6662926', sheep_XCI_skews)

comb_test_df = Reduce(rbind, list(IL3RA_var_1_df, IL3RA_var_2_df, IL3RA_var_3_df, IL3RA_var_4_df, IL3RA_var_5_df, IL3RA_var_6_df, IL3RA_var_7_df, IL3RA_var_8_df))
ggplot(comb_test_df, aes(y = skews, x = variant, fill = data_label)) + geom_boxplot() + xlab('sheep IL3RA') + ylab('XCI skews') + ggtitle('sheep IL3RA')

save(comb_mxra5_df,
     file = '/home/werner/projects/cross_species_XCI/final_plots/R/data_for_plots/sheep_individual_variant_auc_boxplots.Rdata')



```










